### 2024-04-08 README.qmd
Starter for 10

---
format: gfm
---

```{r}
library(tidyverse)
u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
download_and_read = function(u) {
    tmp = tempfile(fileext = ".xlsx")
    download.file(u, tmp, mode = "wb")
    readxl::read_excel(tmp)
}
d = download_and_read(u)
names(d)
#  [1] "REPORTER TRANSPORT MODE" "OFFENDER VEHICLE MAKE"  
#  [3] "OFFENDER VEHICLE MODEL"  "OFFENDER VEHICLE COLOUR"
#  [5] "OFFENCE"                 "DISTRICT"               
#  [7] "DISPOSAL"                "DATE OF SUBMISSION"     
#  [9] "...9"                    "OFF LOCATION"  
d_locations = d |>
  dplyr::pull(`OFF LOCATION`) |>
  head(10) |>
  lapply(FUN = stplanr::geo_code)
# d_geocoded = d |>
#   tidygeocoder::geocode("OFF LOCATION")
d_locations[[1]]
matrix(d_locations[[1]], ncol = 2)
sf::st_point(matrix(d_locations[[1]], ncol = 2))
d_locations_lengths = sapply(d_locations, function(x) {
    length(x)
    })

d_sf = lapply(d_locations[d_locations_lengths > 0], function(x) {
  sf::st_point(matrix(x, ncol = 2))
})
d_sfc = sf::st_sfc(d_sf) |>
    sf::st_sf(crs = "EPSG:4326")
d_sf = sf::st_sf(
    d |> slice(1:10) |> filter(d_locations_lengths > 0),
    geometry = d_sfc |> sf::st_geometry()
)

mapview::mapview(d_sf)
```

```{r}
table(d$OFFENCE) |>
  sort()
```

```{r}
d |>
  # Reduce nchar of OFFENCE
  mutate(OFFENCE = stringr::str_sub(OFFENCE, 1, 60)) |>
  group_by(OFFENCE) |>
  # Count number of rows in each group
  mutate(n = n()) |>
  filter(n > nrow(d) / 50)|>
  ggplot() +
  geom_bar(aes(OFFENCE)) +
  # Make x labels vertical
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### 2024-04-08 README.qmd
Add pkg basics

@@
 ```{r}
+#| echo: false
+#| eval: false
+# Package setup:
+usethis::use_description()
+usethis::use_package("sf")
+usethis::use_package("stplanr")
+usethis::use_package("ggplot2")
+usethis::use_package("dplyr")
+usethis::use_r("opsnap")
+```
+
+# Installation
+
+Install the package from GitHub:
+
+```{r}
+remotes::install_github("ITSLeeds/opsnap")
+```
+
+```{r}
 library(tidyverse)
 u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
-download_and_read = function(u) {
-    tmp = tempfile(fileext = ".xlsx")
-    download.file(u, tmp, mode = "wb")
-    readxl::read_excel(tmp)
-}
-d = download_and_read(u)
+d = opsnap:::download_and_read(u)
 names(d)

### 2024-04-08 R/opsnap.R

@@ -1,0 +1,5 @@
+download_and_read = function(u) {
+    tmp = tempfile(fileext = ".xlsx")
+    download.file(u, tmp, mode = "wb")
+    readxl::read_excel(tmp)
+}

### 2024-04-08 README.qmd
Update with cleaning

@@
 
-```{r}
+```r
 remotes::install_github("ITSLeeds/opsnap")
@@
 ```{r}
+#| include: false
 library(tidyverse)
+remotes::install_github("ITSLeeds/opsnap")
+```
+
+# Opsnap data
+
+The `opsnap` package provides a function to download and read in data from the West Yorkshire Police Operation Snap database.
+The data is available at the following URL: https://www.westyorkshire.police.uk/SaferRoadsSubmissions
+
+The data is open acess and looks like this:
+
+```{r}
 u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
@@
 #  [9] "...9"                    "OFF LOCATION"
+
+# Function to clean up column names
+
 d_locations = d |>

### 2024-04-08 R/opsnap.R

@@ -1,5 +1,28 @@
-download_and_read = function(u) {
+download_and_read = function(u, remove_nas = TRUE) {
     tmp = tempfile(fileext = ".xlsx")
     download.file(u, tmp, mode = "wb")
-    readxl::read_excel(tmp)
+    d = readxl::read_excel(tmp)
+    names(d) = clean_names(names(d))
+    d = select_columns(d)
+    d = d |> filter_nas()
+    return(d)
+}
+
+clean_names = function(x) {
+  x |>
+    gsub("REPORTER TRANSPORT ", "", .) |>
+    gsub("OFFENDER VEHICLE ", "", .) |>
+    gsub("OFF ", "", .) |>
+    gsub("DATE OF SUBMISSION", "DATE", .) |>
+    tolower()
+}
+
+select_columns = function(d) {
+  names_remove = stringr::str_detect(names(d), "\\.\\.")
+  d = d[, !names_remove]
+}
+
+filter_nas = function(d) {
+  d |>
+    dplyr::filter(offence == "N/A")
 }

### 2024-04-08 README.qmd
Fix typo in packaging

@@
 usethis::use_r("opsnap")
+
+# Add CI
+usethis::use_github_action_check_standard()
+# Add pkgdown
+usethis::use_pkgdown()
+# Add pkgdown action
+usethis::use_github_action("pkgdown")
+# Add gh pages:
+usethis::use_github_pages()
 ```
@@
 
+```{r}
+#| eval: false
+#| echo: false
+devtools::load_all()
+```
+
 # Opsnap data
@@
 
-The data is open acess and looks like this:
+The data is open acess and looks like this, with names cleaned up by the package:
 
@@
 names(d)
+# Old names:
 #  [1] "REPORTER TRANSPORT MODE" "OFFENDER VEHICLE MAKE"  
@@
 #  [9] "...9"                    "OFF LOCATION"
+```
 

### 2024-04-08 R/opsnap.R

@@ -1,6 +1,6 @@
 download_and_read = function(u, remove_nas = TRUE) {
     tmp = tempfile(fileext = ".xlsx")
-    download.file(u, tmp, mode = "wb")
+    utils::download.file(u, tmp, mode = "wb")
     d = readxl::read_excel(tmp)
     names(d) = clean_names(names(d))
     d = select_columns(d)
@@ -10,10 +10,10 @@ download_and_read = function(u, remove_nas = TRUE) {
 
 clean_names = function(x) {
   x |>
-    gsub("REPORTER TRANSPORT ", "", .) |>
-    gsub("OFFENDER VEHICLE ", "", .) |>
-    gsub("OFF ", "", .) |>
-    gsub("DATE OF SUBMISSION", "DATE", .) |>
+    gsub("REPORTER TRANSPORT ", "", x = _) |>
+    gsub("OFFENDER VEHICLE ", "", x = _) |>
+    gsub("OFF ", "", x = _) |>
+    gsub("DATE OF SUBMISSION", "DATE", x = _) |>
     tolower()
 }
 
@@ -25,4 +25,12 @@ select_columns = function(d) {
 filter_nas = function(d) {
   d |>
     dplyr::filter(offence == "N/A")
+}
+
+op_plot_offence = function(d) {
+  d |>
+    ggplot::ggplot() +
+    ggplot::geom_bar(ggplot2::aes(x = offence)) +
+    # Make x labels vertical
+    ggplot::theme(axis.text.x = element_text(angle = 90, hjust = 1))
 }

### 2024-04-08 README.qmd
No warnings

@@
 usethis::use_r("opsnap")
+# MIT license:
+usethis::use_mit_license("Leeds Institute for Transport Studies")
 
@@
 
+```{r}
 # Function to clean up column names
-
-d_locations = d |>
-  dplyr::pull(`OFF LOCATION`) |>
-  head(10) |>
-  lapply(FUN = stplanr::geo_code)
-# d_geocoded = d |>
-#   tidygeocoder::geocode("OFF LOCATION")
-d_locations[[1]]
-matrix(d_locations[[1]], ncol = 2)
-sf::st_point(matrix(d_locations[[1]], ncol = 2))
-d_locations_lengths = sapply(d_locations, function(x) {
-    length(x)
-    })
-
-d_sf = lapply(d_locations[d_locations_lengths > 0], function(x) {
-  sf::st_point(matrix(x, ncol = 2))
-})
-d_sfc = sf::st_sfc(d_sf) |>
-    sf::st_sf(crs = "EPSG:4326")
-d_sf = sf::st_sf(
-    d |> slice(1:10) |> filter(d_locations_lengths > 0),
-    geometry = d_sfc |> sf::st_geometry()
-)
-
+d_sample = d[1:50, ]
+d_sf = opsnap:::op_geocode(d_sample)
 mapview::mapview(d_sf)

### 2024-04-08 R/opsnap.R

@@ -28,9 +28,35 @@ filter_nas = function(d) {
 }
 
 op_plot_offence = function(d) {
-  d |>
-    ggplot::ggplot() +
-    ggplot::geom_bar(ggplot2::aes(x = offence)) +
+  ggplot2::ggplot(d) +
+    ggplot2::geom_bar(ggplot2::aes(x = offence)) +
     # Make x labels vertical
-    ggplot::theme(axis.text.x = element_text(angle = 90, hjust = 1))
+    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1))
+}
+
+op_geocode = function(d, region = "West Yorkshire") {
+    d_locations_string = d |>
+        dplyr::pull(location)
+    d_locations_string = paste(d_locations_string, region, sep = ", ")
+        
+    d_locations = lapply(d_locations_string, FUN = stplanr::geo_code)
+    
+    d_locations_lengths = sapply(d_locations, function(x) {
+        length(x)
+    })
+    d_is_geo = d_locations_lengths > 0
+    d_with_geo = d[d_is_geo, ]
+    d_locations = d_locations[d_is_geo]
+    
+    d_point = lapply(d_locations, function(x) {
+        sf::st_point(matrix(x, ncol = 2))
+    })
+    d_sfc = sf::st_sfc(d_point) |>
+        sf::st_sf(crs = "EPSG:4326")
+    d_sf = sf::st_sf(
+        d_with_geo,
+        geometry = d_sfc |> sf::st_geometry()
+    )
+    
+    return(d_sf)
 }

### 2024-04-08 README.qmd
Small fixes

@@
 
+The data looks like this (first 3 rows shown):
+
+```{r}
+d |>
+  head(3) |>
+  knitr::kable()
+```
+
 ```{r}
@@
 ```{r}
-table(d$OFFENCE) |>
+table(d$offence) |>
   sort()
@@
 d |>
-  # Reduce nchar of OFFENCE
-  mutate(OFFENCE = stringr::str_sub(OFFENCE, 1, 60)) |>
-  group_by(OFFENCE) |>
+  # Reduce nchar of offence
+  mutate(offence = stringr::str_sub(offence, 1, 60)) |>
+  group_by(offence) |>
   # Count number of rows in each group
@@
   ggplot() +
-  geom_bar(aes(OFFENCE)) +
+  geom_bar(aes(offence)) +
   # Make x labels vertical

### 2024-04-08 R/opsnap.R

@@ -1,7 +1,11 @@
 download_and_read = function(u, remove_nas = TRUE) {
-    tmp = tempfile(fileext = ".xlsx")
-    utils::download.file(u, tmp, mode = "wb")
-    d = readxl::read_excel(tmp)
+    tmp = file.path(tempdir(), basename(u))
+    if (!file.exists(tmp)) {
+        utils::download.file(u, tmp, mode = "wb")
+    }
+    suppressMessages({
+        d = readxl::read_excel(tmp)
+    })
     names(d) = clean_names(names(d))
     d = select_columns(d)
     d = d |> filter_nas()
@@ -24,7 +28,7 @@ select_columns = function(d) {
 
 filter_nas = function(d) {
   d |>
-    dplyr::filter(offence == "N/A")
+    dplyr::filter(!offence == "N/A")
 }
 
 op_plot_offence = function(d) {

### 2024-04-08 README.qmd
Update readme

@@
 
+We provide a function to geocode the records:
+
 ```{r}
@@
 
+You can query the data downloaded with `opsnap` functions, e.g. as follows (results not shown):
+
 ```{r}
+#| eval: false
 table(d$offence) |>
@@
 ```
+
+Let's make a plot of the data:
 
@@
   # Make x labels vertical
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 60, hjust = 1))
 

### 2024-04-08 README.qmd
Update with rmd check status
Add analyis

--- 2024-04-08 README.qmd
@@
 format: gfm
+execute: 
+  message: false
+  warning: false
+  echo: false
 ---
+
+<!-- badges: start -->
+[![R-CMD-check](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml)
+<!-- badges: end -->
 
@@
 #| include: false
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.path = "man/figures/README-",
+  out.width = "100%"
+)
 library(tidyverse)
@@
 
+Data for the following years are provided:
+
+```{r}
+#| echo: false
+library(rvest)
+query = ".file-link a"
+url = "https://www.westyorkshire.police.uk/SaferRoadsSubmissions"
+files_available = read_html(url) |>
+  html_nodes(query) |>
+  html_text()
+links_available = read_html(url) |>
+  html_nodes(query) |>
+  html_attr("href")
+urls = paste0("https://www.westyorkshire.police.uk", links_available)
+file_names = basename(links_available)
+tibble::tibble(file_names) |>
+  knitr::kable()
+
+# In preparation for reading-in all data:
+date_str = format(Sys.Date(), "%Y-%m")
+file_name = paste0("data/west-yorkshire/operation_snap_", date_str, ".csv")
+```
+
+```{r}
+#| eval: false
+#| echo: false
+# Download them to raw_data/west-yorkshire:
+dir.create("raw_data/west-yorkshire", recursive = TRUE, showWarnings = FALSE)
+pbapply::pblapply(urls, function(u) {
+  download.file(u, paste0("raw_data/west-yorkshire/", basename(u)))
+})
+
+# Test with 2nd url:
+u = urls[2]
+d2 = opsnap:::download_and_read(u)
+d_all = purrr::map_df(urls, opsnap:::download_and_read)
+dir.create("data/west-yorkshire", recursive = TRUE, showWarnings = FALSE)
+# Date to nearest month:
+write_csv(d_all, file_name)
+```
+
 The data is open acess and looks like this, with names cleaned up by the package:
@@
 ```{r}
+#| eval: false
+#| echo: true
 u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
@@
 #  [9] "...9"                    "OFF LOCATION"
+# New names:
+# [1] "mode"     "make"     "model"    "colour"   "offence"  "district" "disposal"
+# [8] "date"     "location"
 ```
 
+<!-- The data looks like this (first 3 rows shown): -->
+
 ```{r}
-# Function to clean up column names
-d_sample = d[1:50, ]
+#| echo: false
+d_all = read_csv(file_name)
+d_all |>
+  head(3) |>
+  knitr::kable()
+```
+
+There are `r nrow(d_all)` records in the data, with increasing numbers of records over time (average n. records per month shown below):
+
+```{r}
+d_all_monthly = d_all |>
+  mutate(month = lubridate::floor_date(date, "month")) |>
+  group_by(month) |>
+  summarise(n = n()) |>
+  mutate(records = "all")
+d = d_all |>
+  opsnap:::filter_nas()
+d_complete_monthly = d |>
+  mutate(month = lubridate::floor_date(date, "month")) |>
+  group_by(month) |>
+  summarise(n = n()) |>
+  mutate(records = "complete")
+d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
+d_monthly |>
+  ggplot() +
+  geom_line(aes(month, n, colour = records)) +
+  geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
+  labs(title = "Number of records in West Yorkshire Police Operation Snap data",
+       x = "Date",
+       y = "Number of records per month")
+```
+
+As shown in the graph above, some records `r round(nrow(d) / nrow(d_all) * 100, 1)`% have missing values for locations (these records did not classify as an offence).
+Filtering these out leaves `r nrow(d)` complete records.
+
+There are `r unique(d$location) |> length()` unique locations in the data, with the most common locations shown below:
+
+```{r}
+#| echo: false
+d |>
+  count(location, sort = TRUE) |>
+  mutate(percent_records = n / nrow(d)) |>
+  mutate(percent_records = scales::percent(percent_records)) |>
+  arrange(desc(n)) |>
+  head(10) |>
+  knitr::kable()
+```
+
+# Geocoding
+
+We provide a function to geocode the records:
+
+```{r}
+#| eval: false
+#| echo: true
+d_sample = d[1:5, ]
 d_sf = opsnap:::op_geocode(d_sample)
@@
 ```{r}
-table(d$OFFENCE) |>
+#| eval: false
+#| echo: false
+# Geocode all data:
+d_geocoded = opsnap:::op_geocode(d)
+d_geometries = d_geocoded |>
+  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
+d_sf = sf::st_sf(
+  d,
+  geometry = d_geometries$geometry
+)
+west_yorkshire = pct::get_pct_zones("west-yorkshire")
+sf::sf_use_s2(FALSE)
+d_sf_wy = d_sf[west_yorkshire, ]
+# Sanity check
+d_sf_wy |>
+  sample_n(1000) |>
+  # mapview::mapview()
+  plot()
+# Save to file:
+sf::write_sf(d_sf_wy, paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
+```
+
+# Analysis
+
+Due to inaccuracy in the geocoding, we only know the locations of the records to within around 500m of each crash (although we can link to specific roads).
+We'll present the geographic distribution of crashes using a 500m grid:
+
+```{r}
+d_sf = sf::read_sf(paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
+d_projected = sf::st_transform(d_sf, 27700)
+library(terra)
+raster_template = rast(ext(d_projected), res = 500)
+raster_count = rasterize(d_projected, raster_template, fun = "length")
+plot(raster_count)
+```
+
+The results show there is one outlier with a very high number of crashes. We can remove this and plot the data again:
+
+```{r}
+raster_count_no_outlier = raster_count |>
+  clamp(-Inf, 100)
+plot(raster_count_no_outlier)
+```
+
+<!-- You can query the data downloaded with `opsnap` functions, e.g. as follows (results not shown): -->
+
+```{r}
+#| eval: false
+#| echo: false
+table(d$offence) |>
   sort()
@@
 
+<!-- Let's make a plot of the data: -->
+
 ```{r}
+#| eval: false
+#| echo: false
 d |>
-  # Reduce nchar of OFFENCE
-  mutate(OFFENCE = stringr::str_sub(OFFENCE, 1, 60)) |>
-  group_by(OFFENCE) |>
+  # Reduce nchar of offence
+  mutate(offence = stringr::str_sub(offence, 1, 60)) |>
+  group_by(offence) |>
   # Count number of rows in each group
@@
   ggplot() +
-  geom_bar(aes(OFFENCE)) +
+  geom_bar(aes(offence)) +
   # Make x labels vertical
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 60, hjust = 1))
 

### 2024-04-08 R/opsnap.R

@@ -1,5 +1,5 @@
-download_and_read = function(u, remove_nas = TRUE) {
-    tmp = file.path(tempdir(), basename(u))
+download_and_read = function(u, remove_nas = TRUE, dir = "raw_data/west-yorkshire", filter_nas = FALSE) {
+    tmp = file.path(dir, basename(u))
     if (!file.exists(tmp)) {
         utils::download.file(u, tmp, mode = "wb")
     }
@@ -8,7 +8,9 @@ download_and_read = function(u, remove_nas = TRUE) {
     })
     names(d) = clean_names(names(d))
     d = select_columns(d)
-    d = d |> filter_nas()
+    if (filter_nas) {
+        d = filter_nas(d)
+    }
     return(d)
 }
 
@@ -17,6 +19,8 @@ clean_names = function(x) {
     gsub("REPORTER TRANSPORT ", "", x = _) |>
     gsub("OFFENDER VEHICLE ", "", x = _) |>
     gsub("OFF ", "", x = _) |>
+    gsub("OFFENCE ", "", x = _) |>
+    gsub(toupper("recommended disposal at point of triage"), "disposal", x = _) |>
     gsub("DATE OF SUBMISSION", "DATE", x = _) |>
     tolower()
 }
@@ -38,29 +42,39 @@ op_plot_offence = function(d) {
     ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1))
 }
 
-op_geocode = function(d, region = "West Yorkshire") {
-    d_locations_string = d |>
+op_geocode = function(d, region = "West Yorkshire", method = "google") {
+    address = d |>
         dplyr::pull(location)
-    d_locations_string = paste(d_locations_string, region, sep = ", ")
-        
-    d_locations = lapply(d_locations_string, FUN = stplanr::geo_code)
-    
-    d_locations_lengths = sapply(d_locations, function(x) {
-        length(x)
-    })
-    d_is_geo = d_locations_lengths > 0
-    d_with_geo = d[d_is_geo, ]
-    d_locations = d_locations[d_is_geo]
-    
-    d_point = lapply(d_locations, function(x) {
-        sf::st_point(matrix(x, ncol = 2))
-    })
-    d_sfc = sf::st_sfc(d_point) |>
-        sf::st_sf(crs = "EPSG:4326")
-    d_sf = sf::st_sf(
-        d_with_geo,
-        geometry = d_sfc |> sf::st_geometry()
+    address = paste(address, region, sep = ", ")
[…]

### 2024-04-08 README.qmd
Filter out locations

@@
 ```{r}
-#| eval: false
+#| include: false
 #| echo: false
@@
   mutate(records = "all")
-d = d_all |>
-  opsnap:::filter_nas()
+d_offence = d_all |>
+  opsnap:::filter_offence_nas()
+d_offence_monthly = d_offence |>
+  mutate(month = lubridate::floor_date(date, "month")) |>
+  group_by(month) |>
+  summarise(n = n()) |>
+  mutate(records = "complete_offence")
+d = d_offence |>
+  opsnap:::filter_location_nas()
 d_complete_monthly = d |>
@@
   mutate(records = "complete")
-d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
+d_monthly = bind_rows(d_all_monthly, d_offence_monthly, d_complete_monthly)
 d_monthly |>
   ggplot() +
-  geom_line(aes(month, n, colour = records)) +
-  geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
+  geom_line(aes(month, n, colour = records), alpha = 0.5, size = 2) +
+  # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
   labs(title = "Number of records in West Yorkshire Police Operation Snap data",
        x = "Date",
-       y = "Number of records per month")
+       y = "Number of records per month") +
+  theme_minimal()
 ```
 
-As shown in the graph above, some records `r round(nrow(d) / nrow(d_all) * 100, 1)`% have missing values for locations (these records did not classify as an offence).
+As shown in the graph above, some records (`r round(nrow(d_offence) / nrow(d_all) * 100, 1)`%) have missing values for the 'offence' column (these records did not classify as an offence). 
+Many records lack either an offence or a location (`r round(nrow(d) / nrow(d_all) * 100, 1)`%).
 Filtering these out leaves `r nrow(d)` complete records.
@@
 d_geocoded = opsnap:::op_geocode(d)
+table(d_geocoded$address) |>
+  sort() |>
+  tail(1)
 d_geometries = d_geocoded |>
+  filter(address != "NA, West Yorkshire") |>
   sf::st_as_sf(coords = c("long", "lat"), crs = 4326)

### 2024-04-08 R/opsnap.R

@@ -30,9 +30,16 @@ select_columns = function(d) {
   d = d[, !names_remove]
 }
 
-filter_nas = function(d) {
+filter_offence_nas = function(d) {
   d |>
-    dplyr::filter(!offence == "N/A")
+    dplyr::filter(!offence == "N/A") |>
+    dplyr::filter(!is.na(offence))
+}
+
+filter_location_nas = function(d) {
+  d |>
+    dplyr::filter(!location == "N/A") |>
+    dplyr::filter(!is.na(location))
 }
 
 op_plot_offence = function(d) {

### 2024-04-09 README.qmd
Update results

@@
 
-As shown in the graph above, some records (`r round(nrow(d_offence) / nrow(d_all) * 100, 1)`%) have missing values for the 'offence' column (these records did not classify as an offence). 
-Many records lack either an offence or a location (`r round(nrow(d) / nrow(d_all) * 100, 1)`%).
-Filtering these out leaves `r nrow(d)` complete records.
+As shown in the graph above, `r round(nrow(d_offence) / nrow(d_all) * 100, 1)`% have values for the 'offence' column. 
+Many records lack either an offence or a location, leaving only `r round(nrow(d) / nrow(d_all) * 100, 1)`% or `r nrow(d)` complete records.
 
-There are `r unique(d$location) |> length()` unique locations in the data, with the most common locations shown below:
+There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:
 
@@
 
-The results show there is one outlier with a very high number of crashes. We can remove this and plot the data again:
+<!-- The results show there is one outlier with a very high number of crashes. We can remove this and plot the data again: -->
 
 ```{r}
-raster_count_no_outlier = raster_count |>
-  clamp(-Inf, 100)
-plot(raster_count_no_outlier)
+# raster_count_no_outlier = raster_count |>
+#   clamp(-Inf, 100)
+# plot(raster_count_no_outlier)
 ```

### 2024-04-09 README.qmd
update readme

@@
 usethis::use_r("opsnap")
+
+# Rbuildignore the data folder:
+usethis::use_build_ignore("data")
 # MIT license:
@@
 d_all = read_csv(file_name)
+d_all = d_all |>
+  mutate(
+    mode = tolower(mode),
+    offence = tolower(offence),
+    disposal = tolower(disposal)
+  )
 d_all |>
@@
 ```
+
+# Preliminary analysis
 
@@
 
+The breakdown of records by mode of transport (of the observer) is shown below:
+
+```{r}
+d_all |>
+  count(mode, sort = TRUE) |>
+  mutate(percent_records = n / nrow(d_all)) |>
+  mutate(percent_records = scales::percent(percent_records)) |>
+  arrange(desc(n)) |>
+  knitr::kable()
+```
+
+The offence text strings are quite long, with the most common offences shown below:
+
+```{r}
+d_all |>
+  count(offence, sort = TRUE) |>
+  mutate(percent_records = n / nrow(d_all)) |>
+  mutate(percent_records = scales::percent(percent_records)) |>
+  arrange(desc(n)) |>
+  head(10) |>
+  knitr::kable()
+```
+
+The equivalent table excluding records with missing offence data is shown below:
+
+```{r}
+d_offence |>
+  count(offence, sort = TRUE) |>
+  mutate(percent_records = n / nrow(d_offence)) |>
+  mutate(percent_records = scales::percent(percent_records)) |>
+  arrange(desc(n)) |>
+  head(10) |>
+  knitr::kable()
+```
+
+In terms 'disposal', the most common values are shown below:
+
+```{r}
+d_all |>
+  count(disposal, sort = TRUE) |>
+  mutate(percent_records = n / nrow(d_all)) |>
+  mutate(percent_records = scales::percent(percent_records)) |>
+  arrange(desc(n)) |>
+  knitr::kable()
+```
+
 There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:
@@
   tail(1)
-d_geometries = d_geocoded |>
-  filter(address != "NA, West Yorkshire") |>
+# d_geometries = d_geocoded |>
+#   filter(address != "NA, West Yorkshire") |>
   sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
@@
 d_sf_wy = d_sf[west_yorkshire, ]
+proportion_outside_wy = 1 - nrow(d_sf_wy) / nrow(d_sf)
 # Sanity check
@@
 
-# Analysis
+After geocoding all records we kept only those within the boundary of West Yorkshire, which removed another 3% of records.
+
+# Location of incidents
 
@@
 ```
+
+The map above represents `r nrow(d_sf)` incidents in West Yorkshire with an offence that could be geocoded.
 

### 2024-04-09 README.qmd
Update with offence types

@@
 ```{r}
-d_offence |>
+d_all |>
+  opsnap:::filter_offence_nas() |>
+  filter(offence != "n/a") |>
   count(offence, sort = TRUE) |>
@@
   head(10) |>
+  knitr::kable()
+```
+
+The equivalent for cyclists, with the least common offences categorised as 'other', is shown below:
+
+```{r}
+d_all |>
+  opsnap:::filter_offence_nas() |>
+  filter(offence != "n/a") |>
+  filter(mode == "cyclist") |>
+  count(offence, sort = TRUE) |>
+  mutate(percent_records = n / nrow(d_offence)) |>
+  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(
+    offence = ifelse(n < 20, "other", offence)
+  ) |>
+  group_by(offence) |>
+  summarise(n = sum(n), n_hybrid = n()) |>
+  arrange(n_hybrid, desc(n)) |>
+  select(-n_hybrid) |>
+  mutate(`% of total` = scales::percent(n / sum(n), accuracy = 0.1)) |>
   knitr::kable()

### 2024-04-22 paper.qmd
Add paper

--- 2024-04-09 README.qmd
@@
 ---
-format: gfm
-execute: 
-  message: false
-  warning: false
-  echo: false
+title: "Cyclist video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda"
+format: pdf
+number-sections: true
+bibliography: references.bib
 ---
 
-<!-- badges: start -->
-[![R-CMD-check](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml)
-<!-- badges: end -->
-
-```{r}
-#| echo: false
-#| eval: false
-# Package setup:
-usethis::use_description()
-usethis::use_package("sf")
-usethis::use_package("stplanr")
-usethis::use_package("ggplot2")
-usethis::use_package("dplyr")
-usethis::use_r("opsnap")
-
-# Rbuildignore the data folder:
-usethis::use_build_ignore("data")
-# MIT license:
-usethis::use_mit_license("Leeds Institute for Transport Studies")
-
-# Add CI
-usethis::use_github_action_check_standard()
-# Add pkgdown
-usethis::use_pkgdown()
-# Add pkgdown action
-usethis::use_github_action("pkgdown")
-# Add gh pages:
-usethis::use_github_pages()
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  echo = FALSE,
+  warning = FALSE
+)
 ```
 
-# Installation
+```{r setup}
+library(opsnap)
+```
 
-Install the package from GitHub:
+# Abstract {.unnumbered}
 
-```r
-remotes::install_github("ITSLeeds/opsnap")
-```
+# Introduction
+
+As shown in @sec-methods.
+
+test test [@prati2019].
+
+# Methods and data {#sec-methods}
 
@@
 #| echo: false
+#| include: false
 library(rvest)
@@
 #| echo: true
+#| include: false
 u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
@@
 #| echo: false
+#| include: false
 d_all = read_csv(file_name)
@@
 
-There are `r nrow(d_all)` records in the data, with increasing numbers of records over time (average n. records per month shown below):
+There are `r nrow(d_all)` records in the data, with increasing numbers of records over time (average n. records per month shown in @fig-time):
 
 ```{r}
+#| label: fig-time
+#| fig-cap: "Number of records per month in the opSnap dataset for West Yorkshire"
 d_all_monthly = d_all |>
@@
   # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
-  labs(title = "Number of records in West Yorkshire Police Operation Snap data",
+  labs(title = "Number of records in West Yorkshire\nPolice Operation Snap data",
        x = "Date",
@@
 
-# Geocoding
+# Results
 
-We provide a function to geocode the records:
+Test results are shown in @fig-test.
 
 ```{r}
-#| eval: false
-#| echo: true
-d_sample = d[1:5, ]
-d_sf = opsnap:::op_geocode(d_sample)
-mapview::mapview(d_sf)
+#| label: fig-test
+#| fig-cap: "Illustration of crashes over time"
+plot(1:9)
 ```
 
-```{r}
-#| eval: false
-#| echo: false
-# Geocode all data:
-d_geocoded = opsnap:::op_geocode(d)
-table(d_geocoded$address) |>
-  sort() |>
-  tail(1)
-# d_geometries = d_geocoded |>
-#   filter(address != "NA, West Yorkshire") |>
-  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
-d_sf = sf::st_sf(
-  d,
-  geometry = d_geometries$geometry
-)
-west_yorkshire = pct::get_pct_zones("west-yorkshire")
-sf::sf_use_s2(FALSE)
-d_sf_wy = d_sf[west_yorkshire, ]
-proportion_outside_wy = 1 - nrow(d_sf_wy) / nrow(d_sf)
-# Sanity check
-d_sf_wy |>
-  sample_n(1000) |>
-  # mapview::mapview()
-  plot()
-# Save to file:
-sf::write_sf(d_sf_wy, paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
-```
+# Discussion
 
-After geocoding all records we kept only those within the boundary of West Yorkshire, which removed another 3% of records.
-
-# Location of incidents
-
-Due to inaccuracy in the geocoding, we only know the locations of the records to within around 500m of each crash (although we can link to specific roads).
-We'll present the geographic distribution of crashes using a 500m grid:
-
-```{r}
-d_sf = sf::read_sf(paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
-d_projected = sf::st_transform(d_sf, 27700)
-library(terra)
-raster_template = rast(ext(d_projected), res = 500)
-raster_count = rasterize(d_projected, raster_template, fun = "length")
-plot(raster_count)
-```
-
-The map above represents `r nrow(d_sf)` incidents in West Yorkshire with an offence that could be geocoded.
-
-<!-- The results show there is one outlier with a very high number of crashes. We can remove this and plot the data again: -->
-
-```{r}
-# raster_count_no_outlier = raster_count |>
-#   clamp(-Inf, 100)
-# plot(raster_count_no_outlier)
-```
-
-<!-- You can query the data downloaded with `opsnap` functions, e.g. as follows (results not shown): -->
-
-```{r}
-#| eval: false
-#| echo: false
-table(d$offence) |>
-  sort()
-```
-
-<!-- Let's make a plot of the data: -->
-
-```{r}
-#| eval: false
-#| echo: false
-d |>
-  # Reduce nchar of offence
-  mutate(offence = stringr::str_sub(offence, 1, 60)) |>
-  group_by(offence) |>
-  # Count number of rows in each group
-  mutate(n = n()) |>
-  filter(n > nrow(d) / 50)|>
-  ggplot() +
-  geom_bar(aes(offence)) +
-  # Make x labels vertical
-  theme(axis.text.x = element_text(angle = 60, hjust = 1))
-
-```
+# References

### 2024-04-22 paper.qmd
Improve table

@@
   mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
   arrange(desc(n)) |>
-  knitr::kable()
+  knitr::kable(digits = 1)
 ```

### 2024-04-22 paper.qmd
Wrap text

--- 2024-04-22 paper.qmd
@@
 bibliography: references.bib
+editor:
+  markdown: 
+    wrap: sentence
 ---

### 2024-04-22 paper.qmd
Add abstract

@@
 # Abstract {.unnumbered}
+
+This study uses data from Operation Snap (Op Snap), the UK police’s national system to receive road users’ video evidence of road traffic offences.
+Data from one police force area for three calendar years, from its launch in 2021 (N= 18,363 records) is analysed.
+34.4% were submitted by cyclists (n=6,312).
+69% of records resulted in an offence being registered and of those and 63% were for driving ‘without reasonable consideration to others’ or ‘without due care and attention’.
+Three quarters (75.2%) resulted in the recommended disposal of an educational course (including conditional offers), a quarter in no further action, and fewer than 1% for court appearance.
+A preliminary research agenda using Op Snap data is outlined.
+In the first instance, comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional and local levels, replicating the findings of this study.
+Studies of disposal-related decision-making and the role of factors such as video quality and supporting evidence might identify good practice.
+Research into video content, offence concentration, recidivism, repeat submitters of evidence, and case progression including court cases, hold significant potential to inform policy and practice locally, nationally and internationally.
 

### 2024-04-22 paper.qmd
Add corbett paper

@@
     wrap: sentence
+# knitr:
+#   opts_chunk:
+#     fig.path: images/
+# # # Uncomment for formatted version
+  # elsevier-pdf:
+  #   linestretch: 2
+  #   keep-tex: true
+  #   journal:
+  #     name: European Transport Research Review
+  #     formatting: preprint
+  #     model: 3p
+  #     cite-style: super
+# # # #   # pdf: default
+
+# # Uncomment the following line to produce the cover page
+author:
+  - name: Graham Farrell
+    affiliations:
+      - name: University of Leeds
+        address: TBC
+        orcid: TBC
+  - name: Robin Lovelace
+    email: r.lovelace@leeds.ac.uk
+    corresponding: true
+    affiliations:
+      - name: University of Leeds
+        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
+        orcid: 0000-0001-5679-6536
+  - name: Steve O'Hern
+    affiliations:
+      - name: University of Leeds
+        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
+        orcid: tbc
+keywords: 
+  - Cycling
+  - Hotspots 
+  - Road Safety
+  - Active Travel
+  - Transport Planning
+  - Near Misses
+date: last-modified
 ---
@@
 
-As shown in @sec-methods.
-
-test test [@prati2019].
+Dangerous and criminal driving are significant problems that take many forms [@simon1996] [@corbett2010; @corbett2003] (Corbett 2003, 2010).
+Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities (ref).
+Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (ref).
 

### 2024-04-22 paper.qmd
Next paragraph

@@
 Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (ref).
+
+The context for the study is that, between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year according to official records recorded by police forces across Great Britain [@departmentfortransport].
+Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
+The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” [@departmentfortransport].
 

### 2024-04-22 paper.qmd
Updates to paper

@@
 
+Furthermore, it is noted that cyclist crashes are underreported in police recorded datasets (refs) and the extent of injuries sustained by cyclists may be significantly higher when also considering hospital reported cases (references).
+Furthermore, under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions (reference).
+These incidents are often referred to as “near-misses” (reference).
+
+In the UK it has been estimated that commuter cyclists experience approximately one near miss for every 6 miles of riding (reference).
+Previous research has also demonstrated that the concern regarding near misses is a barrier reported for people choosing not to cycle (...), with near misses associated with various driver behaviours including inattentive driving, aggressive driving, drivers driving too fast, drivers passing too close, being car doored, and being cut off by turning drivers (Refs).
+In their recent literature review of near miss cycling crashes, Ibrahim et al (ref) highlight the need for data on near misses to be recorded, as they provide rich information with which to study cyclist crash risk and identify factors associated with risk.
+
+Close passes are the most common type of near miss reported by cyclists, and are associated with collisions resulting in injury [@aldred2016] .
+A ‘close pass’ refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph) (XX REF). 
+Close passes take different forms including the ‘punishment pass’ by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
+There is no specific law in the UK Road Traffic Act 1988 for driving to close to cyclist, however there are two laws for careless driving which are commonly applied: RT88575- Drive without due care and attention; and RT99576 – Driver without reasonable consideration to others (ref).
+
+-   A paragraph or two about justice/ road crime??
+
+This study highlights the potential of using open access data from Operation Snap, the UK police’s national system to receive road users’ video evidence of road traffic offences.
+Operation Snap, often referred to informally as ‘Op Snap’, was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 2018 (Road Safety GB 2018).
+It is now in operation nationally across England and Wales, each police force offering its own submission portal (XX NEED TO CHECK IF ALL HAVE IT)  for road users to submit video evidence.
+The nature of video submissions and the related expectations are summarised on the website of one Police and Crime Commissioner as follows:
+
+-   The secure form is for traffic offences, it is NOT for submitting footage of road traffic collisions, any other offences or for parking issues.
+
+-   The car registration number of the offending vehicle must be clearly visible.
+
+-   The public should be prepared to sign a witness statement and possibly give evidence in court.
+
+-   Statements for OpSnap can only be accepted from persons aged 18 or over.
+    If you are under 18 the incident should be reported by email to \[email\] [@davidson2022].
+
+This is, to our knowledge, the first peer-reviewed study to use this dataset.
+As such, the study is offered as proof-of-concept of the potential for analysis of this data, and of its further potential, including understanding the factors associated with cyclist near miss incidents.
+Following analysis and discussion of three years of data for one police force area, the study outlines a research agenda designed to inform policy and practice.
+
 # Methods and data {#sec-methods}
+
+Open access data from West Yorkshire Safer Roads – Op Snap (West Yorkshire Police 2024) was used for this study, the media submissions portal having opened in July 2020 (West Yorkshire Police 2020).
+The West Yorkshire Police (WYP) data used here spanned calendar years 2021 and 2023.
+Data for 2021 was included, however it was far less numerous, with less than half the cases of either 2022 or 2023.
+This could relate to reduced road use during the COVID-19 pandemic .
+
+The dataset consists of a tabulation of reported cases that have been submitted to the West Yorkshire Safer Roads Op Snap web portal.
+
+The terms ‘record’ and ‘case’ are used interchangeably here to refer to a record in the Op Snap database, each of which represents the separate submission of video evidence by a road user.
+
+The portal allows members of the public to submit video footage of suspected traffic offences committed by motor vehicle drivers.
+Video footage is commonly recorded from on-board cameras. 
+For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, and are often referred to as ‘dash cams’.
+One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam (iCompario, 2024).
+For cyclists, footage is commonly recorded using helmet or handlebar mounted cameras, the proportion of cyclists using these cameras is unknown, however anecdotal evidence suggests that there is increased usage, with many choosing to record rides as proof if an incident does occur.
+
+Complainants upload the footage and complete a short form that includes their personal details, the details of the vehicle involved including registration, make model and colour, the location and time of the incident, and details of the camera used to record the footage (ref).
+It is noted that only offences committed by registered motor vehicles can be reported as complainants are required to submit the registration number of the offending vehicle, which must also be readable in the uploaded footage.
+
+A deidentified summary of the submitted cases is available for download, which also includes details of the reported offence and the disposal decision by police.
+This deidentified dataset forms the bases for the analysis presented in this study.
+
+Each record contained an ‘Offence Location’ field.
+This was typically a street name and town or city name, or an intersection.
+Example offence locations included: ‘A58 Godley Road, Halifax’, ‘Keighley Road, Silsden’ , ‘Woodhouse Lane A660, Leeds’.
+Approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
+
+However, it is a noted limitation that these are not precise geolocations but allow some broad analysis of the clustering of offences. 
+XX THERE WAS MISSING LOCATION DATA FOR SOME MONTHS – SO IT WASN’T ‘LOCATION NOT AVAILABLE’ MISSINGS BUT LOOKED MORE LIKE IT WAS OMITTED ACCIDENTALLY OR CUT?
 
@@
 ```
-
-# Opsnap data
-
-The `opsnap` package provides a function to download and read in data from the West Yorkshire Police Operation Snap database.
-The data is available at the following URL: https://www.westyorkshire.police.uk/SaferRoadsSubmissions
-
-Data for the following years are provided:
 
@@
 
-# Preliminary analysis
+# Results
 
@@
 
-# Results
-
-Test results are shown in @fig-test.
-
-```{r}
-#| label: fig-test
-#| fig-cap: "Illustration of crashes over time"
-plot(1:9)
-```
-
 # Discussion

### 2024-04-22 README.qmd

@@
 # Package setup:
-usethis::use_description()
-usethis::use_package("sf")
-usethis::use_package("stplanr")
-usethis::use_package("ggplot2")
-usethis::use_package("dplyr")
-usethis::use_r("opsnap")
-
-# Rbuildignore the data folder:
-usethis::use_build_ignore("data")
-# MIT license:
-usethis::use_mit_license("Leeds Institute for Transport Studies")
-
-# Add CI
-usethis::use_github_action_check_standard()
-# Add pkgdown
-usethis::use_pkgdown()
-# Add pkgdown action
-usethis::use_github_action("pkgdown")
-# Add gh pages:
-usethis::use_github_pages()
+# usethis::use_description()
+# usethis::use_package("sf")
+# usethis::use_package("stplanr")
+# usethis::use_package("ggplot2")
+# usethis::use_package("dplyr")
+# usethis::use_r("opsnap")
+# 
+# # Rbuildignore the data folder:
+# usethis::use_build_ignore("data")
+# # MIT license:
+# usethis::use_mit_license("Leeds Institute for Transport Studies")
+# 
+# # Add CI
+# usethis::use_github_action_check_standard()
+# # Add pkgdown
+# usethis::use_pkgdown()
+# # Add pkgdown action
+# usethis::use_github_action("pkgdown")
+# # Add gh pages:
+# usethis::use_github_pages()
 ```

### 2024-04-22 paper.qmd

@@
 
-Dangerous and criminal driving are significant problems that take many forms [@simon1996] [@corbett2010; @corbett2003] (Corbett 2003, 2010).
+Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
 Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities (ref).
@@
 Approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
-
-However, it is a noted limitation that these are not precise geolocations but allow some broad analysis of the clustering of offences. 
-XX THERE WAS MISSING LOCATION DATA FOR SOME MONTHS – SO IT WASN’T ‘LOCATION NOT AVAILABLE’ MISSINGS BUT LOOKED MORE LIKE IT WAS OMITTED ACCIDENTALLY OR CUT?
 
@@
   head(3) |>
+  # Convert column names to title case:
+  rename_all(snakecase::to_title_case) |>
+  # snakecase::to_title_case
   knitr::kable()
@@
 
-There are `r nrow(d_all)` records in the data, with increasing numbers of records over time (average n. records per month shown in @fig-time):
+There are `r nrow(d_all)` records in the data, with increasing numbers of records over time, as illustrated in @fig-time:
 
@@
   mutate(records = "complete")
-d_monthly = bind_rows(d_all_monthly, d_offence_monthly, d_complete_monthly)
+d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
 d_monthly |>
+  mutate(records = snakecase::to_title_case(records)) |> 
+  rename_all(snakecase::to_title_case) |> 
   ggplot() +
-  geom_line(aes(month, n, colour = records), alpha = 0.5, size = 2) +
+  geom_line(aes(Month, N, colour = Records), alpha = 0.5, size = 2) +
   # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
-  labs(title = "Number of records in West Yorkshire\nPolice Operation Snap data",
+  labs(title = "Number of records in West Yorkshire Police\nOperation Snap data",
        x = "Date",
@@
 
-The breakdown of records by mode of transport (of the observer) is shown below:
+The breakdown of records by mode of transport (of the observer) is shown @tbl-mode.
 
 ```{r}
+#| label: tbl-mode
+#| tbl-cap: Number of OpSnap records by observer mode.
 d_all |>
@@
   mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
   arrange(desc(n)) |>
+  rename_all(snakecase::to_title_case) |>
   knitr::kable()
@@
 
-The offence text strings are quite long, with the most common offences shown below:
+The offence text strings are quite long, with the most common offences shown in @tbl-offences:
 
 ```{r}
+#| label: tbl-offences
+#| tbl-cap: Number and percentages of OpSnap records by offence type.
 d_all |>
@@
   mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
   arrange(desc(n)) |>
   head(10) |>
+  rename_all(snakecase::to_title_case) |>
   knitr::kable()
@@
 
-The equivalent table excluding records with missing offence data is shown below:
+<!-- The equivalent table excluding records with missing offence data is shown below: -->
 
 ```{r}
+#| include: false
 d_all |>
@@
   mutate(percent_records = n / nrow(d_offence)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
   arrange(desc(n)) |>
   head(10) |>
+  rename_all(snakecase::to_title_case) |>
   knitr::kable()
@@
 
-The equivalent for cyclists, with the least common offences categorised as 'other', is shown below:
+The equivalent for cyclists, with the least common offences categorised as 'other', is shown in @tbl-offences-cyclist-observer.
 
 ```{r}
+#| label: tbl-offences-cyclist-observer
+#| tbl-cap: "Number and percentages of OpSnap records, submitted by cyclists, by offence type."
 d_all |>
@@
   mutate(percent_records = n / nrow(d_offence)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
   mutate(
@@
   mutate(`% of total` = scales::percent(n / sum(n), accuracy = 0.1)) |>
+  rename_all(snakecase::to_title_case) |> 
   knitr::kable()
@@
 
-In terms 'disposal', the most common values are shown below:
+In terms 'disposal', the most common values are shown in @tbl-disposal.
 
 ```{r}
+#| label: tbl-disposal
+#| tbl-cap: "Most common disposal values in the OpSnap dataset."
 d_all |>
@@
   mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
   arrange(desc(n)) |>
+  rename_all(snakecase::to_title_case) |> 
   knitr::kable()
@@
 
-There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:
+There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown in @tbl-locations:
 
@@
 #| echo: false
+#| label: tbl-locations
+#| tbl-cap: "Most common locations recorded in the OpSnap dataset"
 d |>
@@
   mutate(percent_records = n / nrow(d)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
   arrange(desc(n)) |>
   head(10) |>
+  rename_all(snakecase::to_title_case) |> 
   knitr::kable()

### 2024-04-22 R/opsnap.R
Add function reference, close #3

@@ -1,4 +1,16 @@
-download_and_read = function(u, remove_nas = TRUE, dir = "raw_data/west-yorkshire", filter_nas = FALSE) {
+#' Download and read data from opsnap project
+#' 
+#' @param u URL of data
+#' @param remove_nas Remove NA values?
+#' @param dir Download directory
+#' @return Data frame
+#' @export
+#' @examples
+#' if (FALSE) {
+#' u = "https://www.westyorkshire.police.uk/sites/default/files/2023-10/operation_snap_july-sept_2023.xlsx"
+#' download_and_read(, dir = ".")
+#' }
+download_and_read = function(u, remove_nas = TRUE, dir = "raw_data/west-yorkshire") {
     tmp = file.path(dir, basename(u))
     if (!file.exists(tmp)) {
         utils::download.file(u, tmp, mode = "wb")
@@ -8,7 +20,7 @@ download_and_read = function(u, remove_nas = TRUE, dir = "raw_data/west-yorkshir
     })
     names(d) = clean_names(names(d))
     d = select_columns(d)
-    if (filter_nas) {
+    if (remove_nas) {
         d = filter_nas(d)
     }
     return(d)

### 2024-05-03 README.qmd
Update readme

@@
 # In preparation for reading-in all data:
-date_str = format(Sys.Date(), "%Y-%m")
+# date_str = format(Sys.Date(), "%Y-%m")
+date_str = format(as.Date("2024-04-01"), "%Y-%m")
 file_name = paste0("data/west-yorkshire/operation_snap_", date_str, ".csv")
@@
 d = opsnap:::download_and_read(u)
-names(d)
+names(d_with_location)
 # Old names:
@@
 d_all = d_all |>
-  mutate(
-    mode = tolower(mode),
-    offence = tolower(offence),
-    disposal = tolower(disposal)
-  )
+  mutate_if(is.character, tolower)
 d_all |>
@@
   summarise(n = n()) |>
-  mutate(records = "all")
-d_offence = d_all |>
+  mutate(records = "All")
+d_with_offence = d_all |>
   opsnap:::filter_offence_nas()
-d_offence_monthly = d_offence |>
+# nrow(d_all) - nrow(d_with_offence) # only 1 record -> ignore
+d_with_location = d_all |>
+  opsnap:::filter_location_nas()
+d_with_location_monthly = d_with_location |>
   mutate(month = lubridate::floor_date(date, "month")) |>
@@
   summarise(n = n()) |>
-  mutate(records = "complete_offence")
-d = d_offence |>
-  opsnap:::filter_location_nas()
-d_complete_monthly = d |>
-  mutate(month = lubridate::floor_date(date, "month")) |>
-  group_by(month) |>
-  summarise(n = n()) |>
-  mutate(records = "complete")
-d_monthly = bind_rows(d_all_monthly, d_offence_monthly, d_complete_monthly)
+  mutate(records = "With location")
+d_monthly = bind_rows(d_all_monthly, d_with_location_monthly)
 d_monthly |>
@@
        y = "Number of records per month") +
-  theme_minimal()
+  theme_minimal() +
+  ylim(c(0, NA))
 ```
 
-As shown in the graph above, `r round(nrow(d_offence) / nrow(d_all) * 100, 1)`% have values for the 'offence' column. 
-Many records lack either an offence or a location, leaving only `r round(nrow(d) / nrow(d_all) * 100, 1)`% or `r nrow(d)` complete records.
+As shown in the graph above not all (`r round(nrow(d_with_location) / nrow(d_all) * 100, 1)`%) records have values for the 'location' column.
 
-The breakdown of records by mode of transport (of the observer) is shown below:
+The breakdown of all records by mode of transport (of the observer) is shown below:
 
@@
   mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
   arrange(desc(n)) |>
@@
 
-The equivalent table excluding records with missing offence data is shown below:
+The equivalent table excluding records with missing location data is shown below:
 
@@
   count(offence, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_offence)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = n / nrow(d_with_location)) |>
+  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
   arrange(desc(n)) |>
@@
 
-The equivalent for cyclists, with the least common offences categorised as 'other', is shown below:
+The equivalent for cyclists, with location present and the least common offences categorised as 'other', is shown below:
 
@@
   count(offence, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_offence)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = n / nrow(d_with_location)) |>
+  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
   mutate(
@@
 
-There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:
+There are `r unique(d_with_location$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:
 
@@
 #| echo: false
-d |>
+d_with_location |>
+  mutate(location = tolower(location)) |> 
   count(location, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d)) |>
-  mutate(percent_records = scales::percent(percent_records)) |>
+  mutate(percent_records = n / nrow(d_with_location)) |>
+  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
   arrange(desc(n)) |>
@@
 #| echo: true
-d_sample = d[1:5, ]
+d_sample = d_with_location[1:5, ]
 d_sf = opsnap:::op_geocode(d_sample)
@@
 # Geocode all data:
-d_geocoded = opsnap:::op_geocode(d)
+d_geocoded = opsnap:::op_geocode(d_with_location)
 table(d_geocoded$address) |>
@@
 d_sf = sf::st_sf(
-  d,
+  d_with_location,
   geometry = d_geometries$geometry
@@
 d_sf = sf::read_sf(paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
+```
+
+```{r}
+#| eval: false
+d_sf_aggregated = d_sf |> 
+  group_by(location) |> 
+  count(mode)
+library(tmap)
+tmap_mode("view")
+d_sf |> 
+  sample_n(1000) |> 
+  tm_shape() +
+  tm_dots()
+```
+
+```{r}
 d_projected = sf::st_transform(d_sf, 27700)
@@
 raster_count = rasterize(d_projected, raster_template, fun = "length")
+# mapview::mapview(raster_count)
 plot(raster_count)
@@
 #| echo: false
-table(d$offence) |>
+table(d_with_location$offence) |>
   sort()
@@
   mutate(n = n()) |>
-  filter(n > nrow(d) / 50)|>
+  filter(n > nrow(d_with_location) / 50)|>
   ggplot() +

### 2024-05-13 paper.qmd
Update template

@@
 title: "Cyclist video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda"
-format: pdf
-number-sections: true
 bibliography: references.bib
@@
     wrap: sentence
-# knitr:
-#   opts_chunk:
-#     fig.path: images/
-# # # Uncomment for formatted version
-  # elsevier-pdf:
-  #   linestretch: 2
-  #   keep-tex: true
-  #   journal:
-  #     name: European Transport Research Review
-  #     formatting: preprint
-  #     model: 3p
-  #     cite-style: super
-# # # #   # pdf: default
-
-# # Uncomment the following line to produce the cover page
 author:
@@
       - name: University of Leeds
-        address: TBC
-        orcid: TBC
+        address: School of Law, University of Leeds, Leeds, UK
+        orcid: 0000-0002-3987-8457
   - name: Robin Lovelace
@@
 date: last-modified
+format:
+  # pdf: default
+  arxiv-pdf:
+    keep-tex: true
+linenumbers: true # Add (continuous) line numbers?
+doublespacing: true # Double space the PDF output?
+runninghead: "Preprint" # The text on the top of each page of the output
+authorcols: false # Should authors be listed in a single column (default) or in multiple columns (`authorcols: true`)
 ---

### 2024-05-13 paper.qmd
Addresses

@@
   - name: Graham Farrell
-    affiliations:
-      - name: University of Leeds
-        address: School of Law, University of Leeds, Leeds, UK
-        orcid: 0000-0002-3987-8457
-  - name: Robin Lovelace
-    email: r.lovelace@leeds.ac.uk
     corresponding: true
@@
       - name: University of Leeds
-        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
+        department: School of Law
+        address: The Liberty Building
+        city: Leeds
+        country: United Kingdom
+        postal-code: LS2 9JT
+    orcid: 0000-0002-3987-8457
+  - name: Robin Lovelace
+    affiliations:
+      - name: University of Leeds
+        department: Institute for Transport Studies
+        address: University Road
+        city: Leeds
+        country: United Kingdom
+        postal-code: LS2 9JT
+      - name: Active Travel England
+        # department: Data and Digital
+        # address: West Offices Station Rise
+        # city: York
+        # country: United Kingdom
+        # postal-code: YO1 6GA
     orcid: 0000-0001-5679-6536
@@
       - name: University of Leeds
-        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
-        orcid: tbc
+        department: Institute for Transport Studies
+        address: University Road
+        city: Leeds
+        country: United Kingdom
+        postal-code: LS2 9JT
+      - name: Monash University Accident Research Centre
+        address: Clayton, VIC 3800, Australia
+    orcid: 0000-0002-7961-3875
 keywords: 
@@
 linenumbers: true # Add (continuous) line numbers?
-doublespacing: true # Double space the PDF output?
+doublespacing: false # Double space the PDF output?
 runninghead: "Preprint" # The text on the top of each page of the output

### 2024-05-13 paper.qmd
Intro-start

--- 2024-05-03 paper.qmd
@@
     wrap: sentence
-# knitr:
-#   opts_chunk:
-#     fig.path: images/
-# # # Uncomment for formatted version
-  # elsevier-pdf:
-  #   linestretch: 2
-  #   keep-tex: true
-  #   journal:
-  #     name: European Transport Research Review
-  #     formatting: preprint
-  #     model: 3p
-  #     cite-style: super
-# # # #   # pdf: default
-
-# # Uncomment the following line to produce the cover page
 author:
   - name: Graham Farrell
-    affiliations:
-      - name: University of Leeds
-        address: TBC
-        orcid: TBC
-  - name: Robin Lovelace
-    email: r.lovelace@leeds.ac.uk
     corresponding: true
@@
   - Hotspots 
-  - Road Safety
-  - Active Travel
-  - Transport Planning
-  - Near Misses
+  - Road safety
+  - Video evidence
+  - Dashcam
+  - Near misses
 date: last-modified
+format:
+  # pdf: default
+  arxiv-pdf:
+    keep-tex: true
+linenumbers: true # Add (continuous) line numbers?
+doublespacing: false # Double space the PDF output?
+runninghead: "Preprint" # The text on the top of each page of the output
+authorcols: true # Should authors be listed in a single column (default) or in multiple columns (`authorcols: true`)
 ---
@@
 
-This study uses data from Operation Snap (Op Snap), the UK police’s national system to receive road users’ video evidence of road traffic offences.
-Data from one police force area for three calendar years, from its launch in 2021 (N= 18,363 records) is analysed.
-34.4% were submitted by cyclists (n=6,312).
-69% of records resulted in an offence being registered and of those and 63% were for driving ‘without reasonable consideration to others’ or ‘without due care and attention’.
-Three quarters (75.2%) resulted in the recommended disposal of an educational course (including conditional offers), a quarter in no further action, and fewer than 1% for court appearance.
-A preliminary research agenda using Op Snap data is outlined.
-In the first instance, comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional and local levels, replicating the findings of this study.
-Studies of disposal-related decision-making and the role of factors such as video quality and supporting evidence might identify good practice.
-Research into video content, offence concentration, recidivism, repeat submitters of evidence, and case progression including court cases, hold significant potential to inform policy and practice locally, nationally and internationally.
+This study uses data from Operation Snap (OpSnap), the UK police’s national system to receive road users’ video evidence of road traffic offences.
+Data from one police force area for three calendar years, from its launch in 2021 (N= 18,363 records), is analysed.
+Of submitted cases, 34.4% were from cyclists (n=6,312) of which 69% resulted in an offence being registered, mostly for driving ‘without reasonable consideration to others’ or ‘without due care and attention’.
+Three quarters (75.2%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), a quarter in no further action, and fewer than 1% for court appearance.
+A research agenda using OpSnap data is outlined: comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional, and local levels; research into disposal-related decision-making, video quality, and the role of supporting evidence should identify good practice; further study of video content, offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases, holds significant potential to inform policy and practice locally, nationally and internationally.
 
@@
 Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
-Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities (ref).
+Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities @dft-xxxx.
 Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (ref).

### 2024-05-13 paper.qmd
Intro

@@
 Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
-Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities @dft-xxxx.
-Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (ref).
+Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@dft-xxxx].
+Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (Department for Transport, 2023a).
+Between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year in Britain according to police records (Department for Transport, 2023b).
+Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
+The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” (Department for Transport, 2023b).
 
-The context for the study is that, between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year according to official records recorded by police forces across Great Britain [@departmentfortransport].
+```{=html}
+<!-- The context for the study is that, between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year according to official records recorded by police forces across Great Britain [@departmentfortransport].
 Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
-The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” [@departmentfortransport].
+The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” [@departmentfortransport].  -->
+```
 
-Furthermore, it is noted that cyclist crashes are underreported in police recorded datasets (refs) and the extent of injuries sustained by cyclists may be significantly higher when also considering hospital reported cases (references).
-Furthermore, under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions (reference).
-These incidents are often referred to as “near-misses” (reference).
+Cyclist crashes are underreported in police recorded datasets (Elvik & Mysen, 1999) and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted (Janstrup et al., 2016).
+Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions (Ibrahim et al., 2021).
+These incidents are often referred to as “near-misses” (Ibrahim et al., 2021).
 
-In the UK it has been estimated that commuter cyclists experience approximately one near miss for every 6 miles of riding (reference).
-Previous research has also demonstrated that the concern regarding near misses is a barrier reported for people choosing not to cycle (...), with near misses associated with various driver behaviours including inattentive driving, aggressive driving, drivers driving too fast, drivers passing too close, being car doored, and being cut off by turning drivers (Refs).
-In their recent literature review of near miss cycling crashes, Ibrahim et al (ref) highlight the need for data on near misses to be recorded, as they provide rich information with which to study cyclist crash risk and identify factors associated with risk.
+Commuter cyclists in the UK experience a near miss for every six miles of riding (Aldred & Crosweller, 2015), and concern over near misses are a key reason people choose not to cycle (Sanders, 2015), being associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers (Sanders, 2015).
+In their review of near miss cycling crashes, Ibrahim et al., (2021) highlighted the need for better data on near misses to inform safety research.
 
-Close passes are the most common type of near miss reported by cyclists, and are associated with collisions resulting in injury [@aldred2016] .
-A ‘close pass’ refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph) (XX REF). 
+Close passes are the most common type of near miss reported by cyclists, and are associated with collisions resulting in injury [@aldred2016].
+A ‘close pass’ refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
 Close passes take different forms including the ‘punishment pass’ by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
-There is no specific law in the UK Road Traffic Act 1988 for driving to close to cyclist, however there are two laws for careless driving which are commonly applied: RT88575- Drive without due care and attention; and RT99576 – Driver without reasonable consideration to others (ref).
-
--   A paragraph or two about justice/ road crime??
+There is no specific law in the UK Road Traffic Act 1988 for driving too close to cyclist, but two are commonly applied for careless driving: RT88575- Drive without due care and attention; and RT99576 – Driver without reasonable consideration to others - and these are prominent in the analysis that follows.
 
 This study highlights the potential of using open access data from Operation Snap, the UK police’s national system to receive road users’ video evidence of road traffic offences.
-Operation Snap, often referred to informally as ‘Op Snap’, was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 2018 (Road Safety GB 2018).
-It is now in operation nationally across England and Wales, each police force offering its own submission portal (XX NEED TO CHECK IF ALL HAVE IT)  for road users to submit video evidence.
+Operation Snap, often referred to informally as ‘OpSnap’, was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 20181.
+It is now in operation nationally across England and Wales, each police force offering its own submission portal for road users to submit video evidence.
 The nature of video submissions and the related expectations are summarised on the website of one Police and Crime Commissioner as follows:
@@
 - The secure form is for traffic offences, it is NOT for submitting footage of road traffic collisions, any other offences or for parking issues. 
+- The car registration number of the offending vehicle must be clearly visible. 
+- The public should be prepared to sign a witness statement and possibly give evidence in court. 
+- Statements for OpSnap can only be accepted from persons aged 18 or over. If you are under 18 the incident should be reported by email to [email]2.
 
--   The car registration number of the offending vehicle must be clearly visible.
-
--   The public should be prepared to sign a witness statement and possibly give evidence in court.
-
--   Statements for OpSnap can only be accepted from persons aged 18 or over.
-    If you are under 18 the incident should be reported by email to \[email\] [@davidson2022].
-
-This is, to our knowledge, the first peer-reviewed study to use this dataset.
+This is, to our knowledge, the first research study to use this dataset.
 As such, the study is offered as proof-of-concept of the potential for analysis of this data, and of its further potential, including understanding the factors associated with cyclist near miss incidents.

### 2024-05-13 paper.qmd
Update text, close #4

@@
 
-Open access data from West Yorkshire Safer Roads – Op Snap (West Yorkshire Police 2024) was used for this study, the media submissions portal having opened in July 2020 (West Yorkshire Police 2020).
-The West Yorkshire Police (WYP) data used here spanned calendar years 2021 and 2023.
-Data for 2021 was included, however it was far less numerous, with less than half the cases of either 2022 or 2023.
-This could relate to reduced road use during the COVID-19 pandemic .
+Open access data from West Yorkshire Safer Roads – OpSnap was used for this study, the media submissions portal having opened in July 2020.
+The West Yorkshire Police (WYP) data used here spanned calendar years 2021, 2022 and 2023.
+For the year 2021 there was less than half the cases of either 2022 or 2023, which could reflect reduced road use during the COVID-19 pandemic and fewer people being aware of OpSnap when it commenced.
 
-The dataset consists of a tabulation of reported cases that have been submitted to the West Yorkshire Safer Roads Op Snap web portal.
-
-The terms ‘record’ and ‘case’ are used interchangeably here to refer to a record in the Op Snap database, each of which represents the separate submission of video evidence by a road user.
+The dataset is a tabulation of cases submitted to the West Yorkshire Safer Roads OpSnap web portal.
+The terms ‘record’ and ‘case’ are used interchangeably here to refer to a record in the OpSnap database, each of which represents the separate submission of video evidence by a road user.
 
@@
 Video footage is commonly recorded from on-board cameras.
-For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, and are often referred to as ‘dash cams’.
+For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as ‘dash cams’.
 One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam (iCompario, 2024).
-For cyclists, footage is commonly recorded using helmet or handlebar mounted cameras, the proportion of cyclists using these cameras is unknown, however anecdotal evidence suggests that there is increased usage, with many choosing to record rides as proof if an incident does occur.
+For cyclists, footage is commonly recorded using helmet or handlebar-mounted cameras.
+The proportion of cyclists using these cameras is unknown but anecdotal evidence suggests increased usage, with many choosing to record rides in case an incident occurs.
 
-Complainants upload the footage and complete a short form that includes their personal details, the details of the vehicle involved including registration, make model and colour, the location and time of the incident, and details of the camera used to record the footage (ref).
-It is noted that only offences committed by registered motor vehicles can be reported as complainants are required to submit the registration number of the offending vehicle, which must also be readable in the uploaded footage.
-
-A deidentified summary of the submitted cases is available for download, which also includes details of the reported offence and the disposal decision by police.
-This deidentified dataset forms the bases for the analysis presented in this study.
-
-Each record contained an ‘Offence Location’ field.
-This was typically a street name and town or city name, or an intersection.
-Example offence locations included: ‘A58 Godley Road, Halifax’, ‘Keighley Road, Silsden’ , ‘Woodhouse Lane A660, Leeds’.
-Approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
+Complainants upload footage and complete a short form that includes their personal details, the details of the vehicle involved including registration, make, model and colour, the location and time of the incident, and details of the camera used to record the footage.
+Only vehicle offences can be reported as the registration number of any offending vehicle is required and must be legible in footage.
+The open access data is a deidentified summary of submitted cases with informtion on: mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
+The offence location is typically a street name and town or city name, or an intersection, and examples included: ‘A58 Godley Road, Halifax’, ‘Keighley Road, Silsden’ , ‘Woodhouse Lane A660, Leeds’.
+For this study, approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
+Further aspects of the data, their uses and limitations are discussed in what follows.
 
@@
 
-There are `r nrow(d_all)` records in the data, with increasing numbers of records over time, as illustrated in @fig-time:
-
 ```{r}
 #| label: fig-time
-#| fig-cap: "Number of records per month in the opSnap dataset for West Yorkshire"
+#| fig-cap: "Monthly count of Operation Snap records, West Yorkshire"
 d_all_monthly = d_all |>
@@
 
-As shown in the graph above, `r round(nrow(d_offence) / nrow(d_all) * 100, 1)`% have values for the 'offence' column.
-Many records lack either an offence or a location, leaving only `r round(nrow(d) / nrow(d_all) * 100, 1)`% or `r nrow(d)` complete records.
+There were `r nrow(d_all)` ecords in the dataset for the three-year study period, shown by month in @fig-time.
+The number of records increased over time.
+Some records lacked either an offence or a location, leaving `r round(nrow(d) / nrow(d_all) * 100, 1)`% or `r nrow(d)` complete records.
+There was a distinct seasonal pattern to reporting, with significant increases in summer months.
 
-The breakdown of records by mode of transport (of the observer) is shown @tbl-mode.
+Mode of transport of the complainant is shown in @tbl-mode.
+Half of cases were reported by vehicle drivers, a third by cyclists, seven percent by pedestrians, with over two percent by horse riders and less than one percent by motorcyclists.
 
@@
 
-The offence text strings are quite long, with the most common offences shown in @tbl-offences:
+Classification of the committed offence is shown in @tbl-offences.
+s previously noted, a high proportion of cases did not include information on the reported office (31.1%), with the most common offences being Driving without reasonable consideration to others (rt88576), and Driving without due care and attention (rt88575).
+Within the road traffic act these offences are related to careless driving and drivers are subject to similar penalties.
 
@@
 
-The equivalent for cyclists, with the least common offences categorised as 'other', is shown in @tbl-offences-cyclist-observer.
+For cases submitted by people riding cycles (shown in @tbl-offences-cyclist-observer), the most common offences were also both associated with careless driving, particularly driving without reasonable consideration to others (78.7%), there were also a small proportion of cases associated with drivers using mobile phones (3.7%), failing to comply with traffic signals (3.5%) and contravening regulator signage (0.6%).
 
@@
 
-In terms 'disposal', the most common values are shown in @tbl-disposal.
+Disposal categories assigned by police are shown in @tbl-disposal.
+ Roughly a third of cases resulted in no further action, and for most of the remainder drivers were required to undertake an education course.
+Conditional offers, that is, drivers being offered a reduced penalty for admitting guilt, were the third most common outcome (XX EXPLAIN FURTHER?).
+Close to two percent of cases went to court and a further one percent underwent further investigation.
 
@@
 
-There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown in @tbl-locations:
+There are `r unique(d$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
+
+INSERT Table 5 AND/OR Map (TO BE INCLUDED)
 
@@
 
+Drawing on the work and findings described above, this discussion outlines a research agenda.
+The breadth and depth of potential policy-relevant work that could be undertaken is, we suggest, enormous.
+While this preliminary list will not prove exhaustive, if it stimulates or informs further research then it will have achieved its objective.
+
+Cyclists undertake two percent of road miles in the UK (Department for Transport, 2023b).
+Other things equal, this suggests that with respect to distance travelled, cyclists were around 20 times more likely to submit evidence of a road traffic offence than vehicle drivers and other road users in West Yorkshire.
+
+Over 90 percent of offences reported by cyclists were for driving without consideration for other road users and without due care and attention.
+This is consistent with the phenomenon of ‘close passes’ in the cycling safety literature (Aldred, 2016; Cubbin et al., 2024).
+It suggests the OpSnap data holds the potential for further analysis to inform knowledge about the nature of close passes generally.
+For example, it should be possible to identify hotspot locations.
+As with much of the further research outlined here, this the potential to inform preventive policy and practice relating to police interventions, driver behaviour, the design of roads and roadside environments.
+
+There was an increase in submissions over time between 2021 and 2023.
+This may reflect changing levels of road use during and after the Covid-19 pandemic, increased awareness of OpSnap, or increased video camera ownership levels.
+Comparative analysis including other police jurisdictions will be able to determine whether this experience is common or isolated.
+We hypothesise that it will prove to be a common experience and that that the volume of submissions will continue to increase in the near and middle term (though ideally, a decline would occur as roads become safer).
+
+If submission volumes continue to increase, means of informing triage and prioritisation of cases will be increasingly valuable.
+Triage systems are probably already in use in different areas, and research to identify and inform best practice should be undertaken.
+Likewise, research to determine best practice in determining disposal recommendations will prove fruitful.
+It may provide feasible, perhaps using machine learning, to develop automated means of triage, determination of offence type and disposal.
+Research to gauge the differential effectiveness of disposal measures is needed, starting with measures of recividism.
+
+Within the trend of increasing video submissions there were distinct seasonal patterns.
+Seasonal variation in cyclists’ submission reflect seasonal (weather and other) influences on cycling prevalence and duration.
+This suggests seasonal variation in preventive responses might be tailored to need.
+It is well established that crime is highly concentrated along whatever dimension is examined (Farrell & Pease, 2017).
+With respect to video evidence this will include recidivists (repeat offenders), repeat submitters of evidence (who may or may not be repeat victims/survivors), the concentration of incidents (close passes, near misses and crashes) at certain times and places, with different types of experiences concentrated among certain types of road users.
+For example, we hypothesise that close passes are more likely to be reported by cyclists and horse riders than vehicle drivers, both due to the nature of the road user interactions and the relative risk associated with the manoeuvre.
+As such, the types of incidents reported by vehicle drivers would be systematically different, indicating a need for different investigative and preventive approaches.
+Policing and preventions efforts focused on where problems are concentrated are more resource efficient – it is the foundation of problem-oriented policing (see e.g., (Laing, 2010; Scott & Maddox, 2009)- and research using video submissions holds the potential to analysis of problem concentrations.
+
+Evidence submitted by cyclists was the focus here.
+However, Operation Snap data offer potential for analysis of data from other road user sources.
+Studies focused on specific types of road user will prove informative with respect to policy and practice.
+Dash cam submissions by motorists offers great potential to inform other aspects of road safety.
+Research into submissions by pedestrians, horse riders and motorcyclists should prove viable to improve the safety of these parties.
+There were over four hundred submissions by horse riders in West Yorkshire, which means there will be thousands nationwide.
+The British Horse Society^[
+https://www.bhs.org.uk/
+]
+and others may be interested in this data being used to inform the safety of horses and riders. The British Motorcyclists Federation^[
+https://www.britishmotorcyclists.co.uk/
+]
+may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent.
+
+This was a case study of West Yorkshire but each of the 43 police forces in England and Wales collates OpSnap data, and our preliminary scoping of individual force websites suggests OpSnap data is publicly available nationwide.
+This suggests the potential for development of a national OpSnap dataset.
+A feasibility study is needed to establish the availability and compatibility of the data from different forces.
+Should a national dataset prove feasible it holds potential for national-level, cross-regional, and comparative analyses of patterns and trends.
+A national dataset will enhance the possibility for cross-national comparative analysis and international cooperation in road safety.
+It holds the potential for the development of rankings according to different criteria and, thereby, potential performance indicators.
+
+Comparison by complainant mode of transport, reported offense type and disposal, subdivided by other characteristics, is needed.
+Determining which type of submission - by which type of road user for example - is more likely to result in a recommendation of court proceedings, may help refine police investigations.
+For example, are submissions by cyclists (or horse riders) more or less likely to result in court proceedings than those by motorists?
+If so, why?
+What is the role of substantive issues, and to what extent are decisions affected by video quality?
+Further work with police investigators should inform best practice in the processing and further investigation of submissions.
+How are disposal options determined?
+Cross-police force comparative analysis may inform national best practice guidelines.
+
+A constraint on the present analysis was the nature of the publicly-available data.
+A pilot project should be undertaken in collaboration with police to establish the potential to further enhance policing and public safety using non-public aspects of the submissions data.
+This would require working partnerships and a secure research platform to meet GDPR requirements.
+Such collaborative approaches are increasingly common in health, medicine, and policing research.
+
+Video footage holds significant potential for further analysis, both qualitative and quantitative.
+Police investigations might be improved by research to identify and promote good practice in the assessment of footage, its use in determining disposals, further investigations, and prosecutions.
+What type of footage works best in the courtroom, and how is it best identified and prepared?
+What is the potential for machine learning to identify, clean, and prepare footage of the most serious submissions?
+There are also likely to be lessons that can be learned for how footage is gathered, edited and submitted by road users.
+More generally, video footage might be used to identification risk factors, that is, the types of situations in which crashes, near misses, and close passes occur.
+This promises the potential to inform policy and practice in ways that ameliorate risk factors.
+Within this area of research, analysis from dash cams is likely to inform different practices than that from cycle-cams, that from pedestrians, horse riders and motorcyclists, and so on.
+
+Some research into the relationship between video footage and supporting written evidence is needed.
+Is written supported evidence always needed?
+Which is deemed most important by police, and which by courts?
+What are the characteristics of strong supporting written evidence, and what are the characteristics of strong video evidence?
+Are both needed or can a weakness in one be overcome by particularly strong aspects in the other?
+
+The spatial analysis undertaken here was obliged to use data for ‘all’ video submissions.
+We were unable to separate the submissions by cyclists because, in the publicly-available data, locations could not be matched to individual cases: they were scrambled to ensure anonymity for GDPR purposes.
+An obvious next step, in the context of a secure research platform, would be spatial analysis for different types of road users, for different types of incidents, for incidents resulting in different disposals, and so on.
+Hotspots and spatial clustering are likely to vary by type of road user, type of incident, day or week, time of day, and so on.
+
+Creating opportunities to access video submission data would be an alternate approach that facilitates research into the nature of footage.
+Another option would be research using existing online videos, of which there appear to be many thousands on social media including those posted by police forces.
+Such research will face different methodological challenges but could provide proof-of-concept of the potential for analysis of OpSnap footage.
+
+OpSnap data should be compared to, and integrated with, other road safety datasets.
+For example, spatial analysis that was Integrated with Strava data (road use data by cyclists and pedestrians), Google road use and other data, may facilitate exposure-based measures rather than the counts used here.
+The use of rates will enhance the identification locations with higher risk because based on their volume of traffic: different types of risk need to be gauged.
+Comparison and integration of OpSnap and Stats193 is an obvious starting point.
+To what extent do OpSnap and Crashmap incidents overlap?
+To what extent are they complementary?
+To what extent do they identify the same or different types of problems?
+
+More generally, OpSnap data holds potential for use in the evaluation of experimental interventions.
+Depending on the level of submissions there is the potential to use the data to assess effects before and after interventions, and to identify comparison (control group) sites.
+Different road safety interventions imply different resource needs.
+Ideally, evaluations would include cost-benefit analysis of the portfolio of social and economic costs involved, including the cost of death and injury.
+
+# Conclusion
+
+This study offers what we believe is the first study using the open access Operation Snap data.
+It provides proof-of-concept that OpSnap data holds considerable research potential.
+It has the potential to inform policy and practice promoting road safety.
+Drawing on the analysis of the West Yorkshire data, a research agenda was outlined.
+This preliminary list will prove far from exhaustive.
+While aspects of that agenda may appear ambitious at the time of writing, we suggest that the present study will have achieved one of its aims if this agenda is rapidly superseded by further research.
+
+The focus here was video evidence submissions by cyclists.
+Relative to distance travelled by cyclists, their submissions were around 20 times over-represented in the West Yorkshire data.
+This was the justification for our focus.
+The disproportionality likely reflects the vulnerability of cyclists relative to motor vehicle users when the two forms of transport interact, and when close passes, near misses and crashes occur.
+Our primary conclusion is that further research into video submissions to Operation Snap holds significant potential to inform the future of road safety for a broad range of road users.
+
 # References

### 2024-05-20 paper.qmd

@@
 ```{r}
-#| label: fig-time
-#| fig-cap: "Monthly count of Operation Snap records, West Yorkshire"
 d_all_monthly = d_all |>
@@
 d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
+```
+
+```{r}
+#| label: stats19
+#| include: false
+years = 2021:2022
+dir.create("data/stats19", recursive = TRUE, showWarnings = FALSE)
+collisions = stats19::get_stats19(year = 2022)
+# collisions_2021 = stats19::get_stats19(year = 2021)
+# collisions = purrr::map_df(years, stats19::get_stats19, type = "collision", data_dir = "data/stats19")
+stats19_monthly = collisions |>
+  filter(police_force == "West Yorkshire") |>
+  mutate(month = lubridate::floor_date(date, "month")) |>
+  group_by(month) |>
+  summarise(n = n()) |>
+  mutate(records = "stats19")
+```
+
+```{r}
+#| label: fig-time
+#| fig-cap: "Monthly count of Operation Snap records, West Yorkshire"
+d_monthly = bind_rows(d_monthly, stats19_monthly)
 d_monthly |>

### 2024-05-20 paper.qmd
Update paper.qmd
tweaked abstract

@@
 authorcols: true # Should authors be listed in a single column (default) or in multiple columns (`authorcols: true`)
+editor_options: 
+  markdown: 
+    wrap: sentence
 ---
@@
 
-This study uses data from Operation Snap (OpSnap), the UK police’s national system to receive road users’ video evidence of road traffic offences.
-Data from one police force area for three calendar years, from its launch in 2021 (N= 18,363 records), is analysed.
-Of submitted cases, 34.4% were from cyclists (n=6,312) of which 69% resulted in an offence being registered, mostly for driving ‘without reasonable consideration to others’ or ‘without due care and attention’.
+This study uses data from Operation Snap (OpSnap), the UK police's national system to receive road users' video evidence of road traffic offences.
+Data from one police force area for the three calendar years 2021 to 2023 (N= 18,363 records), is analysed.
+Of submitted cases, 34.4% were from cyclists (n=6,312) of which 69% resulted in an offence being registered, mostly for driving 'without reasonable consideration to others' or 'without due care and attention'.
 Three quarters (75.2%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), a quarter in no further action, and fewer than 1% for court appearance.
@@
 Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
-The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” (Department for Transport, 2023b).
+The 2023 report noted that "the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was 'driver or rider failed to look properly'" (Department for Transport, 2023b).
 
@@
 ```
-
 Cyclist crashes are underreported in police recorded datasets (Elvik & Mysen, 1999) and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted (Janstrup et al., 2016).
 Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions (Ibrahim et al., 2021).
-These incidents are often referred to as “near-misses” (Ibrahim et al., 2021).
+These incidents are often referred to as "near-misses" (Ibrahim et al., 2021).
 
@@
 Close passes are the most common type of near miss reported by cyclists, and are associated with collisions resulting in injury [@aldred2016].
-A ‘close pass’ refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
-Close passes take different forms including the ‘punishment pass’ by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
-There is no specific law in the UK Road Traffic Act 1988 for driving too close to cyclist, but two are commonly applied for careless driving: RT88575- Drive without due care and attention; and RT99576 – Driver without reasonable consideration to others - and these are prominent in the analysis that follows.
+A 'close pass' refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
+Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
+There is no specific law in the UK Road Traffic Act 1988 for driving too close to cyclist, but two are commonly applied for careless driving: RT88575- Drive without due care and attention; and RT99576 -- Driver without reasonable consideration to others - and these are prominent in the analysis that follows.
 
-This study highlights the potential of using open access data from Operation Snap, the UK police’s national system to receive road users’ video evidence of road traffic offences.
-Operation Snap, often referred to informally as ‘OpSnap’, was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 20181.
+This study highlights the potential of using open access data from Operation Snap, the UK police's national system to receive road users' video evidence of road traffic offences.
+Operation Snap, often referred to informally as 'OpSnap', was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 20181.
 It is now in operation nationally across England and Wales, each police force offering its own submission portal for road users to submit video evidence.
@@
 -   The public should be prepared to sign a witness statement and possibly give evidence in court. 
-- Statements for OpSnap can only be accepted from persons aged 18 or over. If you are under 18 the incident should be reported by email to [email]2.
+-   Statements for OpSnap can only be accepted from persons aged 18 or over. If you are under 18 the incident should be reported by email to \[email\]2.
 
@@
 
-Open access data from West Yorkshire Safer Roads – OpSnap was used for this study, the media submissions portal having opened in July 2020.
+Open access data from West Yorkshire Safer Roads -- OpSnap was used for this study, the media submissions portal having opened in July 2020.
 The West Yorkshire Police (WYP) data used here spanned calendar years 2021, 2022 and 2023.
@@
 The dataset is a tabulation of cases submitted to the West Yorkshire Safer Roads OpSnap web portal.
-The terms ‘record’ and ‘case’ are used interchangeably here to refer to a record in the OpSnap database, each of which represents the separate submission of video evidence by a road user.
+The terms 'record' and 'case' are used interchangeably here to refer to a record in the OpSnap database, each of which represents the separate submission of video evidence by a road user.
 
@@
 Video footage is commonly recorded from on-board cameras.
-For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as ‘dash cams’.
+For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as 'dash cams'.
 One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam (iCompario, 2024).
@@
 The open access data is a deidentified summary of submitted cases with informtion on: mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
-The offence location is typically a street name and town or city name, or an intersection, and examples included: ‘A58 Godley Road, Halifax’, ‘Keighley Road, Silsden’ , ‘Woodhouse Lane A660, Leeds’.
+The offence location is typically a street name and town or city name, or an intersection, and examples included: 'A58 Godley Road, Halifax', 'Keighley Road, Silsden' , 'Woodhouse Lane A660, Leeds'.
 For this study, approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
@@
 Over 90 percent of offences reported by cyclists were for driving without consideration for other road users and without due care and attention.
-This is consistent with the phenomenon of ‘close passes’ in the cycling safety literature (Aldred, 2016; Cubbin et al., 2024).
+This is consistent with the phenomenon of 'close passes' in the cycling safety literature (Aldred, 2016; Cubbin et al., 2024).
 It suggests the OpSnap data holds the potential for further analysis to inform knowledge about the nature of close passes generally.
@@
 Within the trend of increasing video submissions there were distinct seasonal patterns.
-Seasonal variation in cyclists’ submission reflect seasonal (weather and other) influences on cycling prevalence and duration.
+Seasonal variation in cyclists' submission reflect seasonal (weather and other) influences on cycling prevalence and duration.
 This suggests seasonal variation in preventive responses might be tailored to need.
@@
 As such, the types of incidents reported by vehicle drivers would be systematically different, indicating a need for different investigative and preventive approaches.
-Policing and preventions efforts focused on where problems are concentrated are more resource efficient – it is the foundation of problem-oriented policing (see e.g., (Laing, 2010; Scott & Maddox, 2009)- and research using video submissions holds the potential to analysis of problem concentrations.
+Policing and preventions efforts focused on where problems are concentrated are more resource efficient -- it is the foundation of problem-oriented policing (see e.g., (Laing, 2010; Scott & Maddox, 2009)- and research using video submissions holds the potential to analysis of problem concentrations.
 
@@
 There were over four hundred submissions by horse riders in West Yorkshire, which means there will be thousands nationwide.
-The British Horse Society^[
-https://www.bhs.org.uk/
-]
-and others may be interested in this data being used to inform the safety of horses and riders. The British Motorcyclists Federation^[
-https://www.britishmotorcyclists.co.uk/
-]
-may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent.
+The British Horse Society[^1] and others may be interested in this data being used to inform the safety of horses and riders.
+The British Motorcyclists Federation[^2] may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent
+.
+
+[^1]:  https://www.bhs.org.uk/
+
+[^2]:  https://www.britishmotorcyclists.co.uk/
 
@@
 
-The spatial analysis undertaken here was obliged to use data for ‘all’ video submissions.
+The spatial analysis undertaken here was obliged to use data for 'all' video submissions.
 We were unable to separate the submissions by cyclists because, in the publicly-available data, locations could not be matched to individual cases: they were scrambled to ensure anonymity for GDPR purposes.

### 2024-05-20 paper.qmd
tweak title

@@
 ---
-title: "Cyclist video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda"
+title: "Road User video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda"
 bibliography: references.bib

### 2024-05-20 paper.qmd
Update paper.qmd

@@
 #| include: false
+file_exists = file.exists(file_name)
+if (!file_exists) {
+  # stop("try downloading the data first!")
+  u = "https://github.com/ITSLeeds/opsnap/releases/download/v1/operation_snap_2024-04.csv"
+  f = basename(u)
+  download.file(url = u, destfile = f)
+  dir.create("data/west-yorkshire", recursive = TRUE)
+  file.copy(f, "data/west-yorkshire/")
+}
 d_all = read_csv(file_name)

### 2024-05-20 paper.qmd
Update paper.qmd

@@
 ```
+new sentence here.
+
 Cyclist crashes are underreported in police recorded datasets (Elvik & Mysen, 1999) and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted (Janstrup et al., 2016).

### 2024-07-01 paper.qmd
Test
Update with cjs codes

@@
 format:
+  html: default
   # pdf: default
-  arxiv-pdf:
-    keep-tex: true
+  # arxiv-pdf:
+  #   keep-tex: true
 linenumbers: true # Add (continuous) line numbers?
@@
 )
+```
+
+```{r}
+#| eval: false
+#| echo: false
+browseURL("paper.docx")
 ```
@@
     offence = tolower(offence),
-    disposal = tolower(disposal)
+    disposal = tolower(disposal),
+    # First word of text in offence column:
+    offence_code = str_extract(offence, "\\w+")
   )
@@
   mutate(records = "complete_offence")
-d = d_offence |>
+d_with_location = d_offence |>
   opsnap:::filter_location_nas()
-d_complete_monthly = d |>
+d_complete_monthly = d_with_location |>
   mutate(month = lubridate::floor_date(date, "month")) |>
@@
 Half of cases were reported by vehicle drivers, a third by cyclists, seven percent by pedestrians, with over two percent by horse riders and less than one percent by motorcyclists.
+
+```{r}
+#| label: tbl-mode-offences
+#| include: true
+# Aim: get table of n. offences by mode
+d_mode_offence_count = d_all |>
+  count(mode, offence, sort = TRUE) 
+# offences in order of n. offences
+d_mode_offence_count |>
+  group_by(offence) |>
+  summarise(n = sum(n)) |>
+  arrange(desc(n)) |>
+  head(20) |>
+  knitr::kable()
+```
+
+```{r}
+# reclassify offences into types
+x = d_all
+regroup_offences = function(x) {
+ x |> 
+   dplyr::mutate(
+     offence_classified = dplyr::case_when(
+       stringr::str_detect(offence, "")
+     )
+   )
+}
+```
 

### 2024-07-01 paper.qmd
Add opsnap bib

@@
 title: "Road User video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda"
-bibliography: references.bib
+bibliography: [references.bib, opsnap.bib]
 editor:
@@
 Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
-Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@dft-xxxx].
-Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (Department for Transport, 2023a).
-Between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year in Britain according to police records (Department for Transport, 2023b).
+Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@departmentfortransport_reported_2023].
+Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics [@departmentfortransport_reported_2023].
+Between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year in Britain according to police records [@departmentfortransport].
 Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
-The 2023 report noted that "the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was 'driver or rider failed to look properly'" (Department for Transport, 2023b).
+The 2023 report noted that "the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was 'driver or rider failed to look properly'" [@departmentfortransport].
 
@@
 
-There were `r nrow(d_all)` ecords in the dataset for the three-year study period, shown by month in @fig-time.
+There were `r nrow(d_all)` records in the dataset for the three-year study period, shown by month in @fig-time.
 The number of records increased over time.
-Some records lacked either an offence or a location, leaving `r round(nrow(d) / nrow(d_all) * 100, 1)`% or `r nrow(d)` complete records.
+Some records lacked either an offence or a location, leaving `r round(nrow(d_with_location) / nrow(d_all) * 100, 1)`% or `r nrow(d_with_location)` complete records.
 There was a distinct seasonal pattern to reporting, with significant increases in summer months.
@@
 
-There are `r unique(d$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
+There are `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
 
@@
 #| tbl-cap: "Most common locations recorded in the OpSnap dataset"
-d |>
+d_with_location |>
   count(location, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d)) |>
+  mutate(percent_records = n / nrow(d_with_location)) |>
   mutate(percent_records = scales::percent(round(percent_records, 3))) |>
@@
 
-Cyclists undertake two percent of road miles in the UK (Department for Transport, 2023b).
+Cyclists undertake two percent of road miles in the UK [@departmentfortransport].
 Other things equal, this suggests that with respect to distance travelled, cyclists were around 20 times more likely to submit evidence of a road traffic offence than vehicle drivers and other road users in West Yorkshire.

### 2024-07-01 paper.qmd
Update readme

@@
 
-Cyclist crashes are underreported in police recorded datasets (Elvik & Mysen, 1999) and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted (Janstrup et al., 2016).
+Cyclist crashes are underreported in police recorded datasets [@elvik_incomplete_1999] and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted (Janstrup et al., 2016).
 Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions (Ibrahim et al., 2021).

### 2024-07-01 paper.qmd
Update paper

--- 2024-07-01 paper.qmd
@@
     wrap: sentence
+execute: 
+  cache: true
+  echo: false
+  warning: false
 ---
@@
   comment = "#>",
-  echo = FALSE,
   warning = FALSE
@@
 Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
-Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@departmentfortransport_reported_2023].
-Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics [@departmentfortransport_reported_2023].
+Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@departmentfortransportReportedRoadCasualties2023].
+Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics [@departmentfortransportReportedRoadCasualties2023].
 Between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year in Britain according to police records [@departmentfortransport].
@@
 ```
-new sentence here.
 
-Cyclist crashes are underreported in police recorded datasets (Elvik & Mysen, 1999) and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted (Janstrup et al., 2016).
-Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions (Ibrahim et al., 2021).
-These incidents are often referred to as "near-misses" (Ibrahim et al., 2021).
+Cyclist crashes are underreported in police recorded datasets [@elvikIncompleteAccidentReporting1999] and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted [@janstrupetal.UnderstandingTrafficCrash2016].
+Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions [@ibrahimCyclingMissesReview2021].
+These incidents are often referred to as "near-misses" [@ibrahimCyclingMissesReview2021]
 
-Commuter cyclists in the UK experience a near miss for every six miles of riding (Aldred & Crosweller, 2015), and concern over near misses are a key reason people choose not to cycle (Sanders, 2015), being associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers (Sanders, 2015).
-In their review of near miss cycling crashes, Ibrahim et al., (2021) highlighted the need for better data on near misses to inform safety research.
+Commuter cyclists in the UK experience a near miss for every six miles of riding [@aldredInvestigatingRatesImpacts2015a], and concern over near misses are a key reason people choose not to cycle [@sandersPerceivedTrafficRisk2015a], being associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers [@sandersPerceivedTrafficRisk2015a].
+In their review of near miss cycling crashes, @ibrahimCyclingMissesReview2021 highlighted the need for better data on near misses to inform safety research.
 
@@
 For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as 'dash cams'.
-One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam (iCompario, 2024).
+One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam [@DashCamSubmissions2024].
 For cyclists, footage is commonly recorded using helmet or handlebar-mounted cameras.
@@
   count(mode, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
+  mutate(percent_of_records = n / nrow(d_all)) |>
+  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
   arrange(desc(n)) |>
   rename_all(snakecase::to_title_case) |>
+  # Rename N to "Number of records"
+  rename(`Number of Records` = N) |>
   knitr::kable()
@@
   count(offence, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
+  mutate(percent_of_records = n / nrow(d_all)) |>
+  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
   arrange(desc(n)) |>
@@
   count(offence, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_offence)) |>
-  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
+  mutate(percent_of_records = n / nrow(d_offence)) |>
+  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
   arrange(desc(n)) |>
@@
   count(offence, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_offence)) |>
-  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
+  mutate(percent_of_records = n / nrow(d_offence)) |>
+  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
   mutate(
@@
   count(disposal, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_all)) |>
-  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
+  mutate(percent_of_records = n / nrow(d_all)) |>
+  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
   arrange(desc(n)) |>
@@
   count(location, sort = TRUE) |>
-  mutate(percent_records = n / nrow(d_with_location)) |>
-  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
+  mutate(percent_of_records = n / nrow(d_with_location)) |>
+  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
   arrange(desc(n)) |>
@@
 Over 90 percent of offences reported by cyclists were for driving without consideration for other road users and without due care and attention.
-This is consistent with the phenomenon of 'close passes' in the cycling safety literature (Aldred, 2016; Cubbin et al., 2024).
+This is consistent with the phenomenon of 'close passes' in the cycling safety literature [@aldred2016; @cubbin2024].
 It suggests the OpSnap data holds the potential for further analysis to inform knowledge about the nature of close passes generally.
@@
 This suggests seasonal variation in preventive responses might be tailored to need.
-It is well established that crime is highly concentrated along whatever dimension is examined (Farrell & Pease, 2017).
+It is well established that crime is highly concentrated along whatever dimension is examined [@farrellPreventingRepeatRepeat2017].
 With respect to video evidence this will include recidivists (repeat offenders), repeat submitters of evidence (who may or may not be repeat victims/survivors), the concentration of incidents (close passes, near misses and crashes) at certain times and places, with different types of experiences concentrated among certain types of road users.
@@
 As such, the types of incidents reported by vehicle drivers would be systematically different, indicating a need for different investigative and preventive approaches.
-Policing and preventions efforts focused on where problems are concentrated are more resource efficient -- it is the foundation of problem-oriented policing (see e.g., (Laing, 2010; Scott & Maddox, 2009)- and research using video submissions holds the potential to analysis of problem concentrations.
+Policing and preventions efforts focused on where problems are concentrated are more resource efficient -- it is the foundation of problem-oriented policing (see e.g., (@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010)- and research using video submissions holds the potential to analysis of problem concentrations.
 

### 2024-07-01 R/get_cjs_codes.R
Document

@@ -1,0 +1,20 @@
+#' Get Criminal Justice Codes
+#' 
+#' This function gets the criminal justice codes from the Criminal Justice Hub website.
+#' See [www.criminaljusticehub.org.uk](https://www.criminaljusticehub.org.uk/jargon-buster/cjs-offence-code/)
+#' 
+#' @export 
+#' @examples
+#' get_cjs_codes()
+get_cjs_codes = function(
+    u = paste0(
+        "https://www.criminaljusticehub.org.uk/wp-content/uploads/",
+        "2022/10/cjs-offence-index-sept-2022.ods"
+    )
+) {
+    f = file.path(tempdir(), "cjs_codes.ods")
+    message("Downloading CJS codes from ", u, " to ", f)
+    utils::download.file(u, f, mode = "wb")
+    cjs_codes = readODS::read_ods(f, sheet = 1)
+    cjs_codes
+}

### 2024-07-01 paper.qmd
md version

@@
 format:
-  html: default
+  gfm: default
   # pdf: default
@@
 execute: 
-  cache: true
+  cache: refresh
   echo: false
@@
 ```
-
 Cyclist crashes are underreported in police recorded datasets [@elvikIncompleteAccidentReporting1999] and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted [@janstrupetal.UnderstandingTrafficCrash2016].
@@
 d = opsnap:::download_and_read(u)
-names(d_with_location)
 # Old names:
@@
   opsnap:::filter_offence_nas()
+# table(d_offence$offence) |> sort() |> tail()
+#                         rv86019 use a handheld phone / device whilst driving a motor vehicle on a road 
+#                                                                                                    357 
+#                                  rt88966 motor vehicle fail to comply with endorsable s36 traffic sign 
+#                                                                                                    411 
+#                                                          rt88971 fail to comply with red traffic light 
+#                                                                                                    679 
+# rt88975 drive motor vehicle fail to comply with red / green arrow / lane closure traffic light signals 
+#                                                                                                   1364 
+#                                                           rt88575 drive without due care and attention 
+#                                                                                                   2917 
+#                                               rt88576 drive without reasonable consideration to others 
+#                                                                                                   4992 
 d_offence_monthly = d_offence |>
@@
   summarise(n = n()) |>
-  mutate(records = "complete_offence")
-d_with_location = d_offence |>
+  mutate(records = "with_offence")
+d_with_location = d_all |>
   opsnap:::filter_location_nas()
-d_complete_monthly = d_with_location |>
+d_complete = d_offence |>
+  opsnap:::filter_location_nas()
+d_complete_monthly = d_complete |>
   mutate(month = lubridate::floor_date(date, "month")) |>
@@
 d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
+# d_monthly = bind_rows(d_all_monthly, d_offence_monthly, d_complete_monthly)
 ```
@@
 
-There were `r nrow(d_all)` records in the dataset for the three-year study period, shown by month in @fig-time.
-The number of records increased over time.
-Some records lacked either an offence or a location, leaving `r round(nrow(d_with_location) / nrow(d_all) * 100, 1)`% or `r nrow(d_with_location)` complete records.
+There were `r nrow(d_all)` records in the dataset for the three-year study period, with a strong upward trend, as shown in the monthly counts presented in @fig-time.
+Some records lacked either an offence (`r nrow(d_all) - nrow(d_offence)`, `r round((nrow(d_all) - nrow(d_offence)) / nrow(d_all) * 100, 1)`%) or a location (`r nrow(d_offence) - nrow(d_complete)`, `r round((nrow(d_offence) - nrow(d_complete)) / nrow(d_all) * 100, 1)`%), or both, leaving `r round(nrow(d_complete) / nrow(d_all) * 100, 1)`% or `r nrow(d_complete)` complete records.
 There was a distinct seasonal pattern to reporting, with significant increases in summer months.
@@
 #| include: true
+#| fig-cap: "Number of convictions by offence type (top 6 and other)."
 # Aim: get table of n. offences by mode
@@
 # offences in order of n. offences
-d_mode_offence_count |>
+d_offence_count = d_all |>
+  count(offence, sort = TRUE)
+# d_offence_count |>
+#   arrange(desc(n)) |>
+#   head(20) |>
+#   knitr::kable()
+# |offence                                                                                                |    n|
+# |:------------------------------------------------------------------------------------------------------|----:|
+# |n/a                                                                                                    | 5706|
+# |rt88576 drive without reasonable consideration to others                                               | 4992|
+# |rt88575 drive without due care and attention                                                           | 2917|
+# |rt88975 drive motor vehicle fail to comply with red / green arrow / lane closure traffic light signals | 1364|
+# |rt88971 fail to comply with red traffic light                                                          |  679|
+# |rt88966 motor vehicle fail to comply with endorsable s36 traffic sign                                  |  411|
+# |rv86019 use a handheld phone / device whilst driving a motor vehicle on a road                         |  357|
+# |rt88760 fail to comply with solid white lines                                                          |  265|
+# |rt88751 contravene give way sign                                                                       |  264|
+# |suspected contravene weight restriction.                                                               |  213|
+# |rt88751 contravene mandatory direction arrows                                                          |  212|
+# |me82009 driving on hard shoulder of motorway                                                           |  113|
+# |rt88975 fail to comply with red traffic light                                                          |  109|
+# |rt88751 motor vehicle fail to comply with a non-endorsable traffic sign other (specify)                |   91|
+# |zp97004 fail to comply with red light pelican crossing                                                 |   84|
+# |zp97003 stop within controlled area of pelican crossing                                                |   81|
+# |rc86814 driver not in proper control of vehicle                                                        |   62|
+# |zp97001 stop vehicle within limits of pelican crossing                                                 |   59|
+# |hy35001 drive/ride on footpath beside a road                                                           |   42|
+# |rr84171 vehicle contravene local traffic order other than parking (e.g. bus lane)                      |   41|
+# Pull out the top 6 offences excluding n/a:
+d_offence_top_6 = d_mode_offence_count |>
+  filter(offence != "n/a") |>
   group_by(offence) |>
@@
   arrange(desc(n)) |>
-  head(20) |>
+  head(6)
+d_offence_classified = d_offence_count |>
+  filter(offence != "n/a") |>
+  mutate(
+    Offence = case_when(
+      offence %in% d_offence_top_6$offence ~ offence,
+      TRUE ~ "Other"
+    )
+  ) |>
+  group_by(Offence) |>
+  summarise(`Number of records` = sum(n)) |>
+  # Arrange in descending order of n except for "Other":
+  arrange(Offence == "Other", desc(`Number of records`)) |>
+  mutate(`Percent of records` = round(`Number of records` / sum(`Number of records`) * 100, 1))
+d_offence_classified |>
   knitr::kable()
@@
 x = d_all
-regroup_offences = function(x) {
+regroup_offences_simple = function(x) {
  x |> 
    dplyr::mutate(
-     offence_classified = dplyr::case_when(
-       stringr::str_detect(offence, "")
+     offence_simple = dplyr::case_when(
+       stringr::str_detect(offence, "drive without reasonable consideration") ~ "Inconsiderate driving",
+       stringr::str_detect(offence, "drive without due care") ~ "Careless driving",
+       TRUE ~ "Other"
+
      )
@@
 }
+table(x$mode)
+regroup_modes = function(x) {
+  x |> 
+    dplyr::mutate(
+      mode_simplified = dplyr::case_when(
+      mode == "cyclist" ~ "Cyclist",
+      mode == "vehicle driver" ~ "Driver",
+      TRUE ~ "Other"
+    ))
+}
+
+d_all = regroup_offences_simple(d_all)
+d_all = regroup_modes(d_all)
+
 ```
@@
   mutate(percent_of_records = n / nrow(d_all)) |>
-  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
+  mutate(percent_of_records = round(percent_of_records, 3) * 100) |>
   arrange(desc(n)) |>
@@
 
-For cases submitted by people riding cycles (shown in @tbl-offences-cyclist-observer), the most common offences were also both associated with careless driving, particularly driving without reasonable consideration to others (78.7%), there were also a small proportion of cases associated with drivers using mobile phones (3.7%), failing to comply with traffic signals (3.5%) and contravening regulator signage (0.6%).
+<!-- For cases submitted by people riding cycles (shown in @tbl-offences-cyclist-observer), the most common offences were also both associated with careless driving, particularly driving without reasonable consideration to others (78.7%), there were also a small proportion of cases associated with drivers using mobile phones (3.7%), failing to comply with traffic signals (3.5%) and contravening regulator signage (0.6%).
+-->
+
+```{r}
+#| label: tbl-mode-offences-crosstab
+#| fig-cap: Mode of transport from which video was recorded (columns) and offence type (rows), with counts and percentages within each reporting mode.
+
+# | Offence (grouped)       |               | Total |
+# |-------------------------|---------------|-------|
+# |                         | Dangerous driving | Other offences | 
+# |-------------------------|-------------------|-----------------|
+# | **Reporter transport mode** |                   |                 |
+# | Vehicle driver          | **Count**           | **5511**        | **654** | **6165** |
+# |                         | % within Reporter transport mode | **89.4%** | **10.6%** | **100.0%** |
+# | Cyclist                 | **Count**           | **4469**        | **249** | **4718** |
+# |                         | % within Reporter transport mode | **94.7%** | **5.3%**  | **100.0%** |
+# | Other                   | **Count**           | **1279**        | **220** | **1499** |
+# |                         | % within Reporter transport mode | **85.3%** | **14.7%** | **100.0%** |
+# | **Total**               | **Count**           | **11259**       | **1123** | **12382** |
+# |                         | % within Reporter transport mode | **90.9%** | **9.1%**  | **100.0%** |
+
+# R version:
+pivot_counts = d_all |>
+  filter(offence != "n/a") |>
+#   mutate(
+#     offence_simple = case_when(
+#       offence %in% d_offence_top_6$offence ~ offence,
+#       TRUE ~ "Other"
+#   )
+# ) |>
+  group_by(mode_simplified, offence_simple) |>
+  summarise(n = n()) |>
+  pivot_wider(names_from = mode_simplified, values_from = n) |> 
+  mutate(
+    Total = Cyclist + Driver + Other
+  ) |>
+  arrange(offence_simple == "Other", desc(Total))
+pivot_counts_total = tibble::tibble(
+  offence_simple = "Total",
+  Cyclist = sum(pivot_counts$Cyclist),
+  Driver = sum(pivot_counts$Driver),
+  Other = sum(pivot_counts$Other),
+  Total = sum(pivot_counts$Total)
+)
+
+pivot_percents = pivot_counts |>
+  # Calculate percentage of each mode
+  mutate(
+    `Cyclist (%)` = Cyclist / sum(Cyclist) * 100,
+    `Driver (%)` = Driver / sum(Driver) * 100,
+    `Other (%)` = Other / sum(Other) * 100,
+    `Total (%)` = Total / sum(Total) * 100
+  )
+
+pivot_percents_totals = pivot_percents |>
+  group_by(offence_simple = "Total") |>
+  summarise_if(is.numeric, sum)
+
+pivot_combined = bind_rows(pivot_percents, pivot_percents_totals)
+
+percent_careless = pivot_combined |> filter(offence_simple == "Careless driving") |> pull(`Total (%)`)
+percent_inconsiderate = pivot_combined |> filter(offence_simple == "Inconsiderate driving") |> pull(`Total (%)`)
+percent_careless_cycling = pivot_combined |> filter(offence_simple == "Careless driving") |> pull(`Cyclist (%)`)
+percent_inconsiderate_cycling = pivot_combined |> filter(offence_simple == "Inconsiderate driving") |> pull(`Cyclist (%)`)
+percent_other = pivot_combined |> filter(offence_simple == "Other") |> pull(`Total (%)`)
+percent_other_cyclist = pivot_combined |> filter(offence_simple == "Other") |> pull(`Cyclist (%)`)
+percent_other_driver = pivot_combined |> filter(offence_simple == "Other") |> pull(`Driver (%)`)
+percent_other_other = pivot_combined |> filter(offence_simple == "Other") |> pull(`Other (%)`)
+
+pivot_combined |>
+  knitr::kable(digits = 1)
+
+```
+
+While `r round(percent_inconsiderate, 1)` percent of all offences are for inconsiderate driving, they make up the bulk of offences reported by cyclists, with `r round(percent_inconsiderate_cycling, 1)` percent of all offences reported by cyclists being for inconsiderate driving.
+Drivers are proportionally `r round(percent_other_driver / percent_other_cyclist, 1)` times more likely to report other types of offences as cyclists, while other reporting modes are most likely to report other types of offences, being `r round(percent_other_other / percent_other_cyclist, 1)` times more likely to report other types of offences as cyclists.
+While further research is needed, this makes sense if physically-vulnerable cyclists are most concerned with the dangerous driving of vehicles, whereas drivers also report other types of road traffic offence, with pedestrians and others also more likely to encounter other types of offence.
 
@@
 #| tbl-cap: "Number and percentages of OpSnap records, submitted by cyclists, by offence type."
+#| include: false
 d_all |>

### 2024-07-01 R/opsnap.R

@@ -44,7 +44,8 @@ select_columns = function(d) {
 
 filter_offence_nas = function(d) {
   d |>
-    dplyr::filter(!offence == "N/A") |>
+    # Filter out "n/a" values (case insensitive):
+    dplyr::filter(!grepl("n/a", offence, ignore.case = TRUE)) |>
     dplyr::filter(!is.na(offence))
 }
 

### 2024-07-01 paper.qmd
Update paper

@@
 ```{r setup}
-library(opsnap)
+#| include: false
+devtools::load_all()
 ```
@@
 
-Mode of transport of the complainant is shown in @tbl-mode.
-Half of cases were reported by vehicle drivers, a third by cyclists, seven percent by pedestrians, with over two percent by horse riders and less than one percent by motorcyclists.
+Of the `r nrow(d_offence)` records with an offence, the most common offences were Driving without reasonable consideration to others (rt88576) and Driving without due care and attention (rt88575).
+This is presented in @tbl-offences, which shows the number and percentage of records by offence type, showing the top 6 offence types and the remainder grouped as 'Other'.
+As previously noted, a high proportion of cases did not include information on the reported offence, with the most common offences being Driving without reasonable consideration to others (rt88576), and Driving without due care and attention (rt88575).
+Within the road traffic act these offences are related to careless driving and drivers are subject to similar penalties.
 
 ```{r}
-#| label: tbl-mode-offences
+#| label: tbl-offences
 #| include: true
-#| fig-cap: "Number of convictions by offence type (top 6 and other)."
+#| tbl-cap: "Number of convictions by offence type (top 6 and other)."
 # Aim: get table of n. offences by mode
@@
 }
-table(x$mode)
+# table(x$mode)
 regroup_modes = function(x) {
@@
 
+Vehicle (mostly car and van drivers with dashcams) driver and cyclist reporters dominate reporting for all records, as illustrated in @tbl-mode.
+Half of cases were reported by vehicle drivers, a third by cyclists, seven percent by pedestrians, with over two percent by horse riders and less than one percent by motorcyclists.
+
 ```{r}
@@
   rename(`Number of Records` = N) |>
-  knitr::kable()
-```
-
-Classification of the committed offence is shown in @tbl-offences.
-s previously noted, a high proportion of cases did not include information on the reported office (31.1%), with the most common offences being Driving without reasonable consideration to others (rt88576), and Driving without due care and attention (rt88575).
-Within the road traffic act these offences are related to careless driving and drivers are subject to similar penalties.
-
-```{r}
-#| label: tbl-offences
-#| tbl-cap: Number and percentages of OpSnap records by offence type.
-d_all |>
-  count(offence, sort = TRUE) |>
-  mutate(percent_of_records = n / nrow(d_all)) |>
-  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
-  arrange(desc(n)) |>
-  head(10) |>
-  rename_all(snakecase::to_title_case) |>
   knitr::kable()
@@
 ```{r}
-#| label: tbl-mode-offences-crosstab
-#| fig-cap: Mode of transport from which video was recorded (columns) and offence type (rows), with counts and percentages within each reporting mode.
+#| label: tbl-mode-offences-crosstab-prep
 
@@
 
-pivot_combined |>
-  knitr::kable(digits = 1)
-
 ```
 
+The cross-tabulation of reporting mode and offence type is shown in @tbl-mode-offences-crosstab.
+The table reveals some interesting tendencies.
 While `r round(percent_inconsiderate, 1)` percent of all offences are for inconsiderate driving, they make up the bulk of offences reported by cyclists, with `r round(percent_inconsiderate_cycling, 1)` percent of all offences reported by cyclists being for inconsiderate driving.
@@
 While further research is needed, this makes sense if physically-vulnerable cyclists are most concerned with the dangerous driving of vehicles, whereas drivers also report other types of road traffic offence, with pedestrians and others also more likely to encounter other types of offence.
+
+```{r}
+#| label: tbl-mode-offences-crosstab
+#| tbl-cap: Mode of transport from which video was recorded (columns) and offence type (rows), with counts and percentages within each reporting mode.
+pivot_combined |>
+  knitr::kable(digits = 1)
+```
 

### 2024-07-03 paper.qmd
Small update

@@
 format:
-  gfm: default
+  html: default
   # pdf: default
@@
 execute: 
-  cache: refresh
+  cache: true
   echo: false
@@
 #| echo: false
+quarto::quarto_render("paper.qmd", output_format = "pdf")
+quarto::quarto_render("paper.qmd", output_format = "docx")
+file.rename("paper.docx", "~/OneDrive/opsnap_leeds/paper-v6.docx")
+file.rename("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v6.pdf")
 browseURL("paper.docx")
@@
 ```{r}
-#| label: tbl-mode-offences-crosstab-prep
-
-# | Offence (grouped)       |               | Total |
+# # Offence (grouped)       |               | Total |
 # |-------------------------|---------------|-------|
@@
   knitr::kable(digits = 1)
+
 ```
@@
   count(disposal, sort = TRUE) |>
-  mutate(percent_of_records = n / nrow(d_all)) |>
-  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
+  mutate(percent_of_records = round(n / nrow(d_all), 3) * 100) |>
   arrange(desc(n)) |>
@@
   count(location, sort = TRUE) |>
-  mutate(percent_of_records = n / nrow(d_with_location)) |>
-  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
+  mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
   arrange(desc(n)) |>

### 2024-07-03 R/opsnap.R
Fix issue with filter_nas (#6)

@@ -10,7 +10,7 @@
 #' u = "https://www.westyorkshire.police.uk/sites/default/files/2023-10/operation_snap_july-sept_2023.xlsx"
 #' download_and_read(, dir = ".")
 #' }
-download_and_read = function(u, remove_nas = TRUE, dir = "raw_data/west-yorkshire") {
+download_and_read = function(u, remove_nas = FALSE, dir = "raw_data/west-yorkshire") {
     tmp = file.path(dir, basename(u))
     if (!file.exists(tmp)) {
         utils::download.file(u, tmp, mode = "wb")
@@ -21,7 +21,8 @@ download_and_read = function(u, remove_nas = TRUE, dir = "raw_data/west-yorkshir
     names(d) = clean_names(names(d))
     d = select_columns(d)
     if (remove_nas) {
-        d = filter_nas(d)
+        d = filter_offence_nas(d)
+        d = filter_location_nas(d)
     }
     return(d)
 }
@@ -42,6 +43,7 @@ select_columns = function(d) {
   d = d[, !names_remove]
 }
 
+
 filter_offence_nas = function(d) {
   d |>
     # Filter out "n/a" values (case insensitive):
@@ -51,7 +53,8 @@ filter_offence_nas = function(d) {
 
 filter_location_nas = function(d) {
   d |>
-    dplyr::filter(!location == "N/A") |>
+    # Location is not "N/A" or NA (case insensitive):
+    dplyr::filter(!grepl("n/a", location, ignore.case = TRUE)) |>
     dplyr::filter(!is.na(location))
 }
 

### 2024-07-05 paper.qmd
slight table updates w gt()

@@
 library(tidyverse)
+library(gt)
+library(gtExtras)
 remotes::install_github("ITSLeeds/opsnap")
@@
   mutate(`Percent of records` = round(`Number of records` / sum(`Number of records`) * 100, 1))
+
+# d_offence_classified |>
+#   knitr::kable()
+
 d_offence_classified |>
-  knitr::kable()
+  mutate(
+    `Percent of records`= round(`Percent of records`)
+    ) |> 
+  rename(`# records` = `Number of records`, `% records`=`Percent of records`) |> 
+  gt() |>
+
+  gt_plt_bar_pct(column = `% records`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+  text_transform(
+    fn <- function(x){
+      code <- str_extract(x, "^[^\ ]+")
+      desc <- str_remove(x, "^[^\ ]+")
+      glue::glue("<em><span style='font-size:14px'>{code}</span></em><br><span style='font-size:18px'>{desc}</span>")
+      },
+    locations=cells_body(columns=Offence)
+  ) |> 
+  cols_width(1 ~ px(600), 2 ~ px(80)) |> 
+  gt_theme_espn()
 ```
@@
 #| tbl-cap: Number of OpSnap records by observer mode.
+# d_all |>
+#   count(mode, sort = TRUE) |>
+#   mutate(percent_of_records = n / nrow(d_all)) |>
+#   mutate(percent_of_records = round(percent_of_records, 3) * 100) |>
+#   arrange(desc(n)) |>
+#   rename_all(snakecase::to_title_case) |>
+#   # Rename N to "Number of records"
+#   rename(`Number of Records` = N) |>
+#   knitr::kable()
+
 d_all |>
@@
   # Rename N to "Number of records"
-  rename(`Number of Records` = N) |>
-  knitr::kable()
+  rename(`# records` = N, `% records` = `Percent of Records`) |>
+  gt() |>
+  gt_plt_bar_pct(column = `% records`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
+  tab_style(
+    style = cell_text(size=px(18)),
+    locations = cells_body(columns = Mode)
+  ) |>
+  gt_theme_espn()
 ```
@@
 #| tbl-cap: Mode of transport from which video was recorded (columns) and offence type (rows), with counts and percentages within each reporting mode.
+# pivot_combined |>
+#   knitr::kable(digits = 1)
 pivot_combined |>
-  knitr::kable(digits = 1)
-
+  select(-`Total (%)`) |> 
+  filter(offence_simple!="Total") |> 
+  select(-c(Cyclist:Other)) |> 
+  pivot_longer(cols=`Cyclist (%)`:`Other (%)`, names_to="mode", values_to="perc") |> 
+  mutate(
+    mode=factor(str_extract(mode, "^[^\ ]+"), levels=c("Driver", "Cyclist", "Other")),
+    offence_simple=factor(offence_simple, levels=c("Inconsiderate driving", "Careless driving", "Other"))
+    ) |> 
+  group_by(offence_simple) |> 
+  summarise(perc=list(round(perc)), count=first(Total)) |> 
+  gt() |>
+  gt_plt_bar_stack(perc, width=65, labels = c(" Cyclists ", " Drivers ", " Other "),
+                   palette= c("#e31a1c", "#1f78b4", "#bdbdbd")) |>
+    cols_width(1 ~ px(180), 2 ~ px(400), 3 ~ px(150)) |> 
+  tab_style(
+    style = cell_text(size=px(18)),
+    locations = cells_body(columns = c(offence_simple, count))
+    ) |> 
+    gt_theme_espn() 
 ```
@@
 #| tbl-cap: "Most common disposal values in the OpSnap dataset."
+# d_all |>
+#   count(disposal, sort = TRUE) |>
+#   mutate(percent_of_records = round(n / nrow(d_all), 3) * 100) |>
+#   arrange(desc(n)) |>
+#   rename_all(snakecase::to_title_case) |> 
+#   knitr::kable()
+
 d_all |>
@@
   rename_all(snakecase::to_title_case) |>
-  knitr::kable()
+  # Rename N to "Number of records"
+  rename(`# records` = N, `% records` = `Percent of Records`) |>
+  gt() |>
+  gt_plt_bar_pct(column = `% records`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
+  tab_style(
+    style = cell_text(size=px(18)),
+    locations = cells_body(columns = Disposal)
+  ) |>
+  gt_theme_espn()
 ```
@@
 #| tbl-cap: "Most common locations recorded in the OpSnap dataset"
+# d_with_location |>
+#   count(location, sort = TRUE) |>
+#   mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
+#   arrange(desc(n)) |>
+#   head(10) |>
+#   rename_all(snakecase::to_title_case) |> 
+#   knitr::kable()
 d_with_location |>
@@
   rename_all(snakecase::to_title_case) |>
-  knitr::kable()
+  rename(`# records` = N, `% records` = `Percent of Records`) |>
+  gt() |>
+  gt_plt_bar_pct(column = `% records`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
+  tab_style(
+    style = cell_text(size=px(18)),
+    locations = cells_body(columns = Location)
+  ) |>
+  gt_theme_espn()
 ```

### 2024-07-12 paper.qmd
Try with PDF

@@
 format:
-  html: default
-  # pdf: default
+  # html: default
+  pdf: default
   # arxiv-pdf:
@@
 execute: 
-  cache: true
   echo: false
@@
     ) |> 
-  rename(`# records` = `Number of records`, `% records`=`Percent of records`) |> 
+  rename(`Number` = `Number of records`, `Percent`=`Percent of records`) |> 
   gt() |>
 
-  gt_plt_bar_pct(column = `% records`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
   text_transform(
@@
   # Rename N to "Number of records"
-  rename(`# records` = N, `% records` = `Percent of Records`) |>
+  rename(`Number` = N, `Percent` = `Percent of Records`) |>
   gt() |>
-  gt_plt_bar_pct(column = `% records`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
   cols_width(1 ~ px(400), 2 ~ px(80)) |> 
@@
   rename_all(snakecase::to_title_case) |>
-  rename(`# records` = N, `% records` = `Percent of Records`) |>
+  rename(`Number` = N, `Percent` = `Percent of Records`) |>
   gt() |>
-  gt_plt_bar_pct(column = `% records`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
   cols_width(1 ~ px(400), 2 ~ px(80)) |> 

### 2024-07-12 paper.qmd
Update with PDF compatible versions of tables, close #6

@@
 
-d_offence_classified |>
+tbl = d_offence_classified |>
   mutate(
@@
   gt_theme_espn()
+gt::gtsave(tbl, "tbl-offences.png")
+knitr::include_graphics("tbl-offences.png")
 ```
@@
 
-d_all |>
+tbl = d_all |>
   count(mode, sort = TRUE) |>
@@
   gt_theme_espn()
+gt::gtsave(tbl, "tbl-mode.png")
+knitr::include_graphics("tbl-mode.png")
 ```
@@
 #   knitr::kable(digits = 1)
-pivot_combined |>
+tbl = pivot_combined |>
   select(-`Total (%)`) |> 
@@
     gt_theme_espn() 
+gt::gtsave(tbl, "tbl-mode-offences-crosstab.png")
+knitr::include_graphics("tbl-mode-offences-crosstab.png")
 ```
@@
 
-d_all |>
+tbl = d_all |>
   count(disposal, sort = TRUE) |>
@@
   gt_theme_espn()
+gt::gtsave(tbl, "tbl-disposal.png")
+knitr::include_graphics("tbl-disposal.png")
 ```
@@
 #   knitr::kable()
-d_with_location |>
+tbl = d_with_location |>
   count(location, sort = TRUE) |>
@@
   gt_theme_espn()
+gt::gtsave(tbl, "tbl-locations.png")
+knitr::include_graphics("tbl-locations.png")
 ```

### 2024-07-12 paper.qmd
Assorted changes

@@
   # html: default
-  pdf: default
-  # arxiv-pdf:
-  #   keep-tex: true
+  # docx: default
+  # pdf: default
+  arxiv-pdf:
+    keep-tex: true
 linenumbers: true # Add (continuous) line numbers?
@@
   warning: false
+  cache: refresh
 ---
@@
 ```
-
-# Abstract {.unnumbered}
-
-This study uses data from Operation Snap (OpSnap), the UK police's national system to receive road users' video evidence of road traffic offences.
-Data from one police force area for the three calendar years 2021 to 2023 (N= 18,363 records), is analysed.
-Of submitted cases, 34.4% were from cyclists (n=6,312) of which 69% resulted in an offence being registered, mostly for driving 'without reasonable consideration to others' or 'without due care and attention'.
-Three quarters (75.2%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), a quarter in no further action, and fewer than 1% for court appearance.
-A research agenda using OpSnap data is outlined: comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional, and local levels; research into disposal-related decision-making, video quality, and the role of supporting evidence should identify good practice; further study of video content, offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases, holds significant potential to inform policy and practice locally, nationally and internationally.
-
-# Introduction
-
-Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
-Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@departmentfortransportReportedRoadCasualties2023].
-Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics [@departmentfortransportReportedRoadCasualties2023].
-Between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year in Britain according to police records [@departmentfortransport].
-Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
-The 2023 report noted that "the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was 'driver or rider failed to look properly'" [@departmentfortransport].
-
-```{=html}
-<!-- The context for the study is that, between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year according to official records recorded by police forces across Great Britain [@departmentfortransport].
-Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
-The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” [@departmentfortransport].  -->
-```
-Cyclist crashes are underreported in police recorded datasets [@elvikIncompleteAccidentReporting1999] and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted [@janstrupetal.UnderstandingTrafficCrash2016].
-Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions [@ibrahimCyclingMissesReview2021].
-These incidents are often referred to as "near-misses" [@ibrahimCyclingMissesReview2021]
-
-Commuter cyclists in the UK experience a near miss for every six miles of riding [@aldredInvestigatingRatesImpacts2015a], and concern over near misses are a key reason people choose not to cycle [@sandersPerceivedTrafficRisk2015a], being associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers [@sandersPerceivedTrafficRisk2015a].
-In their review of near miss cycling crashes, @ibrahimCyclingMissesReview2021 highlighted the need for better data on near misses to inform safety research.
-
-Close passes are the most common type of near miss reported by cyclists, and are associated with collisions resulting in injury [@aldred2016].
-A 'close pass' refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
-Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
-There is no specific law in the UK Road Traffic Act 1988 for driving too close to cyclist, but two are commonly applied for careless driving: RT88575- Drive without due care and attention; and RT99576 -- Driver without reasonable consideration to others - and these are prominent in the analysis that follows.
-
-This study highlights the potential of using open access data from Operation Snap, the UK police's national system to receive road users' video evidence of road traffic offences.
-Operation Snap, often referred to informally as 'OpSnap', was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 20181.
-It is now in operation nationally across England and Wales, each police force offering its own submission portal for road users to submit video evidence.
-The nature of video submissions and the related expectations are summarised on the website of one Police and Crime Commissioner as follows:
-
--   The secure form is for traffic offences, it is NOT for submitting footage of road traffic collisions, any other offences or for parking issues. 
--   The car registration number of the offending vehicle must be clearly visible. 
--   The public should be prepared to sign a witness statement and possibly give evidence in court. 
--   Statements for OpSnap can only be accepted from persons aged 18 or over. If you are under 18 the incident should be reported by email to \[email\]2.
-
-This is, to our knowledge, the first research study to use this dataset.
-As such, the study is offered as proof-of-concept of the potential for analysis of this data, and of its further potential, including understanding the factors associated with cyclist near miss incidents.
-Following analysis and discussion of three years of data for one police force area, the study outlines a research agenda designed to inform policy and practice.
-
-# Methods and data {#sec-methods}
-
-Open access data from West Yorkshire Safer Roads -- OpSnap was used for this study, the media submissions portal having opened in July 2020.
-The West Yorkshire Police (WYP) data used here spanned calendar years 2021, 2022 and 2023.
-For the year 2021 there was less than half the cases of either 2022 or 2023, which could reflect reduced road use during the COVID-19 pandemic and fewer people being aware of OpSnap when it commenced.
-
-The dataset is a tabulation of cases submitted to the West Yorkshire Safer Roads OpSnap web portal.
-The terms 'record' and 'case' are used interchangeably here to refer to a record in the OpSnap database, each of which represents the separate submission of video evidence by a road user.
-
-The portal allows members of the public to submit video footage of suspected traffic offences committed by motor vehicle drivers.
-Video footage is commonly recorded from on-board cameras.
-For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as 'dash cams'.
-One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam [@DashCamSubmissions2024].
-For cyclists, footage is commonly recorded using helmet or handlebar-mounted cameras.
-The proportion of cyclists using these cameras is unknown but anecdotal evidence suggests increased usage, with many choosing to record rides in case an incident occurs.
-
-Complainants upload footage and complete a short form that includes their personal details, the details of the vehicle involved including registration, make, model and colour, the location and time of the incident, and details of the camera used to record the footage.
-Only vehicle offences can be reported as the registration number of any offending vehicle is required and must be legible in footage.
-The open access data is a deidentified summary of submitted cases with informtion on: mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
-The offence location is typically a street name and town or city name, or an intersection, and examples included: 'A58 Godley Road, Halifax', 'Keighley Road, Silsden' , 'Woodhouse Lane A660, Leeds'.
-For this study, approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
-Further aspects of the data, their uses and limitations are discussed in what follows.
 
@@
 
-The data is open acess and looks like this, with names cleaned up by the package:
+<!-- The data is open acess and looks like this, with names cleaned up by the package: -->
 
@@
 ```
-
-# Results
 
@@
   mutate(records = "stats19")
+
+perc_vehicle_driver = d_all |>
+  count(mode) |>
+  filter(mode == "vehicle driver") |>
+  pull(n) / nrow(d_all) * 100 
+perc_cyclist = d_all |>
+  count(mode) |>
+  filter(mode == "cyclist") |>
+  pull(n) / nrow(d_all) * 100 
+perc_pedestrian = d_all |>
+  count(mode) |>
+  filter(mode == "pedestrian") |>
+  pull(n) / nrow(d_all) * 100 
+perc_disposal_educational_course_conditional = d_all |>
+  mutate(disposal = case_when(
+    disposal == "conditional offer" ~ "educational course",
+    TRUE ~ disposal
+  )) |>
+  count(disposal) |>
+  filter(disposal == "educational course") |>
+  pull(n) / nrow(d_all) * 100 
+perc_no_further_action = d_all |>
+  count(disposal) |>
+  filter(disposal == "no further action") |>
+  pull(n) / nrow(d_all) * 100
+perc_course = d_all |>
+  count(disposal) |>
+  filter(disposal == "conditional offer") |>
+  pull(n) / nrow(d_all) * 100
 ```
+
+# Abstract {.unnumbered}
+
+This study uses data from Operation Snap (OpSnap), the UK police's national system to receive road users' video evidence of road traffic offences.
+Data from one police force area for the three calendar years 2021 to 2023
+<!-- (N= 18,363 records) -->
+(N = `r nrow(d_all)` records)
+is analysed.
+<!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians. -->
+Of submitted cases, `r round(perc_vehicle_driver, 1)`% were from vehicle drivers, `r round(perc_cyclist, 1)`% were from cyclists, and `r round(perc_pedestrian, 1)`% were from pedestrians.
+The most common offences were driving 'without reasonable consideration to others' or 'without due care and attention'.
+<!-- Two thirds (66.1%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), 31% no further action, and fewer than 1% for court appearance.
+-->
+Two-thirds (`r round(perc_disposal_educational_course_conditional, 1)`%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), `r round(perc_no_further_action, 1)`% no further action, and `r round(perc_course, 1)`% for court appearance.
+A research agenda using OpSnap data is outlined: comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional, and local levels; research into disposal-related decision-making, video quality, and the role of supporting evidence should identify good practice; further study of video content, offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases, holds significant potential to inform policy and practice locally, nationally and internationally.
+
+# Introduction
+
+Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
+Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@departmentfortransportReportedRoadCasualties2023].
+
+Traditionally, road safety research has relied on retrospective analysis of crashes to assess and enhance road safety [@towards2008].
+However, relying on historical crash data poses several challenges.
+Crashes are rare events [@elvik2006], requiring extended periods, often at least five years, of data to obtain statistically significant estimates [@songchitruksa2006].
+Furthermore, the reactive nature of crash analysis means that safety improvements only follow after crashes occur, which is both inefficient and ethically problematic.
+
+Surrogate safety measures, offer a proactive alternative to identify
+road safety issues [@songchitruksa2006].
+These measures do not depend on crash data but instead use more frequently observable, less severe, traffic events to identify road
+safety issues (reference).
+One prominent surrogate measure is traffic
+conflicts (reference).
+A traffic conflict occurs when road users’
+paths intersect, with a risk of collision if no action was taken
+(reference).
+The concept of using traffic conflicts as a surrogate
+measure was notably illustrated by Hyden’s safety pyramid
+(reference).
+This pyramid represents the spectrum of traffic events
+from safe, undisturbed passages at the base, to severe, rare crashes
+at the top.
+Hyden’s model demonstrates that the inverse
+relationship between crash frequency and severity.
+Understanding this
+relationship allows for the prediction of severe crashes based on the
+more common, less sever conflicts.
+
+Traditional methods for observing traffic conflicts have relied on
+manual observation, which is resource intensive.
+This study
+highlights the potential of using open access data from Operation
+Snap, the UK police’s national system to receive road users’
+video evidence of road traffic offences as a surrogate measure of
+road safety.
+
+<!-- Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics [@departmentfortransportReportedRoadCasualties2023].
+Between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year in Britain according to police records [@departmentfortransport].
+Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
+The 2023 report noted that "the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was 'driver or rider failed to look properly'" [@departmentfortransport].
+-->
+
+<!-- The context for the study is that, between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year according to official records recorded by police forces across Great Britain [@departmentfortransport].
+Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
+The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” [@departmentfortransport].  -->
+
+Cyclist crashes are underreported in police recorded datasets [@elvikIncompleteAccidentReporting1999] and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted [@janstrupetal.UnderstandingTrafficCrash2016].
+Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions [@ibrahimCyclingMissesReview2021].
+These incidents are often referred to as "near-misses" [@ibrahimCyclingMissesReview2021]
+
+Commuter cyclists in the UK experience a near miss for every six miles of riding [@aldredInvestigatingRatesImpacts2015a], and concern over near misses are a key reason people choose not to cycle [@sandersPerceivedTrafficRisk2015a], being associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers [@sandersPerceivedTrafficRisk2015a].
+In their review of near miss cycling crashes, @ibrahimCyclingMissesReview2021 highlighted the need for better data on near misses to inform safety research.
+
+Close passes are the most common type of near miss reported by cyclists, and are associated with collisions resulting in injury [@aldred2016].
+A 'close pass' refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
+Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
+There is no specific law in the UK Road Traffic Act 1988 for driving too close to cyclist, but two are commonly applied for careless driving: RT88575- Drive without due care and attention; and RT99576 -- Driver without reasonable consideration to others - and these are prominent in the analysis that follows.
+
+Operation Snap, often referred to informally as 'OpSnap', was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 2018.
+It is now in operation nationally across England and Wales, each police force offering its own submission portal for road users to submit video evidence.
+The nature of video submissions and the related expectations are summarised on the website of one Police and Crime Commissioner as follows:
+
+-   The secure form is for traffic offences, it is NOT for submitting footage of road traffic collisions, any other offences or for parking issues. 
+-   The car registration number of the offending vehicle must be clearly visible. 
+-   The public should be prepared to sign a witness statement and possibly give evidence in court. 
+-   Statements for OpSnap can only be accepted from persons aged 18 or over. If you are under 18 the incident should be reported by email.
+
+This is, to our knowledge, the first research study to use this dataset.
+As such, the study is offered as proof-of-concept of the potential for analysis of this data, and of its further potential, including understanding the factors associated with cyclist near miss incidents.
+Following analysis and discussion of three years of data for one police force area, the study outlines a research agenda designed to inform policy and practice.
+
+# Methods and data {#sec-methods}
+
+Open access data from West Yorkshire Safer Roads -- OpSnap was used for this study, the media submissions portal having opened in July 2020.
+The West Yorkshire Police (WYP) data used here spanned calendar years 2021, 2022 and 2023.
+For the year 2021 there was less than half the cases of either 2022 or 2023, which could reflect reduced road use during the COVID-19 pandemic and fewer people being aware of OpSnap when it commenced.
+
+The dataset is a tabulation of cases submitted to the West Yorkshire Safer Roads OpSnap web portal.
+The terms 'record' and 'case' are used interchangeably here to refer to a record in the OpSnap database, each of which represents the separate submission of video evidence by a road user.
+
+The portal allows members of the public to submit video footage of suspected traffic offences committed by motor vehicle drivers.
+Video footage is commonly recorded from on-board cameras.
+For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as 'dash cams'.
+One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam [@DashCamSubmissions2024].
+For cyclists, footage is commonly recorded using helmet or handlebar-mounted cameras.
+The proportion of cyclists using these cameras is unknown but anecdotal evidence suggests increased usage, with many choosing to record rides in case an incident occurs.
+
+Complainants upload footage and complete a short form that includes their personal details, the details of the vehicle involved including registration, make, model and colour, the location and time of the incident, and details of the camera used to record the footage.
+Only vehicle offences can be reported as the registration number of any offending vehicle is required and must be legible in footage.
+The open access data is a deidentified summary of submitted cases with informtion on: mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
+The offence location is typically a street name and town or city name, or an intersection, and examples included: 'A58 Godley Road, Halifax', 'Keighley Road, Silsden' , 'Woodhouse Lane A660, Leeds'.
+For this study, approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
+Further aspects of the data, their uses and limitations are discussed in what follows.
+
+# Results
 

### 2024-07-12 paper.qmd
Add table 1

@@
   warning: false
-  cache: refresh
+  cache: true
 ---
@@
 
-Open access data from West Yorkshire Safer Roads -- OpSnap was used for this study, the media submissions portal having opened in July 2020.
-The West Yorkshire Police (WYP) data used here spanned calendar years 2021, 2022 and 2023.
+Open access data from West Yorkshire Safer Roads from the OpSnap project was used for this study.
+The media submissions portal opened in July 2020, and available data from the West Yorkshire Police (WYP) used in this paper span the calendar years 2021, 2022 and 2023.
 For the year 2021 there was less than half the cases of either 2022 or 2023, which could reflect reduced road use during the COVID-19 pandemic and fewer people being aware of OpSnap when it commenced.
@@
 Only vehicle offences can be reported as the registration number of any offending vehicle is required and must be legible in footage.
-The open access data is a deidentified summary of submitted cases with informtion on: mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
-The offence location is typically a street name and town or city name, or an intersection, and examples included: 'A58 Godley Road, Halifax', 'Keighley Road, Silsden' , 'Woodhouse Lane A660, Leeds'.
+The open access data is a deidentified summary of submitted cases with information on: mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
+The offence location is typically a street name and town or city name, or an intersection, and examples included: 'A58 Godley Road, Halifax', 'Keighley Road, Silsden', 'Woodhouse Lane A660, Leeds'.
 For this study, approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
 Further aspects of the data, their uses and limitations are discussed in what follows.
+
+A random sample of 5 records from the raw data is shown in @tbl-raw (note: "nfa" refers to "no further action").
+
+```{r}
+#| label: tbl-raw
+#| tbl-cap: "Random sample of 5 records from the raw data."
+set.seed(2024)
+d_all |>
+  sample_n(5) |>
+  select(-district, -`offence_code`) |>
+  mutate(date = as.Date(date)) |>
+  arrange(date) |>
+  rename_all(snakecase::to_title_case) |>
+  knitr::kable() |>
+  # Striped styling and tiny text to fit:
+  kableExtra::kable_styling(latex_options = "striped") |>
+  # Set column widths:
+  kableExtra::column_spec(1, width = "3em") |>
+  kableExtra::column_spec(2, width = "3em") |>
+  kableExtra::column_spec(3, width = "4em") |>
+  kableExtra::column_spec(4, width = "4em") |>
+  kableExtra::column_spec(5, width = "6em") |>
+  kableExtra::column_spec(6, width = "4em") |>
+  kableExtra::column_spec(8, width = "10em")
+
+```
 

### 2024-07-12 paper.qmd
Final tweaks to paper

@@
   warning: false
-  cache: true
+  cache: false
 ---
@@
 <!-- (N= 18,363 records) -->
-(N = `r nrow(d_all)` records)
+(N = `r nrow(d_all) |> scales::number(big.mark = ",")` records)
 is analysed.
@@
 
+There were `r nrow(d_all) |> scales::number(big.mark = ",")` records in the dataset for the three-year study period, with a strong upward trend, as shown in the monthly counts presented in @fig-time.
+
 ```{r}
@@
 
-There were `r nrow(d_all)` records in the dataset for the three-year study period, with a strong upward trend, as shown in the monthly counts presented in @fig-time.
-Some records lacked either an offence (`r nrow(d_all) - nrow(d_offence)`, `r round((nrow(d_all) - nrow(d_offence)) / nrow(d_all) * 100, 1)`%) or a location (`r nrow(d_offence) - nrow(d_complete)`, `r round((nrow(d_offence) - nrow(d_complete)) / nrow(d_all) * 100, 1)`%), or both, leaving `r round(nrow(d_complete) / nrow(d_all) * 100, 1)`% or `r nrow(d_complete)` complete records.
+Some records lacked either an offence (`r format(nrow(d_all) - nrow(d_offence), big.mark = ",", scientific = FALSE)`, `r round((nrow(d_all) - nrow(d_offence)) / nrow(d_all) * 100, 1)`%) or a location (`r format(nrow(d_offence) - nrow(d_complete), big.mark = ",", scientific = FALSE)`, `r round((nrow(d_offence) - nrow(d_complete)) / nrow(d_all) * 100, 1)`%), or both, leaving `r round(nrow(d_complete) / nrow(d_all) * 100, 1)`% or `r format(nrow(d_complete), big.mark = ",", scientific = FALSE)` complete records.
 There was a distinct seasonal pattern to reporting, with significant increases in summer months.
 
-Of the `r nrow(d_offence)` records with an offence, the most common offences were Driving without reasonable consideration to others (rt88576) and Driving without due care and attention (rt88575).
-This is presented in @tbl-offences, which shows the number and percentage of records by offence type, showing the top 6 offence types and the remainder grouped as 'Other'.
-As previously noted, a high proportion of cases did not include information on the reported offence, with the most common offences being Driving without reasonable consideration to others (rt88576), and Driving without due care and attention (rt88575).
-Within the road traffic act these offences are related to careless driving and drivers are subject to similar penalties.
+A summary of the `r nrow(d_offence) |> scales::number(big.mark = ",")` records with an offence is presented in @tbl-offences.
+Included in the table are the number and percentage of records by offence type, showing the top 6 offence types and the remainder grouped as ‘Other’.
+
+The most common offences were ‘Driving without reasonable consideration to others (rt88576)’ and ‘Driving without due care and attention (rt88575)’.
+Within the Road Traffic Act these offences are related to careless driving and drivers are subject to similar penalties.
+The other common offences included failing to comply with traffic signals, traffic signs and the use of a handheld phone while driving. 
+<!-- TODO: Within the other category... -->
 
@@
 Vehicle (mostly car and van drivers with dashcams) driver and cyclist reporters dominate reporting for all records, as illustrated in @tbl-mode.
-Half of cases were reported by vehicle drivers, a third by cyclists, seven percent by pedestrians, with over two percent by horse riders and less than one percent by motorcyclists.
+Half of the cases were reported by vehicle drivers, a third by cyclists, seven percent by pedestrians, with over two percent by horse riders and less than one percent by motorcyclists.
 
@@
 tbl = d_all |>
+  # NA is Unknown:
+  mutate(mode = ifelse(is.na(mode), "Unknown", mode)) |>
   count(mode, sort = TRUE) |>
@@
 
-The cross-tabulation of reporting mode and offence type is shown in @tbl-mode-offences-crosstab.
-The table reveals some interesting tendencies.
-While `r round(percent_inconsiderate, 1)` percent of all offences are for inconsiderate driving, they make up the bulk of offences reported by cyclists, with `r round(percent_inconsiderate_cycling, 1)` percent of all offences reported by cyclists being for inconsiderate driving.
+The cross-tabulation of reporting mode and offence type is shown in Table @tbl-mode-offences-crosstab, revealing some interesting tendencies.
+
+While `r round(percent_inconsiderate, 1)` percent of all offences are for driving without reasonable considered to others (rt88576), they make up the bulk of offences reported by cyclists (`r round(percent_inconsiderate_cycling, 1)`%).
 Drivers are proportionally `r round(percent_other_driver / percent_other_cyclist, 1)` times more likely to report other types of offences as cyclists, while other reporting modes are most likely to report other types of offences, being `r round(percent_other_other / percent_other_cyclist, 1)` times more likely to report other types of offences as cyclists.
-While further research is needed, this makes sense if physically-vulnerable cyclists are most concerned with the dangerous driving of vehicles, whereas drivers also report other types of road traffic offence, with pedestrians and others also more likely to encounter other types of offence.
+While further research is needed to understand the reasons for these tendencies, the results match intuition.
+Physically-vulnerable cyclists are understandably most concerned with the dangerous driving of vehicles, whereas drivers tend to focus on other types of road traffic offence.
+Pedestrians and other reporting modes were also relatively more likely to encounter other types of offence.
 
@@
   knitr::kable()
+
+perc_no_further_action = d_all |>
+  count(disposal) |>
+  filter(disposal == "nfa") |>
+  pull(n) / nrow(d_all) * 100
+
+perc_course = d_all |>
+  count(disposal) |>
+  filter(disposal == "course") |>
+  pull(n) / nrow(d_all) * 100
 ```
@@
 Disposal categories assigned by police are shown in @tbl-disposal.
-Roughly a third of cases resulted in no further action, and for most of the remainder drivers were required to undertake an education course.
-Conditional offers, that is, drivers being offered a reduced penalty for admitting guilt, were the third most common outcome (XX EXPLAIN FURTHER?).
+Roughly a third of cases (`r round(perc_no_further_action, 1)`%) resulted in no further action, and for most of the remainder drivers were required to undertake an education course (`r round(perc_course, 1)`%).
+Conditional offers, that is, drivers being offered a reduced penalty for admitting guilt, were the third most common outcome.
 Close to two percent of cases went to court and a further one percent underwent further investigation.
@@
 
-Cyclists undertake two percent of road miles in the UK [@departmentfortransport].
-Other things equal, this suggests that with respect to distance travelled, cyclists were around 20 times more likely to submit evidence of a road traffic offence than vehicle drivers and other road users in West Yorkshire.
-
-Over 90 percent of offences reported by cyclists were for driving without consideration for other road users and without due care and attention.
-This is consistent with the phenomenon of 'close passes' in the cycling safety literature [@aldred2016; @cubbin2024].
-It suggests the OpSnap data holds the potential for further analysis to inform knowledge about the nature of close passes generally.
-For example, it should be possible to identify hotspot locations.
-As with much of the further research outlined here, this the potential to inform preventive policy and practice relating to police interventions, driver behaviour, the design of roads and roadside environments.
-
 There was an increase in submissions over time between 2021 and 2023.
@@
 
+This was a case study of West Yorkshire but each of the 43 police forces in England and Wales collates OpSnap data, and our preliminary scoping of individual force websites suggests OpSnap data is publicly available nationwide.
+This suggests the potential for development of a national OpSnap dataset.
+A feasibility study is needed to establish the availability and compatibility of the data from different forces.
+Should a national dataset prove feasible it holds potential for national-level, cross-regional, and comparative analyses of patterns and trends.
+A national dataset will enhance the possibility for cross-national comparative analysis and international cooperation in road safety.
+It holds the potential for the development of rankings according to different criteria and, thereby, potential performance indicators.
+
 If submission volumes continue to increase, means of informing triage and prioritisation of cases will be increasingly valuable.
@@
 It may provide feasible, perhaps using machine learning, to develop automated means of triage, determination of offence type and disposal.
-Research to gauge the differential effectiveness of disposal measures is needed, starting with measures of recividism.
+Research to gauge the differential effectiveness of disposal measures is needed, starting with measures of recidivism.
 
@@
 
-Evidence submitted by cyclists was the focus here.
-However, Operation Snap data offer potential for analysis of data from other road user sources.
 Studies focused on specific types of road user will prove informative with respect to policy and practice.
@@
 The British Horse Society[^1] and others may be interested in this data being used to inform the safety of horses and riders.
-The British Motorcyclists Federation[^2] may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent
-.
+The British Motorcyclists Federation[^2] may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent.
 
@@
 [^2]: https://www.britishmotorcyclists.co.uk/
+
+As an example, cyclists undertake two percent of road miles in the UK [@departmentfortransport].
+Other things equal, this suggests that with respect to distance travelled, cyclists were around 20 times more likely to submit evidence of a road traffic offence than vehicle drivers and other road users in West Yorkshire.
+
+Over 90 percent of offences reported by cyclists were for driving without consideration for other road users and without due care and attention.
+This is consistent with the phenomenon of 'close passes' in the cycling safety literature [@aldred2016; @cubbin2024].
+It suggests the OpSnap data holds the potential for further analysis to inform knowledge about the nature of close passes generally.
+For example, it should be possible to identify hotspot locations.
+As with much of the further research outlined here, this the potential to inform preventive policy and practice relating to police interventions, driver behaviour, the design of roads and roadside environments.
 
@@
 The spatial analysis undertaken here was obliged to use data for 'all' video submissions.
-We were unable to separate the submissions by cyclists because, in the publicly-available data, locations could not be matched to individual cases: they were scrambled to ensure anonymity for GDPR purposes.
+We were unable to cross-tabulate the geographic location of submissions by different road users because, in the publicly-available data, locations are not be matched to individual cases: they were scrambled to ensure anonymity for GDPR purposes.
 An obvious next step, in the context of a secure research platform, would be spatial analysis for different types of road users, for different types of incidents, for incidents resulting in different disposals, and so on.

### 2024-07-12 paper.qmd
Docx version

@@
   # html: default
-  # docx: default
+  docx: default
   # pdf: default
-  arxiv-pdf:
-    keep-tex: true
+  # arxiv-pdf:
+  #   keep-tex: true
 linenumbers: true # Add (continuous) line numbers?
@@
 quarto::quarto_render("paper.qmd", output_format = "docx")
-file.rename("paper.docx", "~/OneDrive/opsnap_leeds/paper-v6.docx")
-file.rename("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v6.pdf")
+file.rename("paper.docx", "~/OneDrive/opsnap_leeds/paper-v7.docx")
+file.rename("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v7.pdf")
 browseURL("paper.docx")
@@
   rename_all(snakecase::to_title_case) |>
-  knitr::kable() |>
-  # Striped styling and tiny text to fit:
-  kableExtra::kable_styling(latex_options = "striped") |>
-  # Set column widths:
-  kableExtra::column_spec(1, width = "3em") |>
-  kableExtra::column_spec(2, width = "3em") |>
-  kableExtra::column_spec(3, width = "4em") |>
-  kableExtra::column_spec(4, width = "4em") |>
-  kableExtra::column_spec(5, width = "6em") |>
-  kableExtra::column_spec(6, width = "4em") |>
-  kableExtra::column_spec(8, width = "10em")
+  knitr::kable()
+
+  # |>
+  # # Striped styling and tiny text to fit:
+  # kableExtra::kable_styling(latex_options = "striped") |>
+  # # Set column widths:
+  # kableExtra::column_spec(1, width = "3em") |>
+  # kableExtra::column_spec(2, width = "3em") |>
+  # kableExtra::column_spec(3, width = "4em") |>
+  # kableExtra::column_spec(4, width = "4em") |>
+  # kableExtra::column_spec(5, width = "6em") |>
+  # kableExtra::column_spec(6, width = "4em") |>
+  # kableExtra::column_spec(8, width = "10em")
 
@@
 
-INSERT Table 5 AND/OR Map (TO BE INCLUDED)
-
 ```{r}
@@
   gt_theme_espn()
-gt::gtsave(tbl, "tbl-locations.png")
+# gt::gtsave(tbl, "tbl-locations.png")
 knitr::include_graphics("tbl-locations.png")

### 2024-07-12 paper.qmd
Additional references

@@
   # html: default
-  docx: default
-  # pdf: default
+  # docx: default
+  pdf: default
   # arxiv-pdf:
@@
 This study uses data from Operation Snap (OpSnap), the UK police's national system to receive road users' video evidence of road traffic offences.
-Data from one police force area for the three calendar years 2021 to 2023
-<!-- (N= 18,363 records) -->
-(N = `r nrow(d_all) |> scales::number(big.mark = ",")` records)
-is analysed.
-<!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians. -->
-Of submitted cases, `r round(perc_vehicle_driver, 1)`% were from vehicle drivers, `r round(perc_cyclist, 1)`% were from cyclists, and `r round(perc_pedestrian, 1)`% were from pedestrians.
+Data from one police force area for the three calendar years 2021 to 2023 <!-- (N= 18,363 records) --> (N = `r nrow(d_all) |> scales::number(big.mark = ",")` records) is analysed.
+<!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians.
+--> Of submitted cases, `r round(perc_vehicle_driver, 1)`% were from vehicle drivers, `r round(perc_cyclist, 1)`% were from cyclists, and `r round(perc_pedestrian, 1)`% were from pedestrians.
 The most common offences were driving 'without reasonable consideration to others' or 'without due care and attention'.
 <!-- Two thirds (66.1%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), 31% no further action, and fewer than 1% for court appearance.
--->
-Two-thirds (`r round(perc_disposal_educational_course_conditional, 1)`%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), `r round(perc_no_further_action, 1)`% no further action, and `r round(perc_course, 1)`% for court appearance.
+--> Two-thirds (`r round(perc_disposal_educational_course_conditional, 1)`%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), `r round(perc_no_further_action, 1)`% no further action, and `r round(perc_course, 1)`% for court appearance.
 A research agenda using OpSnap data is outlined: comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional, and local levels; research into disposal-related decision-making, video quality, and the role of supporting evidence should identify good practice; further study of video content, offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases, holds significant potential to inform policy and practice locally, nationally and internationally.
@@
 
-Surrogate safety measures, offer a proactive alternative to identify
-road safety issues [@songchitruksa2006].
-These measures do not depend on crash data but instead use more frequently observable, less severe, traffic events to identify road
-safety issues (reference).
-One prominent surrogate measure is traffic
-conflicts (reference).
-A traffic conflict occurs when road users’
-paths intersect, with a risk of collision if no action was taken
-(reference).
-The concept of using traffic conflicts as a surrogate
-measure was notably illustrated by Hyden’s safety pyramid
-(reference).
-This pyramid represents the spectrum of traffic events
-from safe, undisturbed passages at the base, to severe, rare crashes
-at the top.
-Hyden’s model demonstrates that the inverse
-relationship between crash frequency and severity.
-Understanding this
-relationship allows for the prediction of severe crashes based on the
-more common, less sever conflicts.
+Surrogate safety measures offer an alternative to the use of casualty data to identify safety issues by using more frequently observable, less severe, traffic events to identify road safety issues [@songchitruksa2006].
+One prominent surrogate measure is traffic conflicts [@lord2021].
+A traffic conflict occurs when road users’ paths intersect, with a risk of collision if no action was taken [@tarko2018].
+The concept of using traffic conflicts as a surrogate measure was notably illustrated by Hyden’s safety pyramid [@hyden1987].
+This pyramid represents the spectrum of traffic events from safe, undisturbed passages at the base, to severe, rare crashes at the top.
+Hyden’s model demonstrates that the inverse relationship between crash frequency and severity.
+Understanding this relationship allows for the prediction of severe crashes based on the more common, less sever conflicts.
 
-Traditional methods for observing traffic conflicts have relied on
-manual observation, which is resource intensive.
-This study
-highlights the potential of using open access data from Operation
-Snap, the UK police’s national system to receive road users’
-video evidence of road traffic offences as a surrogate measure of
-road safety.
+Traditional methods for observing traffic conflicts have relied on manual observation, which is resource intensive.
+This study highlights the potential of using open access data from Operation Snap, the UK police’s national system to receive road users’ video evidence of road traffic offences as a surrogate measure of road safety.
 
@@
 -->
-
 <!-- The context for the study is that, between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year according to official records recorded by police forces across Great Britain [@departmentfortransport].
@@
   rename_all(snakecase::to_title_case) |>
-  knitr::kable()
-
-  # |>
-  # # Striped styling and tiny text to fit:
-  # kableExtra::kable_styling(latex_options = "striped") |>
-  # # Set column widths:
-  # kableExtra::column_spec(1, width = "3em") |>
-  # kableExtra::column_spec(2, width = "3em") |>
-  # kableExtra::column_spec(3, width = "4em") |>
-  # kableExtra::column_spec(4, width = "4em") |>
-  # kableExtra::column_spec(5, width = "6em") |>
-  # kableExtra::column_spec(6, width = "4em") |>
-  # kableExtra::column_spec(8, width = "10em")
-
+  knitr::kable() |>
+  # Striped styling and tiny text to fit:
+  kableExtra::kable_styling(latex_options = "striped") |>
+  # Set column widths:
+  kableExtra::column_spec(1, width = "3em") |>
+  kableExtra::column_spec(2, width = "3em") |>
+  kableExtra::column_spec(3, width = "4em") |>
+  kableExtra::column_spec(4, width = "4em") |>
+  kableExtra::column_spec(5, width = "6em") |>
+  kableExtra::column_spec(6, width = "4em") |>
+  kableExtra::column_spec(8, width = "10em")
 ```
@@
 The British Horse Society[^1] and others may be interested in this data being used to inform the safety of horses and riders.
-The British Motorcyclists Federation[^2] may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent.
+The British Motorcyclists Federation[^2] may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent
+.
 

### 2024-07-15 paper.qmd
Add tbl-offences

@@
   gt_theme_espn()
-gt::gtsave(tbl, "tbl-offences.png")
+# gt::gtsave(tbl, "tbl-offences.png")
+# browseURL("tbl-offences.png")
 knitr::include_graphics("tbl-offences.png")

### 2024-07-15 paper.qmd
Add pngs

@@
   gt_theme_espn()
-gt::gtsave(tbl, "tbl-mode.png")
+# gt::gtsave(tbl, "tbl-mode.png")
 knitr::include_graphics("tbl-mode.png")

### 2024-07-15 paper.qmd
Update without gt webshot2 issue

@@
     gt_theme_espn() 
-gt::gtsave(tbl, "tbl-mode-offences-crosstab.png")
+# gt::gtsave(tbl, "tbl-mode-offences-crosstab.png")
 knitr::include_graphics("tbl-mode-offences-crosstab.png")
@@
   gt_theme_espn()
-gt::gtsave(tbl, "tbl-disposal.png")
+# gt::gtsave(tbl, "tbl-disposal.png")
 knitr::include_graphics("tbl-disposal.png")

### 2024-07-15 paper.qmd
Update paper docs

@@
   # docx: default
-  pdf: default
-  # arxiv-pdf:
-  #   keep-tex: true
+  # pdf: default
+  arxiv-pdf:
+    keep-tex: true
 linenumbers: true # Add (continuous) line numbers?
@@
 browseURL("paper.docx")
+# Install extension:
+system("quarto install extension mikemahoney218/quarto-arxiv")
 ```

### 2024-07-15 paper.qmd
New Figure 1

@@
 system("quarto install extension mikemahoney218/quarto-arxiv")
+system("sudo apt install lmodern")
 ```
@@
 dir.create("data/stats19", recursive = TRUE, showWarnings = FALSE)
-collisions = stats19::get_stats19(year = 2022)
+collisions_2022 = stats19::get_stats19(year = 2022)
+collisions_2021 = stats19::get_stats19(year = 2021)
+collisions_2020 = stats19::get_stats19(year = 2020)
+collisions = bind_rows(collisions_2020, collisions_2021, collisions_2022)
 # collisions_2021 = stats19::get_stats19(year = 2021)
@@
 There were `r nrow(d_all) |> scales::number(big.mark = ",")` records in the dataset for the three-year study period, with a strong upward trend, as shown in the monthly counts presented in @fig-time.
+Since early 2022, there have been more monthly records in the OpSnap data than in the official 'STATS19' road traffic collision records for West Yorkshire, highlighting the under-reporting of road traffic incidents in official statistics.
+STATS19 records are from the Department for Transport's database of road traffic collisions reported to the police and only include incidents that result in injury.
+Like OpSnap data, STATS19 records are open access.
+They were downloaded with the `stats19` R package  and filtered to include only records from West Yorkshire to ensure comparability with the OpSnap data.
 
@@
 #| label: fig-time
-#| fig-cap: "Monthly count of Operation Snap records, West Yorkshire"
+#| fig-cap: "Monthly count of Operation Snap (complete and with offence and location data, red and green) and official STATS19 road traffic collision records (blue), West Yorkshire."
 d_monthly = bind_rows(d_monthly, stats19_monthly)
 d_monthly |>
-  mutate(records = snakecase::to_title_case(records)) |> 
+  mutate(records = case_when(
+    records == "all" ~ "OpSnap (all)",
+    records == "complete" ~ "OpSnap (complete)",
+    records == "stats19" ~ "STATS19"
+    )
+  ) |>
   rename_all(snakecase::to_title_case) |> 
@@
   # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
-  labs(title = "Number of records in West Yorkshire Police\nOperation Snap data",
+  labs(
+    # title = "Number of monthly records in West Yorkshire Police\nOperation Snap data",
        x = "Date",

### 2024-07-15 paper.qmd
Update with STATS19 ref

@@
 Like OpSnap data, STATS19 records are open access.
-They were downloaded with the `stats19` R package  and filtered to include only records from West Yorkshire to ensure comparability with the OpSnap data.
+They were downloaded with the `stats19` R package [@lovelace2019] and filtered to include only records from West Yorkshire to ensure comparability with the OpSnap data.
 

### 2024-07-15 paper.qmd
Various tweaks

@@
   pull(n) / nrow(d_all) * 100
+date_range = range(d_all$date)
+# Crashes in 2024:
+d_all_2024 = d_all |>
+  filter(date >= "2024-01-01")
+table(d_all_2024$date)
 ```
@@
 This study uses data from Operation Snap (OpSnap), the UK police's national system to receive road users' video evidence of road traffic offences.
-Data from one police force area for the three calendar years 2021 to 2023 <!-- (N= 18,363 records) --> (N = `r nrow(d_all) |> scales::number(big.mark = ",")` records) is analysed.
+Data from one police force area for the three years (January 2021 to March 2024) <!-- (N= 18,363 records) --> (N = `r nrow(d_all) |> scales::number(big.mark = ",")` records) is analysed.
 <!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians.
@@
 This pyramid represents the spectrum of traffic events from safe, undisturbed passages at the base, to severe, rare crashes at the top.
-Hyden’s model demonstrates that the inverse relationship between crash frequency and severity.
+Hyden’s model demonstrates the inverse relationship between crash frequency and severity.
 Understanding this relationship allows for the prediction of severe crashes based on the more common, less sever conflicts.
@@
 
-Close passes are the most common type of near miss reported by cyclists, and are associated with collisions resulting in injury [@aldred2016].
+Close passes are the most common type of near miss reported by cyclists and are associated with collisions resulting in injury [@aldred2016].
 A 'close pass' refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
@@
 Only vehicle offences can be reported as the registration number of any offending vehicle is required and must be legible in footage.
-The open access data is a deidentified summary of submitted cases with information on: mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
+The open access data is a deidentified summary of submitted cases with information on mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
 The offence location is typically a street name and town or city name, or an intersection, and examples included: 'A58 Godley Road, Halifax', 'Keighley Road, Silsden', 'Woodhouse Lane A660, Leeds'.
@@
 #| tbl-cap: "Random sample of 5 records from the raw data."
-set.seed(2024)
+set.seed(24)
 d_all |>
@@
   kableExtra::column_spec(2, width = "3em") |>
-  kableExtra::column_spec(3, width = "4em") |>
-  kableExtra::column_spec(4, width = "4em") |>
-  kableExtra::column_spec(5, width = "6em") |>
+  kableExtra::column_spec(3, width = "3em") |>
+  kableExtra::column_spec(4, width = "3em") |>
+  kableExtra::column_spec(5, width = "9em") |>
   kableExtra::column_spec(6, width = "4em") |>
-  kableExtra::column_spec(8, width = "10em")
+  kableExtra::column_spec(8, width = "8em")
 ```
@@
 Since early 2022, there have been more monthly records in the OpSnap data than in the official 'STATS19' road traffic collision records for West Yorkshire, highlighting the under-reporting of road traffic incidents in official statistics.
-STATS19 records are from the Department for Transport's database of road traffic collisions reported to the police and only include incidents that result in injury.
+STATS19 record cs are from the Department for Transport's database of road traffic collisions reported to the police and only include incidents that result in injury.
 Like OpSnap data, STATS19 records are open access.
@@
 #| tbl-cap: "Number of convictions by offence type (top 6 and other)."
+#| width: 60%
 # Aim: get table of n. offences by mode
@@
 #| tbl-cap: Number of OpSnap records by observer mode.
+#| width: "80%"
 # d_all |>
@@
   # NA is Unknown:
-  mutate(mode = ifelse(is.na(mode), "Unknown", mode)) |>
+  mutate(mode = case_when(
+    is.na(mode) ~ "unknown",
+    TRUE ~ mode
+  )) |>
   count(mode, sort = TRUE) |>
@@
   mutate(percent_of_records = round(percent_of_records, 3) * 100) |>
+  # Reclassify Unknown to unknown:
   arrange(desc(n)) |>
@@
   gt_theme_espn()
+tbl
 # gt::gtsave(tbl, "tbl-mode.png")

### 2024-07-15 paper.qmd
Re-update table

@@
   gt_theme_espn()
-tbl
+# tbl
 # gt::gtsave(tbl, "tbl-mode.png")
@@
 
-The cross-tabulation of reporting mode and offence type is shown in Table @tbl-mode-offences-crosstab, revealing some interesting tendencies.
+The cross-tabulation of reporting mode and offence type is shown in @tbl-mode-offences-crosstab, revealing some interesting tendencies.
 

### 2024-07-15 paper.qmd
Stats19 tweaks

@@
 #| include: false
-years = 2021:2022
-dir.create("data/stats19", recursive = TRUE, showWarnings = FALSE)
-collisions_2022 = stats19::get_stats19(year = 2022)
-collisions_2021 = stats19::get_stats19(year = 2021)
-collisions_2020 = stats19::get_stats19(year = 2020)
-collisions = bind_rows(collisions_2020, collisions_2021, collisions_2022)
+# years = 2021:2022
+# dir.create("data/stats19", recursive = TRUE, showWarnings = FALSE)
+# collisions_2022 = stats19::get_stats19(year = 2022)
 # collisions_2021 = stats19::get_stats19(year = 2021)
-# collisions = purrr::map_df(years, stats19::get_stats19, type = "collision", data_dir = "data/stats19")
+# collisions_2020 = stats19::get_stats19(year = 2020)
+# collisions = bind_rows(collisions_2020, collisions_2021, collisions_2022)
+# write_csv(collisions, "collisions_2020-2022.csv")
+# piggyback::pb_upload("collisions_2020-2022.csv")
+system("gh release download v1")
+collisions = read_csv("collisions_2020-2022.csv")
 stats19_monthly = collisions |>
@@
 The use of rates will enhance the identification locations with higher risk because based on their volume of traffic: different types of risk need to be gauged.
-Comparison and integration of OpSnap and Stats193 is an obvious starting point.
-To what extent do OpSnap and Crashmap incidents overlap?
-To what extent are they complementary?
+Comparison and integration of OpSnap and STATS19 datasets could provide a logical starting point.
+To what extent are OpSnap and STATS19 records overlapping or complementary?
 To what extent do they identify the same or different types of problems?

### 2024-07-15 paper.qmd
V8

@@
 quarto::quarto_render("paper.qmd", output_format = "docx")
-file.rename("paper.docx", "~/OneDrive/opsnap_leeds/paper-v7.docx")
-file.rename("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v7.pdf")
+file.rename("paper.docx", "~/OneDrive/opsnap_leeds/paper-v8.docx")
+file.rename("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v8.pdf")
 browseURL("paper.docx")

### 2024-07-22 paper.qmd
Update abstract 1

@@
 keywords: 
-  - Cycling
-  - Hotspots 
   - Road safety
@@
 This study uses data from Operation Snap (OpSnap), the UK police's national system to receive road users' video evidence of road traffic offences.
-Data from one police force area for the three years (January 2021 to March 2024) <!-- (N= 18,363 records) --> (N = `r nrow(d_all) |> scales::number(big.mark = ",")` records) is analysed.
+Data from one police force area for 39 months (January 2021 to March 2024) <!-- (N= 18,363 records) --> (N = `r nrow(d_all) |> scales::number(big.mark = ",")` records) is analysed.
 <!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians.

### 2024-07-22 paper.qmd
Update with changes to abstract, in progress

@@
 
+table(d_all$mode)
+
+table(d_all$mode)
+
+          # cyclist       horse rider      motorcyclist        pedestrian 
+          #    7069               456                50              1467 
+          # unknown    vehicle driver vehicle passenger 
+          #     526             10145               650
+
 perc_vehicle_driver = d_all |>
@@
   pull(n) / nrow(d_all) * 100
+perc_horse = d_all |>
+  count(mode) |>
+  filter(mode == "horse rider") |>
+  pull(n) / nrow(d_all) * 100
+perc_motorcyclist = d_all |>
+  count(mode) |>
+  filter(mode == "motorcyclist") |>
+  pull(n) / nrow(d_all) * 100
+perc_other = 100 - perc_vehicle_driver - perc_cyclist - perc_pedestrian - perc_horse - perc_motorcyclist
+
+# table(d_all$disposal)
+#  conditional offer              court dsit investigation educational course 
+#               2563                321                202              10887 
+#               fine                nfa  rpu investigation 
+#                  1               6365                 24
+
 perc_disposal_educational_course_conditional = d_all |>
@@
   filter(disposal == "educational course") |>
+  pull(n) / nrow(d_all) * 100
+# Just educational:
+perc_disposal_educational = d_all |>
+  count(disposal) |>
+  filter(disposal == "educational course") |>
+  pull(n) / nrow(d_all) * 100
+perc_disposal_conditional = d_all |>
+  count(disposal) |>
+  filter(disposal == "conditional offer") |>
+  pull(n) / nrow(d_all) * 100
+perc_disposal_court = d_all |>
+  count(disposal) |>
+  filter(disposal == "court") |>
+  pull(n) / nrow(d_all) * 100
+perc_undergoing_further_investigation = d_all |>
+  count(disposal) |>
+  filter(disposal == "undergoing further investigation") |>
   pull(n) / nrow(d_all) * 100
@@
 Data from one police force area for 39 months (January 2021 to March 2024) <!-- (N= 18,363 records) --> (N = `r nrow(d_all) |> scales::number(big.mark = ",")` records) is analysed.
-<!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians.
---> Of submitted cases, `r round(perc_vehicle_driver, 1)`% were from vehicle drivers, `r round(perc_cyclist, 1)`% were from cyclists, and `r round(perc_pedestrian, 1)`% were from pedestrians.
-The most common offences were driving 'without reasonable consideration to others' or 'without due care and attention'.
+<!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians. -->
+Half were submitted by vehicle drivers (`r round(perc_vehicle_driver, 1)`%), a third by cyclists (`r round(perc_cyclist, 1)`%), `r round(perc_pedestrian, 1)`% by pedestrians, `r round(perc_horse, 1)`% by horse riders,  `r round(perc_motorcyclist, 1)`% by motorcyclists, and `r round(perc_other, 1)`% were unknown.
+We estimate that, relative to road distance travelled, cyclists were 20 times more likely to submit video evidence than vehicle drivers.
+The most common offences overall were driving 'without reasonable consideration to others' or 'without due care and attention'.
 <!-- Two thirds (66.1%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), 31% no further action, and fewer than 1% for court appearance.
---> Two-thirds (`r round(perc_disposal_educational_course_conditional, 1)`%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), `r round(perc_no_further_action, 1)`% no further action, and `r round(perc_course, 1)`% for court appearance.
+-->
+Half (`r round(perc_disposal_educational, 1)`%) of reported cases resulted in the recommended disposal of an educational course, `r round(perc_no_further_action, 1)`% no further action, 
 A research agenda using OpSnap data is outlined: comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional, and local levels; research into disposal-related decision-making, video quality, and the role of supporting evidence should identify good practice; further study of video content, offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases, holds significant potential to inform policy and practice locally, nationally and internationally.

### 2024-07-22 paper.qmd
Update abstract

@@
 -->
-Half (`r round(perc_disposal_educational, 1)`%) of reported cases resulted in the recommended disposal of an educational course, `r round(perc_no_further_action, 1)`% no further action, 
-A research agenda using OpSnap data is outlined: comparative analysis of submission type, source, hotspot locations, and case disposal should be undertaken at national, regional, and local levels; research into disposal-related decision-making, video quality, and the role of supporting evidence should identify good practice; further study of video content, offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases, holds significant potential to inform policy and practice locally, nationally and internationally.
+Half (`r round(perc_disposal_educational, 1)`%) of reported cases resulted in the recommended disposal of an educational course, `r round(perc_no_further_action, 1)`% no further action `r round(perc_disposal_conditional, 1)`% conditional offer, and `r round(perc_disposal_court, 1)`% resulted in court appearance.
+A research agenda using OpSnap data is outlined that could emerge if national datasets are compiled and responsibly opened-up and made available for research and policy-making: data-driven research should identify hotspot locations and other correlates of dangerous and antisocial road use at regional, and local levels; research projects should investigate disposal-related decision-making, video quality, and the role of supporting evidence; offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases should be explored with reference to new video evidence.
+We conclude that datasets derived from publicly-uploaded video submission portals have the potential to transform evidence-based policy and practice locally, nationally and internationally.
 

### 2024-07-22 paper.qmd
Tweak first para of intro

@@
 Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
-Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities [@departmentfortransportReportedRoadCasualties2023].
+There were 1,711 fatalities and 135,480 causalities across all severity categories (slight, serious fatal) in Britain resulting from car crashes and other road traffic collisions reported by the police in 2022  [@departmentfortransportReportedRoadCasualties2023].
+This is just the tip of the iceberg of road traffic incidents, with many more near misses and minor incidents going unreported.
 

### 2024-07-22 paper.qmd
Finish intro tweaks

@@
 
-Surrogate safety measures offer an alternative to the use of casualty data to identify safety issues by using more frequently observable, less severe, traffic events to identify road safety issues [@songchitruksa2006].
+Surrogate safety measures offer an alternative to casualty data by using more frequently observable, less severe, traffic events to identify road safety issues [@songchitruksa2006].
 One prominent surrogate measure is traffic conflicts [@lord2021].
-A traffic conflict occurs when road users’ paths intersect, with a risk of collision if no action was taken [@tarko2018].
-The concept of using traffic conflicts as a surrogate measure was notably illustrated by Hyden’s safety pyramid [@hyden1987].
-This pyramid represents the spectrum of traffic events from safe, undisturbed passages at the base, to severe, rare crashes at the top.
+A traffic conflict occurs when road users’ paths intersect with a collision risk if no action was taken [@tarko2018].
+The use of traffic conflicts is illustrated by Hyden’s safety pyramid [@hyden1987], which represents the spectrum of traffic events from safe, undisturbed passages at the base, to severe, rare crashes at the top.
 Hyden’s model demonstrates the inverse relationship between crash frequency and severity.
-Understanding this relationship allows for the prediction of severe crashes based on the more common, less sever conflicts.
+Understanding this relationship allows for the prediction of severe crashes based on the more common, less severe conflicts.
 
 Traditional methods for observing traffic conflicts have relied on manual observation, which is resource intensive.
-This study highlights the potential of using open access data from Operation Snap, the UK police’s national system to receive road users’ video evidence of road traffic offences as a surrogate measure of road safety.
+The present paper highlights the potential of using open access data from Operation Snap, the UK police’s national system to receive road users’ video evidence of road traffic offences as a surrogate measure of road safety.
 
@@
 
+In what follows, it emerges that video submissions are, relative to road distance travelled, 20 times disproportionately reported by cyclists. 
 Cyclist crashes are underreported in police recorded datasets [@elvikIncompleteAccidentReporting1999] and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted [@janstrupetal.UnderstandingTrafficCrash2016].
@@
 
-Commuter cyclists in the UK experience a near miss for every six miles of riding [@aldredInvestigatingRatesImpacts2015a], and concern over near misses are a key reason people choose not to cycle [@sandersPerceivedTrafficRisk2015a], being associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers [@sandersPerceivedTrafficRisk2015a].
-In their review of near miss cycling crashes, @ibrahimCyclingMissesReview2021 highlighted the need for better data on near misses to inform safety research.
+Commuter cyclists in the UK experience a near miss for every six miles of riding [@aldredInvestigatingRatesImpacts2015a], and concern over near misses is a key reason people choose not to cycle [@sandersPerceivedTrafficRisk2015a].
+Near misses are associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers [@sandersPerceivedTrafficRisk2015a; @cubbin2024].
+A review of near-miss cycling crashes highlighted the need for better data to inform safety research [@ibrahimCyclingMissesReview2021].
 
@@
 A 'close pass' refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
-Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
-There is no specific law in the UK Road Traffic Act 1988 for driving too close to cyclist, but two are commonly applied for careless driving: RT88575- Drive without due care and attention; and RT99576 -- Driver without reasonable consideration to others - and these are prominent in the analysis that follows.
+<!-- Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024]. -->
+There is no specific law in the UK Road Traffic Act 1988 for driving too close to a cyclist, but two are commonly applied for careless driving: RT88575, driving without due care and attention; and RT99576, driving without reasonable consideration to others - and these are prominent in the analysis that follows.
 
@@
 It is now in operation nationally across England and Wales, each police force offering its own submission portal for road users to submit video evidence.
-The nature of video submissions and the related expectations are summarised on the website of one Police and Crime Commissioner as follows:
+The nature of video submissions and the related expectations were summarised on the website of one Police and Crime Commissioner as follows:
 
@@
 This is, to our knowledge, the first research study to use this dataset.
-As such, the study is offered as proof-of-concept of the potential for analysis of this data, and of its further potential, including understanding the factors associated with cyclist near miss incidents.
+As such, the study is offered as proof-of-concept of the potential for its further analysis to improve road safety.
 Following analysis and discussion of three years of data for one police force area, the study outlines a research agenda designed to inform policy and practice.
@@
 Within the Road Traffic Act these offences are related to careless driving and drivers are subject to similar penalties.
-The other common offences included failing to comply with traffic signals, traffic signs and the use of a handheld phone while driving.
+The other common offences included failing to comply with traffic signals, traffic signs and using a handheld phone while driving.
 <!-- TODO: Within the other category... -->

### 2024-07-22 paper.qmd
Assorted updates

@@
 The proportion of cyclists using these cameras is unknown but anecdotal evidence suggests increased usage, with many choosing to record rides in case an incident occurs.
+The proportions of horse riders, pedestrians and motor cyclists recording video footage is, to our knowledge, also unknown.
 
@@
 Since early 2022, there have been more monthly records in the OpSnap data than in the official 'STATS19' road traffic collision records for West Yorkshire, highlighting the under-reporting of road traffic incidents in official statistics.
-STATS19 record cs are from the Department for Transport's database of road traffic collisions reported to the police and only include incidents that result in injury.
+STATS19 records are from the Department for Transport's database of road traffic collisions reported to the police and only include incidents that result in injury.
 Like OpSnap data, STATS19 records are open access.

### 2024-07-25 paper.qmd
Dash-cam and other typos from Graham

@@
 Video footage is commonly recorded from on-board cameras.
-For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as 'dash cams'.
-One source examined GB Driving Licence data to find that by early 2024, close to a third of private and commercial vehicle in the UK had a dash cam [@DashCamSubmissions2024].
+For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as 'dash-cams'.
+One source examined GB Driving Licence data to find that, by early 2024, close to a third of private and commercial vehicle in the UK had a dash-cam [@DashCamSubmissions2024].
 For cyclists, footage is commonly recorded using helmet or handlebar-mounted cameras.
@@
 Studies focused on specific types of road user will prove informative with respect to policy and practice.
-Dash cam submissions by motorists offers great potential to inform other aspects of road safety.
+Dash-cam submissions by motorists offers great potential to inform other aspects of road safety.
 Research into submissions by pedestrians, horse riders and motorcyclists should prove viable to improve the safety of these parties.
@@
 This promises the potential to inform policy and practice in ways that ameliorate risk factors.
-Within this area of research, analysis from dash cams is likely to inform different practices than that from cycle-cams, that from pedestrians, horse riders and motorcyclists, and so on.
+Within this area of research, analysis from dash-cams is likely to inform different practices than that from cycle-cams, that from pedestrians, horse riders and motorcyclists, and so on.
 

### 2024-07-25 paper.qmd
Typo fixes

@@
 Like OpSnap data, STATS19 records are open access.
-They were downloaded with the `stats19` R package [@lovelace2019] and filtered to include only records from West Yorkshire to ensure comparability with the OpSnap data.
+For the results presented in @fig-time, STATS19 datasets were downloaded with the `stats19` R package [@lovelace2019] and filtered to include only records from West Yorkshire.
 
@@
 #| include: true
-#| tbl-cap: "Number of convictions by offence type (top 6 and other)."
+#| tbl-cap: "Offence types reported (top 6 and other)."
 #| width: 60%
@@
 d_offence_top_6 = d_mode_offence_count |>
-  filter(offence != "n/a") |>
+  # filter(offence != "n/a") |>
   group_by(offence) |>
@@
 d_offence_classified = d_offence_count |>
-  filter(offence != "n/a") |>
   mutate(
@@
   arrange(Offence == "Other", desc(`Number of records`)) |>
+  # Rename "n/a" to "No offence":
+  mutate(Offence = case_when(
+    Offence == "n/a" ~ "NA No offence or unknown offence type",
+    TRUE ~ Offence
+  )) |>
   mutate(`Percent of records` = round(`Number of records` / sum(`Number of records`) * 100, 1))
+
+d_offence_totals = d_offence_classified |>
+  summarise(`Number of records` = sum(`Number of records`), `Percent of records` = sum(`Percent of records`)) |>
+  mutate(Offence = "Total")
 
@@
 tbl = d_offence_classified |>
+  bind_rows(d_offence_totals) |>
   mutate(
@@
   gt_theme_espn()
-# gt::gtsave(tbl, "tbl-offences.png")
+# gt::gtsave(tbl, "tbl-offences.html")
+# webshot2::webshot("tbl-offences.html", "tbl-offences.png")
 # browseURL("tbl-offences.png")
@@
        TRUE ~ "Other"
-
      )
@@
 #| label: tbl-mode
-#| tbl-cap: Number of OpSnap records by observer mode.
+#| tbl-cap: Mode of transport of person submitting video evidence
 #| width: "80%"
@@
   # Reclassify Unknown to unknown:
-  arrange(desc(n)) |>
+  arrange(desc(n))
+tbl_totals = tbl |>
+  summarise(n = sum(n), percent_of_records = sum(percent_of_records)) |>
+  mutate(percent_of_records = round(percent_of_records)) |>
+  mutate(mode = "Total")
+tbl = bind_rows(tbl, tbl_totals) |>
   rename_all(snakecase::to_title_case) |>
@@
 # tbl
-# gt::gtsave(tbl, "tbl-mode.png")
+# gt::gtsave(tbl, "tbl-mode.html")
+# webshot2::webshot("tbl-mode.html", "tbl-mode.png")
 knitr::include_graphics("tbl-mode.png")
@@
 
-The cross-tabulation of reporting mode and offence type is shown in @tbl-mode-offences-crosstab, revealing some interesting tendencies.
+A cross-tabulation of transport mode (of the person submitting video evidence) and the offence type is shown in @tbl-mode-offences-crosstab.
+For both variables it shows the two main categories plus an 'other' category.
 
@@
 #| label: tbl-mode-offences-crosstab
-#| tbl-cap: Mode of transport from which video was recorded (columns) and offence type (rows), with counts and percentages within each reporting mode.
+#| tbl-cap: Mode of transport of person submitting video (columns).
 # pivot_combined |>
@@
 Disposal categories assigned by police are shown in @tbl-disposal.
-Roughly a third of cases (`r round(perc_no_further_action, 1)`%) resulted in no further action, and for most of the remainder drivers were required to undertake an education course (`r round(perc_course, 1)`%).
+Roughly a third of cases (`r round(perc_no_further_action, 1)`%) resulted in no further action and, for most of the remainder drivers were required to undertake an education course.
 Conditional offers, that is, drivers being offered a reduced penalty for admitting guilt, were the third most common outcome.
-Close to two percent of cases went to court and a further one percent underwent further investigation.
+Nearly two percent of cases went to court and a further one percent underwent further investigation.
 
@@
 
-There are `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
+There were `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
 
@@
 
-There was an increase in submissions over time between 2021 and 2023.
+There was an increase in submissions to opsnap over time between 2021 and 2023.
 This may reflect changing levels of road use during and after the Covid-19 pandemic, increased awareness of OpSnap, or increased video camera ownership levels.
-Comparative analysis including other police jurisdictions will be able to determine whether this experience is common or isolated.
-We hypothesise that it will prove to be a common experience and that that the volume of submissions will continue to increase in the near and middle term (though ideally, a decline would occur as roads become safer).
+Comparative analysis including other police jurisdictions should be undertaken to determine whether this experience is common or isolated.
+We hypothesise that it is a common experience and also that that the volume of submissions will continue to increase in the near and middle term (though ideally, a decline would occur as roads become safer).
 

### 2024-07-25 paper.qmd
Updates...

@@
 Data from one police force area for 39 months (January 2021 to March 2024) <!-- (N= 18,363 records) --> (N = `r nrow(d_all) |> scales::number(big.mark = ",")` records) is analysed.
-<!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians. -->
-Half were submitted by vehicle drivers (`r round(perc_vehicle_driver, 1)`%), a third by cyclists (`r round(perc_cyclist, 1)`%), `r round(perc_pedestrian, 1)`% by pedestrians, `r round(perc_horse, 1)`% by horse riders,  `r round(perc_motorcyclist, 1)`% by motorcyclists, and `r round(perc_other, 1)`% were unknown.
+<!-- Of submitted cases, 49.9% were from vehicle drivers, 34.4% were from cyclists, and 7.4% were from pedestrians.
+--> Half were submitted by vehicle drivers (`r round(perc_vehicle_driver, 1)`%), a third by cyclists (`r round(perc_cyclist, 1)`%), `r round(perc_pedestrian, 1)`% by pedestrians, `r round(perc_horse, 1)`% by horse riders, `r round(perc_motorcyclist, 1)`% by motorcyclists, and `r round(perc_other, 1)`% were unknown.
 We estimate that, relative to road distance travelled, cyclists were 20 times more likely to submit video evidence than vehicle drivers.
@@
 <!-- Two thirds (66.1%) of reported cases resulted in the recommended disposal of an educational course (including conditional offers), 31% no further action, and fewer than 1% for court appearance.
--->
-Half (`r round(perc_disposal_educational, 1)`%) of reported cases resulted in the recommended disposal of an educational course, `r round(perc_no_further_action, 1)`% no further action `r round(perc_disposal_conditional, 1)`% conditional offer, and `r round(perc_disposal_court, 1)`% resulted in court appearance.
+--> Half (`r round(perc_disposal_educational, 1)`%) of reported cases resulted in the recommended disposal of an educational course, `r round(perc_no_further_action, 1)`% no further action `r round(perc_disposal_conditional, 1)`% conditional offer, and `r round(perc_disposal_court, 1)`% resulted in court appearance.
 A research agenda using OpSnap data is outlined that could emerge if national datasets are compiled and responsibly opened-up and made available for research and policy-making: data-driven research should identify hotspot locations and other correlates of dangerous and antisocial road use at regional, and local levels; research projects should investigate disposal-related decision-making, video quality, and the role of supporting evidence; offence concentration (recidivism, repeat submitters of evidence, spatial hotspots) and case progression including court cases should be explored with reference to new video evidence.
@@
 A 'close pass' refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
-<!-- Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024]. -->
-There is no specific law in the UK Road Traffic Act 1988 for driving too close to a cyclist, but two are commonly applied for careless driving: RT88575, driving without due care and attention; and RT99576, driving without reasonable consideration to others - and these are prominent in the analysis that follows.
+<!-- Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
+--> There is no specific law in the UK Road Traffic Act 1988 for driving too close to a cyclist, but two are commonly applied for careless driving: RT88575, driving without due care and attention; and RT99576, driving without reasonable consideration to others - and these are prominent in the analysis that follows.
 
@@
 
-There was an increase in submissions to opsnap over time between 2021 and 2023.
+There has been an increase in submissions to West Yorkshire Police's OpSnap system since it was set-up in 2021.
 This may reflect changing levels of road use during and after the Covid-19 pandemic, increased awareness of OpSnap, or increased video camera ownership levels.
 Comparative analysis including other police jurisdictions should be undertaken to determine whether this experience is common or isolated.
-We hypothesise that it is a common experience and also that that the volume of submissions will continue to increase in the near and middle term (though ideally, a decline would occur as roads become safer).
-
-This was a case study of West Yorkshire but each of the 43 police forces in England and Wales collates OpSnap data, and our preliminary scoping of individual force websites suggests OpSnap data is publicly available nationwide.
-This suggests the potential for development of a national OpSnap dataset.
-A feasibility study is needed to establish the availability and compatibility of the data from different forces.
-Should a national dataset prove feasible it holds potential for national-level, cross-regional, and comparative analyses of patterns and trends.
-A national dataset will enhance the possibility for cross-national comparative analysis and international cooperation in road safety.
-It holds the potential for the development of rankings according to different criteria and, thereby, potential performance indicators.
-
-If submission volumes continue to increase, means of informing triage and prioritisation of cases will be increasingly valuable.
-Triage systems are probably already in use in different areas, and research to identify and inform best practice should be undertaken.
-Likewise, research to determine best practice in determining disposal recommendations will prove fruitful.
+We hypothesise that it is a common experience for police forces (and other organisations) setting up public video submissions reporting antisocial and dangerous driving and that that the rate of submissions will continue to increase (longer term, ideally, submission rates decline as roads become safer).
+
+This was a case study of West Yorkshire but each of the 43 police forces in England and Wales collates OpSnap data.
+Our preliminary scoping of individual force websites suggests OpSnap data is publicly available nationwide.
+A national dataset should be developed that includes OpSnap data from all police forces.
+ A feasibility study should be undertaken to establish cross-force data availability and compatibility.
+A national dataset holds potential for national-level, cross-regional, and comparative analyses of patterns and trends.
+It would have the potential to promote cross-national comparative analysis and international cooperation in road safety should similar data be collated elsewhere.
+It would hold the potential for the development of rankings according to different criteria and, thereby, potential performance-related metrics [@tiwana2015].
+There will be added-value from a national dataset when multiple incidents involve the same persons in different jursdictions, such as recidivist drivers, and in relation to other cross-jurisdictional issues.
+
+With submission volumes likely to continue to increase, means of informing triage and prioritisation of cases will be increasingly valuable.
+Triage systems are almost certainly used already, and research to identify best practice should be undertaken.
+Similarly, research to determine best practice in determining disposal recommendations should be undertaken.
 It may provide feasible, perhaps using machine learning, to develop automated means of triage, determination of offence type and disposal.
-Research to gauge the differential effectiveness of disposal measures is needed, starting with measures of recidivism.
+Research to gauge the differential effectiveness of disposal measures is also needed.
+A starting point for an evaluation may be easures of recidivism found in the OpSnap data.
 
 Within the trend of increasing video submissions there were distinct seasonal patterns.
-Seasonal variation in cyclists' submission reflect seasonal (weather and other) influences on cycling prevalence and duration.
+Seasonal variation in submissions reflect seasonal (weather and other) influences, and these will vary by transport mode: cycling, for example, is less prevalent in winter.
 This suggests seasonal variation in preventive responses might be tailored to need.
+
+Future research should focus on different types of problem concentration.
 It is well established that crime is highly concentrated along whatever dimension is examined [@farrellPreventingRepeatRepeat2017].
 With respect to video evidence this will include recidivists (repeat offenders), repeat submitters of evidence (who may or may not be repeat victims/survivors), the concentration of incidents (close passes, near misses and crashes) at certain times and places, with different types of experiences concentrated among certain types of road users.
-For example, we hypothesise that close passes are more likely to be reported by cyclists and horse riders than vehicle drivers, both due to the nature of the road user interactions and the relative risk associated with the manoeuvre.
-As such, the types of incidents reported by vehicle drivers would be systematically different, indicating a need for different investigative and preventive approaches.
-Policing and preventions efforts focused on where problems are concentrated are more resource efficient -- it is the foundation of problem-oriented policing (see e.g., (@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010)- and research using video submissions holds the potential to analysis of problem concentrations.
+As discussed in the introduction, close passes are more likely to be reported by cyclists and horse riders than vehicle drivers, both due to the nature of the road user interactions and the relative risk associated with the maneuver.
+
+There was preliminary confirmation of this pattern in our analysis which found that offence types reported by vehicle drivers are systematically different to those of other road users.
+Hence future research should consider that investigative and preventive approaches should be tailored to different contexts and types of road user.
+Rural roads account for a disproportionate amount of fatalities @brake, and further research using video evidence may inform preventive approaches tailored to road type.
+
+Policing and prevention efforts focused on where problems are concentrated are more resource efficient — it is the foundation of problem-oriented policing (see e.g., (@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010).
 

### 2024-07-25 paper.qmd
Update tbls

@@
 
-tbl = d_all |>
+tbl_disposal = d_all |>
   count(disposal, sort = TRUE) |>
   mutate(percent_of_records = round(n / nrow(d_all), 3) * 100) |>
-  arrange(desc(n)) |>
+  arrange(desc(n))
+
+tbl_disposal_totals = tbl_disposal |>
+  summarise(n = sum(n), percent_of_records = round(sum(percent_of_records))) |>
+  mutate(disposal = "Total")
+
+tbl = bind_rows(tbl_disposal, tbl_disposal_totals) |>
   rename_all(snakecase::to_title_case) |>
@@
   gt_theme_espn()
-# gt::gtsave(tbl, "tbl-disposal.png")
+# gt::gtsave(tbl, "tbl-disposal.html")
+# webshot2::webshot("tbl-disposal.html", "tbl-disposal.png")
 knitr::include_graphics("tbl-disposal.png")

### 2024-07-25 paper.qmd
Complete updates to section 4

@@
 
-This was a case study of West Yorkshire but each of the 43 police forces in England and Wales collates OpSnap data, and our preliminary scoping of individual force websites suggests OpSnap data is publicly available nationwide.
-This suggests the potential for development of a national OpSnap dataset.
-A feasibility study is needed to establish the availability and compatibility of the data from different forces.
-Should a national dataset prove feasible it holds potential for national-level, cross-regional, and comparative analyses of patterns and trends.
-A national dataset will enhance the possibility for cross-national comparative analysis and international cooperation in road safety.
-It holds the potential for the development of rankings according to different criteria and, thereby, potential performance indicators.
-
 Comparison by complainant mode of transport, reported offense type and disposal, subdivided by other characteristics, is needed.
@@
 A pilot project should be undertaken in collaboration with police to establish the potential to further enhance policing and public safety using non-public aspects of the submissions data.
-This would require working partnerships and a secure research platform to meet GDPR requirements.
+This would require working partnerships and a secure research platform to meet GDPR requirements and Data Protection Act (2018).
 Such collaborative approaches are increasingly common in health, medicine, and policing research.
@@
 There are also likely to be lessons that can be learned for how footage is gathered, edited and submitted by road users.
-More generally, video footage might be used to identification risk factors, that is, the types of situations in which crashes, near misses, and close passes occur.
-This promises the potential to inform policy and practice in ways that ameliorate risk factors.
-Within this area of research, analysis from dash-cams is likely to inform different practices than that from cycle-cams, that from pedestrians, horse riders and motorcyclists, and so on.
+
+Analysis of video footage should be undertaken to identify risk factors, that is, the types of situations in which crashes, near misses, close passes and other offences occur.
+Such research can inform policy and practice in ways that ameliorate risk.
+Within this area of research, analysis from dash cams is likely to inform different practices than that from cycle-cams, that from pedestrians, horse riders and motorcyclists, and so on.
 
@@
 What are the characteristics of strong supporting written evidence, and what are the characteristics of strong video evidence?
-Are both needed or can a weakness in one be overcome by particularly strong aspects in the other?
+Do both need to be ‘strong’ or can a weakness in one be overcome by particularly strong aspects in the other?
+Research should identify further aspects of good practice for those submitting evidence to police.
 
@@
 
-Creating opportunities to access video submission data would be an alternate approach that facilitates research into the nature of footage.
-Another option would be research using existing online videos, of which there appear to be many thousands on social media including those posted by police forces.
-Such research will face different methodological challenges but could provide proof-of-concept of the potential for analysis of OpSnap footage.
-
-OpSnap data should be compared to, and integrated with, other road safety datasets.
-For example, spatial analysis that was Integrated with Strava data (road use data by cyclists and pedestrians), Google road use and other data, may facilitate exposure-based measures rather than the counts used here.
-The use of rates will enhance the identification locations with higher risk because based on their volume of traffic: different types of risk need to be gauged.
-Comparison and integration of OpSnap and STATS19 datasets could provide a logical starting point.
-To what extent are OpSnap and STATS19 records overlapping or complementary?
-To what extent do they identify the same or different types of problems?
-
-More generally, OpSnap data holds potential for use in the evaluation of experimental interventions.
-Depending on the level of submissions there is the potential to use the data to assess effects before and after interventions, and to identify comparison (control group) sites.
+Some research into the relationship between video footage and supporting written evidence is needed.
+Is written supported evidence always needed?
+Which is deemed most important by police, and which by courts?
+What are the characteristics of strong supporting written evidence, and what are the characteristics of strong video evidence?
+Do both need to be ‘strong’ or can a weakness in one be overcome by particularly strong aspects in the other?
+Research should identify further aspects of good practice for those submitting evidence to police.
+
+We did not include spatial analysis here, but the data hold that potential.
+The public data ued here only allowed spatial analysis for ‘all’ video submissions.
+That is, we were unable to distinguish the geographic location of submissions by different types of road users because, or for different offence types, because the locations were scrambled for data protection purposes.
+Future research, in the context of a secure data platform, should include spatial analysis for different types of road users, for different types of incidents, for incidents resulting in different disposals.
+Hotspots and spatial clustering will vary by type of road user, type of incident, day or week, time of day, and so on, and parsing the data will produce more informed road safety research.
+
+Accessing OpSnap data in collaboration with police is likely to prove the most fruitful approach for future research.
+Were that impractical, research using existing online videos should be undertaken.
+There are many thousands of such videos on social media in the public realm, including those posted by police forces, and while there are different sampling issues to consider, this sources offers a plausible alternative route to informing road safety.
+
+OpSnap data should be compared to, and integrated with, other road use and road safety datasets.
+Here we offered preliminary comparison to the volume of Stats19 data on road crashes.
+patial analysis integrated with Strava data (road use data by cyclists and pedestrians), Google road use and other data, may facilitate exposure-based measures rather than the counts used here.
+The use of rates will enhance the identification of locations with higher risk other than that due to volume of traffic.
+ 
+Future research should recognise that OpSnapdata holds the potential for use in the evaluation of experimental interventions.
+It may offer the potential for pre-post intervention comparative evaluations using control sites.
 Different road safety interventions imply different resource needs.

### 2024-07-25 paper.qmd
Final edits

@@
 quarto::quarto_render("paper.qmd", output_format = "pdf")
+file.rename("paper.pdf", "paper-v10.pdf")
 quarto::quarto_render("paper.qmd", output_format = "docx")
-file.rename("paper.docx", "~/OneDrive/opsnap_leeds/paper-v8.docx")
-file.rename("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v8.pdf")
+# file.rename("paper.docx", "/OneDrive/opsnap_leeds/paper-v10.docx")
+file.rename("paper.docx", "paper-v10.docx")
+system("gh release upload v1 paper-v10.docx paper-v10.pdf")
 browseURL("paper.docx")
@@
   rename_all(snakecase::to_title_case) |>
-  knitr::kable() |>
-  # Striped styling and tiny text to fit:
-  kableExtra::kable_styling(latex_options = "striped") |>
-  # Set column widths:
-  kableExtra::column_spec(1, width = "3em") |>
-  kableExtra::column_spec(2, width = "3em") |>
-  kableExtra::column_spec(3, width = "3em") |>
-  kableExtra::column_spec(4, width = "3em") |>
-  kableExtra::column_spec(5, width = "9em") |>
-  kableExtra::column_spec(6, width = "4em") |>
-  kableExtra::column_spec(8, width = "8em")
+  knitr::kable()
+
+  # |>
+  # # Striped styling and tiny text to fit:
+  # kableExtra::kable_styling(latex_options = "striped") |>
+  # # Set column widths:
+  # kableExtra::column_spec(1, width = "3em") |>
+  # kableExtra::column_spec(2, width = "3em") |>
+  # kableExtra::column_spec(3, width = "3em") |>
+  # kableExtra::column_spec(4, width = "3em") |>
+  # kableExtra::column_spec(5, width = "9em") |>
+  # kableExtra::column_spec(6, width = "4em") |>
+  # kableExtra::column_spec(8, width = "8em")
 ```
@@
 There were `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
+The locations were scrambled by West Yorkshire Police before being made available for data protection purposes, which meant spatial analysis would only possible for ‘all’ video submissions, ignoring different types of road users or offence types.
 
@@
 It would hold the potential for the development of rankings according to different criteria and, thereby, potential performance-related metrics [@tiwana2015].
-There will be added-value from a national dataset when multiple incidents involve the same persons in different jursdictions, such as recidivist drivers, and in relation to other cross-jurisdictional issues.
+There will be added-value from a national dataset which allows identification of cross-jurisdictional issues such as the same vehicles or persons involved in incidents in different police force areas.
 
@@
 Research to gauge the differential effectiveness of disposal measures is also needed.
-A starting point for an evaluation may be easures of recidivism found in the OpSnap data.
-
 Within the trend of increasing video submissions there were distinct seasonal patterns.
@@
 Comparison by complainant mode of transport, reported offense type and disposal, subdivided by other characteristics, is needed.
-Determining which type of submission - by which type of road user for example - is more likely to result in a recommendation of court proceedings, may help refine police investigations.
+Determining which type of submission - by which type of road user for example — is more likely to result in a recommendation of court proceedings, may help refine police investigations.
 For example, are submissions by cyclists (or horse riders) more or less likely to result in court proceedings than those by motorists?
@@
 The use of rates will enhance the identification of locations with higher risk other than that due to volume of traffic.
- 
   Future research should recognise that OpSnapdata holds the potential for use in the evaluation of experimental interventions.
@@
 It provides proof-of-concept that OpSnap data holds considerable research potential.
-It has the potential to inform policy and practice promoting road safety.
-Drawing on the analysis of the West Yorkshire data, a research agenda was outlined.
-This preliminary list will prove far from exhaustive.
-While aspects of that agenda may appear ambitious at the time of writing, we suggest that the present study will have achieved one of its aims if this agenda is rapidly superseded by further research.
-
-The focus here was video evidence submissions by cyclists.
-Relative to distance travelled by cyclists, their submissions were around 20 times over-represented in the West Yorkshire data.
-This was the justification for our focus.
-The disproportionality likely reflects the vulnerability of cyclists relative to motor vehicle users when the two forms of transport interact, and when close passes, near misses and crashes occur.
-Our primary conclusion is that further research into video submissions to Operation Snap holds significant potential to inform the future of road safety for a broad range of road users.
+OpSnap data has, we suggest, significant potential to inform policy and practice promoting road safety.
+Drawing on the preliminary analysis of the West Yorkshire data, a  research agenda was outlined.
+Our preliminary list will prove far from exhaustive.
+And while aspects of that agenda may appear ambitious at the time of writing, we suggest that the present study will have achieved one of its aims if this agenda is rapidly superseded.
 

### 2024-07-25 paper.qmd
Fix brake reference

@@
 Hence future research should consider that investigative and preventive approaches should be tailored to different contexts and types of road user.
-Rural roads account for a disproportionate amount of fatalities @brake, and further research using video evidence may inform preventive approaches tailored to road type.
-
-Policing and prevention efforts focused on where problems are concentrated are more resource efficient — it is the foundation of problem-oriented policing (see e.g., (@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010).
+Rural roads account for a disproportionate amount of fatalities @brakeDirectLineBrake2018, and further research using video evidence may inform preventive approaches tailored to road type.
+Policing and prevention efforts focused on where problems are concentrated are more resource efficient and is the foundation of problem-oriented policing (see e.g., (@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010).
 

### 2024-07-25 paper.qmd
Brake ref

@@
   rename_all(snakecase::to_title_case) |>
-  knitr::kable()
-
-  # |>
-  # # Striped styling and tiny text to fit:
-  # kableExtra::kable_styling(latex_options = "striped") |>
-  # # Set column widths:
-  # kableExtra::column_spec(1, width = "3em") |>
-  # kableExtra::column_spec(2, width = "3em") |>
-  # kableExtra::column_spec(3, width = "3em") |>
-  # kableExtra::column_spec(4, width = "3em") |>
-  # kableExtra::column_spec(5, width = "9em") |>
-  # kableExtra::column_spec(6, width = "4em") |>
-  # kableExtra::column_spec(8, width = "8em")
+  knitr::kable() |>
+  # Striped styling and tiny text to fit:
+  kableExtra::kable_styling(latex_options = "striped") |>
+  # Set column widths:
+  kableExtra::column_spec(1, width = "3em") |>
+  kableExtra::column_spec(2, width = "3em") |>
+  kableExtra::column_spec(3, width = "3em") |>
+  kableExtra::column_spec(4, width = "3em") |>
+  kableExtra::column_spec(5, width = "9em") |>
+  kableExtra::column_spec(6, width = "4em") |>
+  kableExtra::column_spec(8, width = "8em")
 ```

### 2024-07-29 README.qmd
Update with cjs codes
Update readme
update readme
md version
Add bib

@@
 
+# Note
+
+This package is in development and not yet on CRAN.
+If you would like to use it, please install it from GitHub as shown below, and cite it as follows:
+
+Farrell, G., Lovelace, R., & O'Hern, S.
+(2024).
+Road User Video Evidence of Road Traffic Offences: Preliminary Analysis of Operation Snap Data and Suggestions for a Research Agenda.
+https://doi.org/10.31235/osf.io/cgjmr
+
+```
+@misc{farrellRoadUserVideo2024,
+  title = {Road {{User Video Evidence}} of {{Road Traffic Offences}}: {{Preliminary Analysis}} of {{Operation Snap Data}} and {{Suggestions}} for a {{Research Agenda}}},
+  shorttitle = {{{ROAD USER VIDEO EVIDENCE OF ROAD TRAFFIC OFFENCES}}},
+  author = {Farrell, Graham and Lovelace, Robin and O'Hern, Steve},
+  year = {2024},
+  month = jul,
+  publisher = {OSF},
+  doi = {10.31235/osf.io/cgjmr},
+  urldate = {2024-07-29},
+  archiveprefix = {OSF},
+  langid = {american},
+  keywords = {antisocial driving,dangerous driving,near misses,Operation Snap,OpSnap,roas safety,Video evidence}
+}
+
+```
+
+Read the full paper [here](https://osf.io/cgjmr/download/).
+
 # Installation
@@
 
-```r
-remotes::install_github("ITSLeeds/opsnap")
+```{r}
+#| eval: false
+remotes::install_github("ITSLeeds/opsnap", force = TRUE, dependencies = TRUE)
 ```

### 2024-07-29 paper.qmd
Minor updates

@@
     wrap: sentence
-author:
-  - name: Graham Farrell
-    corresponding: true
-    affiliations:
-      - name: University of Leeds
-        department: School of Law
-        address: The Liberty Building
-        city: Leeds
-        country: United Kingdom
-        postal-code: LS2 9JT
-    orcid: 0000-0002-3987-8457
-  - name: Robin Lovelace
-    affiliations:
-      - name: University of Leeds
-        department: Institute for Transport Studies
-        address: University Road
-        city: Leeds
-        country: United Kingdom
-        postal-code: LS2 9JT
-      - name: Active Travel England
-        # department: Data and Digital
-        # address: West Offices Station Rise
-        # city: York
-        # country: United Kingdom
-        # postal-code: YO1 6GA
-    orcid: 0000-0001-5679-6536
-  - name: Steve O'Hern
-    affiliations:
-      - name: University of Leeds
-        department: Institute for Transport Studies
-        address: University Road
-        city: Leeds
-        country: United Kingdom
-        postal-code: LS2 9JT
-      - name: Monash University Accident Research Centre
-        address: Clayton, VIC 3800, Australia
-    orcid: 0000-0002-7961-3875
 keywords: 
@@
   - Video evidence
-  - Dashcam
   - Near misses
+  - Operation Snap
+  - Dangerous driving
+  - Antisocial driving
 date: last-modified
@@
 #| echo: false
-quarto::quarto_render("paper.qmd", output_format = "pdf")
-file.rename("paper.pdf", "paper-v10.pdf")
-quarto::quarto_render("paper.qmd", output_format = "docx")
-# file.rename("paper.docx", "/OneDrive/opsnap_leeds/paper-v10.docx")
-file.rename("paper.docx", "paper-v10.docx")
-system("gh release upload v1 paper-v10.docx paper-v10.pdf")
-browseURL("paper.docx")
-# Install extension:
-system("quarto install extension mikemahoney218/quarto-arxiv")
-system("sudo apt install lmodern")
+# quarto::quarto_render("paper.qmd", output_format = "pdf")
+# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-no-line-numbers-v11.pdf")
+# quarto::quarto_render("paper.qmd", output_format = "arxiv-pdf")
+# file.copy("paper.pdf", "paper-v11.pdf")
+# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v11.pdf")
+# quarto::quarto_render("paper.qmd", output_format = "docx")
+# # file.copy("paper.docx", "~/OneDrive/opsnap_leeds/paper-v11.docx")
+# file.copy("paper.docx", "paper-v11.docx")
+# quarto::quarto_render("title.qmd", output_format = "arxiv-pdf")
+# browseURL("title.pdf")
+# system("gh release upload v1 --clobber paper-v11.pdf title.pdf")
+# browseURL("paper.docx")
+# # Install extension:
+# system("quarto install extension mikemahoney218/quarto-arxiv")
+# system("sudo apt install lmodern")
 ```
@@
 
-There were `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
+There were `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations corresponding to busy roads: Meanwood Road (Leeds), Dewsbury Road (Wakefield) and Chapeltown Road (Leeds), with no single address accounting for more than 0.4% of records.
 The locations were scrambled by West Yorkshire Police before being made available for data protection purposes, which meant spatial analysis would only possible for ‘all’ video submissions, ignoring different types of road users or offence types.
@@
 #| echo: false
+#| eval: false
 #| label: tbl-locations
@@
 # gt::gtsave(tbl, "tbl-locations.png")
-knitr::include_graphics("tbl-locations.png")
+# knitr::include_graphics("tbl-locations.png")
 ```
@@
 Rural roads account for a disproportionate amount of fatalities @brakeDirectLineBrake2018, and further research using video evidence may inform preventive approaches tailored to road type.
-Policing and prevention efforts focused on where problems are concentrated are more resource efficient and is the foundation of problem-oriented policing (see e.g., (@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010).
+Policing and prevention efforts focused on where problems are concentrated are more resource efficient and is the foundation of problem-oriented policing [@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010].
 
@@
 
+# Acknowledgements
+
+[To be added after peer review]
+
+<!--
+
+We would like to thank Ian King for making the connections that led to this paper being written.
+Thanks to Roger Beecham who wrote code for visualising some of the tables/figures in the paper, testing and supporting with reproducibility. 
+Thanks to Simon D'vali and colleagues from City of Bradford Metropolitan Council for their support and advice on the project and colleagues from the Safer Cycling Yorkshire group.
+
+-->
+
 # References

### 2024-07-29 R/get_cjs_codes.R
Document

@@ -3,6 +3,8 @@
 #' This function gets the criminal justice codes from the Criminal Justice Hub website.
 #' See [www.criminaljusticehub.org.uk](https://www.criminaljusticehub.org.uk/jargon-buster/cjs-offence-code/)
 #' 
+#' @param u The URL of the Criminal Justice Hub website.
+#' 
 #' @export 
 #' @examples
 #' get_cjs_codes()

### 2024-07-29 R/opsnap.R
Update

@@ -7,7 +7,10 @@
 #' @export
 #' @examples
 #' if (FALSE) {
-#' u = "https://www.westyorkshire.police.uk/sites/default/files/2023-10/operation_snap_july-sept_2023.xlsx"
+#' u = paste0(
+#'   "https://www.westyorkshire.police.uk/sites/default/files/",
+#'   "2023-10/operation_snap_july-sept_2023.xlsx"
+#' )
 #' download_and_read(, dir = ".")
 #' }
 download_and_read = function(u, remove_nas = FALSE, dir = "raw_data/west-yorkshire") {

### 2024-09-16 paper.qmd
Tests

@@
     wrap: sentence
-keywords: 
-  - Road safety
-  - Video evidence
-  - Near misses
-  - Operation Snap
-  - Dangerous driving
-  - Antisocial driving
+# keywords: 
+#   - Road safety
+#   - Video evidence
+#   - Near misses
+#   - Operation Snap
+#   - Dangerous driving
+#   - Antisocial driving
 date: last-modified
@@
     keep-tex: true
-linenumbers: true # Add (continuous) line numbers?
-doublespacing: false # Double space the PDF output?
-runninghead: "Preprint" # The text on the top of each page of the output
-authorcols: true # Should authors be listed in a single column (default) or in multiple columns (`authorcols: true`)
-editor_options: 
-  markdown: 
-    wrap: sentence
+# linenumbers: true # Add (continuous) line numbers?
+# doublespacing: false # Double space the PDF output?
+# runninghead: "Preprint" # The text on the top of each page of the output
+# authorcols: true # Should authors be listed in a single column (default) or in multiple columns (`authorcols: true`)
 execute: 
@@
 # quarto::quarto_render("paper.qmd", output_format = "pdf")
-# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-no-line-numbers-v11.pdf")
+# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-no-line-numbers-v12.pdf")
 # quarto::quarto_render("paper.qmd", output_format = "arxiv-pdf")
-# file.copy("paper.pdf", "paper-v11.pdf")
-# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v11.pdf")
+# file.copy("paper.pdf", "paper-v12.pdf")
+# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v12.pdf")
 # quarto::quarto_render("paper.qmd", output_format = "docx")
-# # file.copy("paper.docx", "~/OneDrive/opsnap_leeds/paper-v11.docx")
-# file.copy("paper.docx", "paper-v11.docx")
+# # file.copy("paper.docx", "~/OneDrive/opsnap_leeds/paper-v12.docx")
+# file.copy("paper.docx", "paper-v12.docx")
 # quarto::quarto_render("title.qmd", output_format = "arxiv-pdf")
 # browseURL("title.pdf")
-# system("gh release upload v1 --clobber paper-v11.pdf title.pdf")
+# system("gh release upload v1 --clobber paper-v12.docx")
 # browseURL("paper.docx")
@@
 #| echo: false
-#| include: false
+#| eval: false
 library(rvest)
@@
   knitr::kable()
-
+```
+
+```{r}
+#| include: false
 # In preparation for reading-in all data:
@@
   rename_all(snakecase::to_title_case) |>
-  knitr::kable() |>
-  # Striped styling and tiny text to fit:
-  kableExtra::kable_styling(latex_options = "striped") |>
-  # Set column widths:
-  kableExtra::column_spec(1, width = "3em") |>
-  kableExtra::column_spec(2, width = "3em") |>
-  kableExtra::column_spec(3, width = "3em") |>
-  kableExtra::column_spec(4, width = "3em") |>
-  kableExtra::column_spec(5, width = "9em") |>
-  kableExtra::column_spec(6, width = "4em") |>
-  kableExtra::column_spec(8, width = "8em")
+  knitr::kable()
+
+  # # For PDF:
+  # #  |>
+  # # Striped styling and tiny text to fit:
+  # kableExtra::kable_styling(latex_options = "striped") |>
+  # # Set column widths:
+  # kableExtra::column_spec(1, width = "3em") |>
+  # kableExtra::column_spec(2, width = "3em") |>
+  # kableExtra::column_spec(3, width = "3em") |>
+  # kableExtra::column_spec(4, width = "3em") |>
+  # kableExtra::column_spec(5, width = "9em") |>
+  # kableExtra::column_spec(6, width = "4em") |>
+  # kableExtra::column_spec(8, width = "8em")
 ```

### 2024-09-16 paper.qmd
Tests

@@
 We conclude that datasets derived from publicly-uploaded video submission portals have the potential to transform evidence-based policy and practice locally, nationally and internationally.
-
-# Introduction
-
-Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
-There were 1,711 fatalities and 135,480 causalities across all severity categories (slight, serious fatal) in Britain resulting from car crashes and other road traffic collisions reported by the police in 2022 [@departmentfortransportReportedRoadCasualties2023].
-This is just the tip of the iceberg of road traffic incidents, with many more near misses and minor incidents going unreported.
-
-Traditionally, road safety research has relied on retrospective analysis of crashes to assess and enhance road safety [@towards2008].
-However, relying on historical crash data poses several challenges.
-Crashes are rare events [@elvik2006], requiring extended periods, often at least five years, of data to obtain statistically significant estimates [@songchitruksa2006].
-Furthermore, the reactive nature of crash analysis means that safety improvements only follow after crashes occur, which is both inefficient and ethically problematic.
-
-Surrogate safety measures offer an alternative to casualty data by using more frequently observable, less severe, traffic events to identify road safety issues [@songchitruksa2006].
-One prominent surrogate measure is traffic conflicts [@lord2021].
-A traffic conflict occurs when road users’ paths intersect with a collision risk if no action was taken [@tarko2018].
-The use of traffic conflicts is illustrated by Hyden’s safety pyramid [@hyden1987], which represents the spectrum of traffic events from safe, undisturbed passages at the base, to severe, rare crashes at the top.
-Hyden’s model demonstrates the inverse relationship between crash frequency and severity.
-Understanding this relationship allows for the prediction of severe crashes based on the more common, less severe conflicts.
-
-Traditional methods for observing traffic conflicts have relied on manual observation, which is resource intensive.
-The present paper highlights the potential of using open access data from Operation Snap, the UK police’s national system to receive road users’ video evidence of road traffic offences as a surrogate measure of road safety.
-
-<!-- Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics [@departmentfortransportReportedRoadCasualties2023].
-Between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year in Britain according to police records [@departmentfortransport].
-Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
-The 2023 report noted that "the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was 'driver or rider failed to look properly'" [@departmentfortransport].
--->
-<!-- The context for the study is that, between 2004 and 2022, an average of 104 cyclists were killed and 4,212 were seriously injured each year according to official records recorded by police forces across Great Britain [@departmentfortransport].
-Almost half of cyclist fatalities involved collision with a car, with 56% on rural roads (compared to 30% of traffic).
-The 2023 report noted that “the most common contributory factor allocated to pedal cyclists in fatal or serious collisions (FSC) with another vehicle was ‘driver or rider failed to look properly’” [@departmentfortransport].  -->
-
-In what follows, it emerges that video submissions are, relative to road distance travelled, 20 times disproportionately reported by cyclists.
-Cyclist crashes are underreported in police recorded datasets [@elvikIncompleteAccidentReporting1999] and the extent of injuries sustained by cyclists may be higher when hospital-recorded cases are counted [@janstrupetal.UnderstandingTrafficCrash2016].
-Under-reporting in police datasets is likely to be greatest for minor injuries, while records are seldom kept when a collision or injury is avoided due to riders or drivers taking evasive actions [@ibrahimCyclingMissesReview2021].
-These incidents are often referred to as "near-misses" [@ibrahimCyclingMissesReview2021]
-
-Commuter cyclists in the UK experience a near miss for every six miles of riding [@aldredInvestigatingRatesImpacts2015a], and concern over near misses is a key reason people choose not to cycle [@sandersPerceivedTrafficRisk2015a].
-Near misses are associated with inattentive driving, aggressive driving, driving too fast, passing too close, being car-doored, and being cut off by turning drivers [@sandersPerceivedTrafficRisk2015a; @cubbin2024].
-A review of near-miss cycling crashes highlighted the need for better data to inform safety research [@ibrahimCyclingMissesReview2021].
-
-Close passes are the most common type of near miss reported by cyclists and are associated with collisions resulting in injury [@aldred2016].
-A 'close pass' refers to when a vehicle passes too close to a cyclist, which is defined in the UK as less than 1.5 metres away at 30mph (50kph).
-<!-- Close passes take different forms including the 'punishment pass' by angry drivers for a perceived slight such as causing the driver to slow down [@cubbin2024].
---> There is no specific law in the UK Road Traffic Act 1988 for driving too close to a cyclist, but two are commonly applied for careless driving: RT88575, driving without due care and attention; and RT99576, driving without reasonable consideration to others - and these are prominent in the analysis that follows.
-
-Operation Snap, often referred to informally as 'OpSnap', was piloted by North Wales police in October 2016 and adopted by all Welsh forces by 2018.
-It is now in operation nationally across England and Wales, each police force offering its own submission portal for road users to submit video evidence.
-The nature of video submissions and the related expectations were summarised on the website of one Police and Crime Commissioner as follows:
-
--   The secure form is for traffic offences, it is NOT for submitting footage of road traffic collisions, any other offences or for parking issues. 
--   The car registration number of the offending vehicle must be clearly visible. 
--   The public should be prepared to sign a witness statement and possibly give evidence in court. 
--   Statements for OpSnap can only be accepted from persons aged 18 or over. If you are under 18 the incident should be reported by email.
-
-This is, to our knowledge, the first research study to use this dataset.
-As such, the study is offered as proof-of-concept of the potential for its further analysis to improve road safety.
-Following analysis and discussion of three years of data for one police force area, the study outlines a research agenda designed to inform policy and practice.
-
-# Methods and data {#sec-methods}
-
-Open access data from West Yorkshire Safer Roads from the OpSnap project was used for this study.
-The media submissions portal opened in July 2020, and available data from the West Yorkshire Police (WYP) used in this paper span the calendar years 2021, 2022 and 2023.
-For the year 2021 there was less than half the cases of either 2022 or 2023, which could reflect reduced road use during the COVID-19 pandemic and fewer people being aware of OpSnap when it commenced.
-
-The dataset is a tabulation of cases submitted to the West Yorkshire Safer Roads OpSnap web portal.
-The terms 'record' and 'case' are used interchangeably here to refer to a record in the OpSnap database, each of which represents the separate submission of video evidence by a road user.
-
-The portal allows members of the public to submit video footage of suspected traffic offences committed by motor vehicle drivers.
-Video footage is commonly recorded from on-board cameras.
-For motor vehicles these cameras are typically mounted on or near the vehicle front dashboard, known as 'dash-cams'.
-One source examined GB Driving Licence data to find that, by early 2024, close to a third of private and commercial vehicle in the UK had a dash-cam [@DashCamSubmissions2024].
-For cyclists, footage is commonly recorded using helmet or handlebar-mounted cameras.
-The proportion of cyclists using these cameras is unknown but anecdotal evidence suggests increased usage, with many choosing to record rides in case an incident occurs.
-The proportions of horse riders, pedestrians and motor cyclists recording video footage is, to our knowledge, also unknown.
-
-Complainants upload footage and complete a short form that includes their personal details, the details of the vehicle involved including registration, make, model and colour, the location and time of the incident, and details of the camera used to record the footage.
-Only vehicle offences can be reported as the registration number of any offending vehicle is required and must be legible in footage.
-The open access data is a deidentified summary of submitted cases with information on mode of transport of the person reporting, offender vehicle details (make, model, colour), offence code, recommended disposal, date of submission, district, and offence location.
-The offence location is typically a street name and town or city name, or an intersection, and examples included: 'A58 Godley Road, Halifax', 'Keighley Road, Silsden', 'Woodhouse Lane A660, Leeds'.
-For this study, approximate geolocations were obtained using Google API, restricting cases to within West Yorkshire.
-Further aspects of the data, their uses and limitations are discussed in what follows.
-
-A random sample of 5 records from the raw data is shown in @tbl-raw (note: "nfa" refers to "no further action").
-
-```{r}
-#| label: tbl-raw
-#| tbl-cap: "Random sample of 5 records from the raw data."
-set.seed(24)
-d_all |>
-  sample_n(5) |>
-  select(-district, -`offence_code`) |>
-  mutate(date = as.Date(date)) |>
-  arrange(date) |>
-  rename_all(snakecase::to_title_case) |>
-  knitr::kable()
-
-  # # For PDF:
-  # #  |>
-  # # Striped styling and tiny text to fit:
-  # kableExtra::kable_styling(latex_options = "striped") |>
-  # # Set column widths:
-  # kableExtra::column_spec(1, width = "3em") |>
-  # kableExtra::column_spec(2, width = "3em") |>
-  # kableExtra::column_spec(3, width = "3em") |>
-  # kableExtra::column_spec(4, width = "3em") |>
-  # kableExtra::column_spec(5, width = "9em") |>
-  # kableExtra::column_spec(6, width = "4em") |>
-  # kableExtra::column_spec(8, width = "8em")
-```
-
-# Results
-
-There were `r nrow(d_all) |> scales::number(big.mark = ",")` records in the dataset for the three-year study period, with a strong upward trend, as shown in the monthly counts presented in @fig-time.
-Since early 2022, there have been more monthly records in the OpSnap data than in the official 'STATS19' road traffic collision records for West Yorkshire, highlighting the under-reporting of road traffic incidents in official statistics.
-STATS19 records are from the Department for Transport's database of road traffic collisions reported to the police and only include incidents that result in injury.
-Like OpSnap data, STATS19 records are open access.
-For the results presented in @fig-time, STATS19 datasets were downloaded with the `stats19` R package [@lovelace2019] and filtered to include only records from West Yorkshire.
-
-```{r}
-#| label: fig-time
-#| fig-cap: "Monthly count of Operation Snap (complete and with offence and location data, red and green) and official STATS19 road traffic collision records (blue), West Yorkshire."
-d_monthly = bind_rows(d_monthly, stats19_monthly)
-d_monthly |>
-  mutate(records = case_when(
-    records == "all" ~ "OpSnap (all)",
-    records == "complete" ~ "OpSnap (complete)",
-    records == "stats19" ~ "STATS19"
-    )
-  ) |>
-  rename_all(snakecase::to_title_case) |> 
-  ggplot() +
-  geom_line(aes(Month, N, colour = Records), alpha = 0.5, size = 2) +
-  # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
-  labs(
-    # title = "Number of monthly records in West Yorkshire Police\nOperation Snap data",
-       x = "Date",
-       y = "Number of records per month") +
-  theme_minimal()
-```
-
-Some records lacked either an offence (`r format(nrow(d_all) - nrow(d_offence), big.mark = ",", scientific = FALSE)`, `r round((nrow(d_all) - nrow(d_offence)) / nrow(d_all) * 100, 1)`%) or a location (`r format(nrow(d_offence) - nrow(d_complete), big.mark = ",", scientific = FALSE)`, `r round((nrow(d_offence) - nrow(d_complete)) / nrow(d_all) * 100, 1)`%), or both, leaving `r round(nrow(d_complete) / nrow(d_all) * 100, 1)`% or `r format(nrow(d_complete), big.mark = ",", scientific = FALSE)` complete records.
-There was a distinct seasonal pattern to reporting, with significant increases in summer months.
-
-A summary of the `r nrow(d_offence) |> scales::number(big.mark = ",")` records with an offence is presented in @tbl-offences.
-Included in the table are the number and percentage of records by offence type, showing the top 6 offence types and the remainder grouped as ‘Other’.
-
-The most common offences were ‘Driving without reasonable consideration to others (rt88576)’ and ‘Driving without due care and attention (rt88575)’.
-Within the Road Traffic Act these offences are related to careless driving and drivers are subject to similar penalties.
-The other common offences included failing to comply with traffic signals, traffic signs and using a handheld phone while driving.
-<!-- TODO: Within the other category... -->
-
-```{r}
-#| label: tbl-offences
-#| include: true
-#| tbl-cap: "Offence types reported (top 6 and other)."
-#| width: 60%
-# Aim: get table of n. offences by mode
-d_mode_offence_count = d_all |>
-  count(mode, offence, sort = TRUE) 
-# offences in order of n. offences
-d_offence_count = d_all |>
-  count(offence, sort = TRUE)
-# d_offence_count |>
-#   arrange(desc(n)) |>
-#   head(20) |>
-#   knitr::kable()
-# |offence                                                                                                |    n|
-# |:------------------------------------------------------------------------------------------------------|----:|
-# |n/a                                                                                                    | 5706|
-# |rt88576 drive without reasonable consideration to others                                               | 4992|
-# |rt88575 drive without due care and attention                                                           | 2917|
-# |rt88975 drive motor vehicle fail to comply with red / green arrow / lane closure traffic light signals | 1364|
-# |rt88971 fail to comply with red traffic light                                                          |  679|
-# |rt88966 motor vehicle fail to comply with endorsable s36 traffic sign                                  |  411|
-# |rv86019 use a handheld phone / device whilst driving a motor vehicle on a road                         |  357|
-# |rt88760 fail to comply with solid white lines                                                          |  265|
-# |rt88751 contravene give way sign                                                                       |  264|
-# |suspected contravene weight restriction.                                                               |  213|
-# |rt88751 contravene mandatory direction arrows                                                          |  212|
-# |me82009 driving on hard shoulder of motorway                                                           |  113|
-# |rt88975 fail to comply with red traffic light                                                          |  109|
-# |rt88751 motor vehicle fail to comply with a non-endorsable traffic sign other (specify)                |   91|
-# |zp97004 fail to comply with red light pelican crossing                                                 |   84|
-# |zp97003 stop within controlled area of pelican crossing                                                |   81|
-# |rc86814 driver not in proper control of vehicle                                                        |   62|
-# |zp97001 stop vehicle within limits of pelican crossing                                                 |   59|
-# |hy35001 drive/ride on footpath beside a road                                                           |   42|
-# |rr84171 vehicle contravene local traffic order other than parking (e.g. bus lane)                      |   41|
-# Pull out the top 6 offences excluding n/a:
-d_offence_top_6 = d_mode_offence_count |>
-  # filter(offence != "n/a") |>
-  group_by(offence) |>
-  summarise(n = sum(n)) |>
-  arrange(desc(n)) |>
-  head(6)
-d_offence_classified = d_offence_count |>
-  mutate(
-    Offence = case_when(
-      offence %in% d_offence_top_6$offence ~ offence,
-      TRUE ~ "Other"
-    )
-  ) |>
-  group_by(Offence) |>
-  summarise(`Number of records` = sum(n)) |>
-  # Arrange in descending order of n except for "Other":
-  arrange(Offence == "Other", desc(`Number of records`)) |>
-  # Rename "n/a" to "No offence":
-  mutate(Offence = case_when(
-    Offence == "n/a" ~ "NA No offence or unknown offence type",
-    TRUE ~ Offence
-  )) |>
-  mutate(`Percent of records` = round(`Number of records` / sum(`Number of records`) * 100, 1))
-
-d_offence_totals = d_offence_classified |>
-  summarise(`Number of records` = sum(`Number of records`), `Percent of records` = sum(`Percent of records`)) |>
-  mutate(Offence = "Total")
-
-# d_offence_classified |>
-#   knitr::kable()
-
-tbl = d_offence_classified |>
-  bind_rows(d_offence_totals) |>
-  mutate(
-    `Percent of records`= round(`Percent of records`)
-    ) |> 
-  rename(`Number` = `Number of records`, `Percent`=`Percent of records`) |> 
-  gt() |>
-
-  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
-  text_transform(
-    fn <- function(x){
-      code <- str_extract(x, "^[^\ ]+")
-      desc <- str_remove(x, "^[^\ ]+")
-      glue::glue("<em><span style='font-size:14px'>{code}</span></em><br><span style='font-size:18px'>{desc}</span>")
-      },
-    locations=cells_body(columns=Offence)
-  ) |> 
-  cols_width(1 ~ px(600), 2 ~ px(80)) |> 
-  gt_theme_espn()
-# gt::gtsave(tbl, "tbl-offences.html")
-# webshot2::webshot("tbl-offences.html", "tbl-offences.png")
-# browseURL("tbl-offences.png")
-knitr::include_graphics("tbl-offences.png")
-```
-
-```{r}
-# reclassify offences into types
-x = d_all
-regroup_offences_simple = function(x) {
- x |> 
-   dplyr::mutate(
-     offence_simple = dplyr::case_when(
-       stringr::str_detect(offence, "drive without reasonable consideration") ~ "Inconsiderate driving",
-       stringr::str_detect(offence, "drive without due care") ~ "Careless driving",
-       TRUE ~ "Other"
-     )
-   )
-}
-# table(x$mode)
-regroup_modes = function(x) {
-  x |> 
-    dplyr::mutate(
-      mode_simplified = dplyr::case_when(
-      mode == "cyclist" ~ "Cyclist",
-      mode == "vehicle driver" ~ "Driver",
-      TRUE ~ "Other"
-    ))
-}
-
-d_all = regroup_offences_simple(d_all)
-d_all = regroup_modes(d_all)
-
-```
-
-Vehicle (mostly car and van drivers with dashcams) driver and cyclist reporters dominate reporting for all records, as illustrated in @tbl-mode.
-Half of the cases were reported by vehicle drivers, a third by cyclists, seven percent by pedestrians, with over two percent by horse riders and less than one percent by motorcyclists.
-
-```{r}
-#| label: tbl-mode
-#| tbl-cap: Mode of transport of person submitting video evidence
-#| width: "80%"
-# d_all |>
-#   count(mode, sort = TRUE) |>
-#   mutate(percent_of_records = n / nrow(d_all)) |>
-#   mutate(percent_of_records = round(percent_of_records, 3) * 100) |>
-#   arrange(desc(n)) |>
-#   rename_all(snakecase::to_title_case) |>
-#   # Rename N to "Number of records"
-#   rename(`Number of Records` = N) |>
-#   knitr::kable()
-
-tbl = d_all |>
-  # NA is Unknown:
-  mutate(mode = case_when(
-    is.na(mode) ~ "unknown",
-    TRUE ~ mode
-  )) |>
-  count(mode, sort = TRUE) |>
-  mutate(percent_of_records = n / nrow(d_all)) |>
-  mutate(percent_of_records = round(percent_of_records, 3) * 100) |>
-  # Reclassify Unknown to unknown:
-  arrange(desc(n))
-tbl_totals = tbl |>
-  summarise(n = sum(n), percent_of_records = sum(percent_of_records)) |>
-  mutate(percent_of_records = round(percent_of_records)) |>
-  mutate(mode = "Total")
-tbl = bind_rows(tbl, tbl_totals) |>
-  rename_all(snakecase::to_title_case) |>
-  # Rename N to "Number of records"
-  rename(`Number` = N, `Percent` = `Percent of Records`) |>
-  gt() |>
-  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
-  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
-  tab_style(
-    style = cell_text(size=px(18)),
-    locations = cells_body(columns = Mode)
-  ) |>
-  gt_theme_espn()
-# tbl
-# gt::gtsave(tbl, "tbl-mode.html")
-# webshot2::webshot("tbl-mode.html", "tbl-mode.png")
-knitr::include_graphics("tbl-mode.png")
-```
-
-<!-- The equivalent table excluding records with missing offence data is shown below: -->
-
-```{r}
-#| include: false
-d_all |>
-  opsnap:::filter_offence_nas() |>
-  filter(offence != "n/a") |>
-  count(offence, sort = TRUE) |>
-  mutate(percent_of_records = n / nrow(d_offence)) |>
-  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
-  arrange(desc(n)) |>
-  head(10) |>
-  rename_all(snakecase::to_title_case) |>
-  knitr::kable()
-```
-
-<!-- For cases submitted by people riding cycles (shown in @tbl-offences-cyclist-observer), the most common offences were also both associated with careless driving, particularly driving without reasonable consideration to others (78.7%), there were also a small proportion of cases associated with drivers using mobile phones (3.7%), failing to comply with traffic signals (3.5%) and contravening regulator signage (0.6%).
--->
-
-```{r}
-# # Offence (grouped)       |               | Total |
-# |-------------------------|---------------|-------|
-# |                         | Dangerous driving | Other offences | 
-# |-------------------------|-------------------|-----------------|
-# | **Reporter transport mode** |                   |                 |
-# | Vehicle driver          | **Count**           | **5511**        | **654** | **6165** |
-# |                         | % within Reporter transport mode | **89.4%** | **10.6%** | **100.0%** |
-# | Cyclist                 | **Count**           | **4469**        | **249** | **4718** |
-# |                         | % within Reporter transport mode | **94.7%** | **5.3%**  | **100.0%** |
-# | Other                   | **Count**           | **1279**        | **220** | **1499** |
-# |                         | % within Reporter transport mode | **85.3%** | **14.7%** | **100.0%** |
-# | **Total**               | **Count**           | **11259**       | **1123** | **12382** |
-# |                         | % within Reporter transport mode | **90.9%** | **9.1%**  | **100.0%** |
-
-# R version:
-pivot_counts = d_all |>
-  filter(offence != "n/a") |>
-#   mutate(
-#     offence_simple = case_when(
-#       offence %in% d_offence_top_6$offence ~ offence,
-#       TRUE ~ "Other"
-#   )
-# ) |>
-  group_by(mode_simplified, offence_simple) |>
-  summarise(n = n()) |>
-  pivot_wider(names_from = mode_simplified, values_from = n) |> 
-  mutate(
-    Total = Cyclist + Driver + Other
-  ) |>
-  arrange(offence_simple == "Other", desc(Total))
-pivot_counts_total = tibble::tibble(
-  offence_simple = "Total",
-  Cyclist = sum(pivot_counts$Cyclist),
-  Driver = sum(pivot_counts$Driver),
-  Other = sum(pivot_counts$Other),
-  Total = sum(pivot_counts$Total)
-)
-
-pivot_percents = pivot_counts |>
-  # Calculate percentage of each mode
-  mutate(
-    `Cyclist (%)` = Cyclist / sum(Cyclist) * 100,
-    `Driver (%)` = Driver / sum(Driver) * 100,
-    `Other (%)` = Other / sum(Other) * 100,
-    `Total (%)` = Total / sum(Total) * 100
-  )
-
-pivot_percents_totals = pivot_percents |>
-  group_by(offence_simple = "Total") |>
-  summarise_if(is.numeric, sum)
-
-pivot_combined = bind_rows(pivot_percents, pivot_percents_totals)
-
-percent_careless = pivot_combined |> filter(offence_simple == "Careless driving") |> pull(`Total (%)`)
-percent_inconsiderate = pivot_combined |> filter(offence_simple == "Inconsiderate driving") |> pull(`Total (%)`)
-percent_careless_cycling = pivot_combined |> filter(offence_simple == "Careless driving") |> pull(`Cyclist (%)`)
-percent_inconsiderate_cycling = pivot_combined |> filter(offence_simple == "Inconsiderate driving") |> pull(`Cyclist (%)`)
-percent_other = pivot_combined |> filter(offence_simple == "Other") |> pull(`Total (%)`)
-percent_other_cyclist = pivot_combined |> filter(offence_simple == "Other") |> pull(`Cyclist (%)`)
-percent_other_driver = pivot_combined |> filter(offence_simple == "Other") |> pull(`Driver (%)`)
-percent_other_other = pivot_combined |> filter(offence_simple == "Other") |> pull(`Other (%)`)
-
-```
-
-A cross-tabulation of transport mode (of the person submitting video evidence) and the offence type is shown in @tbl-mode-offences-crosstab.
-For both variables it shows the two main categories plus an 'other' category.
-
-While `r round(percent_inconsiderate, 1)` percent of all offences are for driving without reasonable considered to others (rt88576), they make up the bulk of offences reported by cyclists (`r round(percent_inconsiderate_cycling, 1)`%).
-Drivers are proportionally `r round(percent_other_driver / percent_other_cyclist, 1)` times more likely to report other types of offences as cyclists, while other reporting modes are most likely to report other types of offences, being `r round(percent_other_other / percent_other_cyclist, 1)` times more likely to report other types of offences as cyclists.
-While further research is needed to understand the reasons for these tendencies, the results match intuition.
-Physically-vulnerable cyclists are understandably most concerned with the dangerous driving of vehicles, whereas drivers tend to focus on other types of road traffic offence.
-Pedestrians and other reporting modes were also relatively more likely to encounter other types of offence.
-
-```{r}
-#| label: tbl-mode-offences-crosstab
-#| tbl-cap: Mode of transport of person submitting video (columns).
-# pivot_combined |>
-#   knitr::kable(digits = 1)
-tbl = pivot_combined |>
-  select(-`Total (%)`) |> 
-  filter(offence_simple!="Total") |> 
-  select(-c(Cyclist:Other)) |> 
-  pivot_longer(cols=`Cyclist (%)`:`Other (%)`, names_to="mode", values_to="perc") |> 
-  mutate(
-    mode=factor(str_extract(mode, "^[^\ ]+"), levels=c("Driver", "Cyclist", "Other")),
-    offence_simple=factor(offence_simple, levels=c("Inconsiderate driving", "Careless driving", "Other"))
-    ) |> 
-  group_by(offence_simple) |> 
-  summarise(perc=list(round(perc)), count=first(Total)) |> 
-  gt() |>
-  gt_plt_bar_stack(perc, width=65, labels = c(" Cyclists ", " Drivers ", " Other "),
-                   palette= c("#e31a1c", "#1f78b4", "#bdbdbd")) |>
-    cols_width(1 ~ px(180), 2 ~ px(400), 3 ~ px(150)) |> 
-  tab_style(
-    style = cell_text(size=px(18)),
-    locations = cells_body(columns = c(offence_simple, count))
-    ) |> 
-    gt_theme_espn() 
-# gt::gtsave(tbl, "tbl-mode-offences-crosstab.png")
-knitr::include_graphics("tbl-mode-offences-crosstab.png")
-```
-
-```{r}
-#| label: tbl-offences-cyclist-observer
-#| tbl-cap: "Number and percentages of OpSnap records, submitted by cyclists, by offence type."
-#| include: false
-d_all |>
-  opsnap:::filter_offence_nas() |>
-  filter(offence != "n/a") |>
-  filter(mode == "cyclist") |>
-  count(offence, sort = TRUE) |>
-  mutate(percent_of_records = n / nrow(d_offence)) |>
-  mutate(percent_of_records = scales::percent(round(percent_of_records, 3))) |>
-  mutate(
-    offence = ifelse(n < 20, "other", offence)
-  ) |>
-  group_by(offence) |>
-  summarise(n = sum(n), n_hybrid = n()) |>
-  arrange(n_hybrid, desc(n)) |>
-  select(-n_hybrid) |>
-  mutate(`% of total` = scales::percent(n / sum(n), accuracy = 0.1)) |>
-  rename_all(snakecase::to_title_case) |> 
-  knitr::kable()
-
-perc_no_further_action = d_all |>
-  count(disposal) |>
-  filter(disposal == "nfa") |>
-  pull(n) / nrow(d_all) * 100
-
-perc_course = d_all |>
-  count(disposal) |>
-  filter(disposal == "course") |>
-  pull(n) / nrow(d_all) * 100
-```
-
-Disposal categories assigned by police are shown in @tbl-disposal.
-Roughly a third of cases (`r round(perc_no_further_action, 1)`%) resulted in no further action and, for most of the remainder drivers were required to undertake an education course.
-Conditional offers, that is, drivers being offered a reduced penalty for admitting guilt, were the third most common outcome.
-Nearly two percent of cases went to court and a further one percent underwent further investigation.
-
-```{r}
-#| label: tbl-disposal
-#| tbl-cap: "Most common disposal values in the OpSnap dataset."
-# d_all |>
-#   count(disposal, sort = TRUE) |>
-#   mutate(percent_of_records = round(n / nrow(d_all), 3) * 100) |>
-#   arrange(desc(n)) |>
-#   rename_all(snakecase::to_title_case) |> 
-#   knitr::kable()
-
-tbl_disposal = d_all |>
-  count(disposal, sort = TRUE) |>
-  mutate(percent_of_records = round(n / nrow(d_all), 3) * 100) |>
-  arrange(desc(n))
-
-tbl_disposal_totals = tbl_disposal |>
-  summarise(n = sum(n), percent_of_records = round(sum(percent_of_records))) |>
-  mutate(disposal = "Total")
-
-tbl = bind_rows(tbl_disposal, tbl_disposal_totals) |>
-  rename_all(snakecase::to_title_case) |>
-  # Rename N to "Number of records"
-  rename(`Number` = N, `Percent` = `Percent of Records`) |>
-  gt() |>
-  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
-  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
-  tab_style(
-    style = cell_text(size=px(18)),
-    locations = cells_body(columns = Disposal)
-  ) |>
-  gt_theme_espn()
-# gt::gtsave(tbl, "tbl-disposal.html")
-# webshot2::webshot("tbl-disposal.html", "tbl-disposal.png")
-knitr::include_graphics("tbl-disposal.png")
-```
-
-There were `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations corresponding to busy roads: Meanwood Road (Leeds), Dewsbury Road (Wakefield) and Chapeltown Road (Leeds), with no single address accounting for more than 0.4% of records.
-The locations were scrambled by West Yorkshire Police before being made available for data protection purposes, which meant spatial analysis would only possible for ‘all’ video submissions, ignoring different types of road users or offence types.
-
-```{r}
-#| echo: false
-#| eval: false
-#| label: tbl-locations
-#| tbl-cap: "Most common locations recorded in the OpSnap dataset"
-# d_with_location |>
-#   count(location, sort = TRUE) |>
-#   mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
-#   arrange(desc(n)) |>
-#   head(10) |>
-#   rename_all(snakecase::to_title_case) |> 
-#   knitr::kable()
-tbl = d_with_location |>
-  count(location, sort = TRUE) |>
-  mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
-  arrange(desc(n)) |>
-  head(10) |>
-  rename_all(snakecase::to_title_case) |>
-  rename(`Number` = N, `Percent` = `Percent of Records`) |>
-  gt() |>
-  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
-  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
-  tab_style(
-    style = cell_text(size=px(18)),
-    locations = cells_body(columns = Location)
-  ) |>
-  gt_theme_espn()
-# gt::gtsave(tbl, "tbl-locations.png")
-# knitr::include_graphics("tbl-locations.png")
-```
-
-# Discussion
-
-Drawing on the work and findings described above, this discussion outlines a research agenda.
-The breadth and depth of potential policy-relevant work that could be undertaken is, we suggest, enormous.
-While this preliminary list will not prove exhaustive, if it stimulates or informs further research then it will have achieved its objective.
-
-There has been an increase in submissions to West Yorkshire Police's OpSnap system since it was set-up in 2021.
-This may reflect changing levels of road use during and after the Covid-19 pandemic, increased awareness of OpSnap, or increased video camera ownership levels.
-Comparative analysis including other police jurisdictions should be undertaken to determine whether this experience is common or isolated.
-We hypothesise that it is a common experience for police forces (and other organisations) setting up public video submissions reporting antisocial and dangerous driving and that that the rate of submissions will continue to increase (longer term, ideally, submission rates decline as roads become safer).
-
-This was a case study of West Yorkshire but each of the 43 police forces in England and Wales collates OpSnap data.
-Our preliminary scoping of individual force websites suggests OpSnap data is publicly available nationwide.
-A national dataset should be developed that includes OpSnap data from all police forces.
- A feasibility study should be undertaken to establish cross-force data availability and compatibility.
-A national dataset holds potential for national-level, cross-regional, and comparative analyses of patterns and trends.
-It would have the potential to promote cross-national comparative analysis and international cooperation in road safety should similar data be collated elsewhere.
-It would hold the potential for the development of rankings according to different criteria and, thereby, potential performance-related metrics [@tiwana2015].
-There will be added-value from a national dataset which allows identification of cross-jurisdictional issues such as the same vehicles or persons involved in incidents in different police force areas.
-
-With submission volumes likely to continue to increase, means of informing triage and prioritisation of cases will be increasingly valuable.
-Triage systems are almost certainly used already, and research to identify best practice should be undertaken.
-Similarly, research to determine best practice in determining disposal recommendations should be undertaken.
-It may provide feasible, perhaps using machine learning, to develop automated means of triage, determination of offence type and disposal.
-Research to gauge the differential effectiveness of disposal measures is also needed.
-Within the trend of increasing video submissions there were distinct seasonal patterns.
-Seasonal variation in submissions reflect seasonal (weather and other) influences, and these will vary by transport mode: cycling, for example, is less prevalent in winter.
-This suggests seasonal variation in preventive responses might be tailored to need.
-
-Future research should focus on different types of problem concentration.
-It is well established that crime is highly concentrated along whatever dimension is examined [@farrellPreventingRepeatRepeat2017].
-With respect to video evidence this will include recidivists (repeat offenders), repeat submitters of evidence (who may or may not be repeat victims/survivors), the concentration of incidents (close passes, near misses and crashes) at certain times and places, with different types of experiences concentrated among certain types of road users.
-As discussed in the introduction, close passes are more likely to be reported by cyclists and horse riders than vehicle drivers, both due to the nature of the road user interactions and the relative risk associated with the maneuver.
-
-There was preliminary confirmation of this pattern in our analysis which found that offence types reported by vehicle drivers are systematically different to those of other road users.
-Hence future research should consider that investigative and preventive approaches should be tailored to different contexts and types of road user.
-Rural roads account for a disproportionate amount of fatalities @brakeDirectLineBrake2018, and further research using video evidence may inform preventive approaches tailored to road type.
-Policing and prevention efforts focused on where problems are concentrated are more resource efficient and is the foundation of problem-oriented policing [@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010].
-
-Studies focused on specific types of road user will prove informative with respect to policy and practice.
-Dash-cam submissions by motorists offers great potential to inform other aspects of road safety.
-Research into submissions by pedestrians, horse riders and motorcyclists should prove viable to improve the safety of these parties.
-There were over four hundred submissions by horse riders in West Yorkshire, which means there will be thousands nationwide.
-The British Horse Society[^1] and others may be interested in this data being used to inform the safety of horses and riders.
-The British Motorcyclists Federation[^2] may be interested in research promoting the safety of its members and in why motorcyclist submissions are relatively infrequent
-.
-
-[^1]: https://www.bhs.org.uk/
-
-[^2]: https://www.britishmotorcyclists.co.uk/
-
-As an example, cyclists undertake two percent of road miles in the UK [@departmentfortransport].
-Other things equal, this suggests that with respect to distance travelled, cyclists were around 20 times more likely to submit evidence of a road traffic offence than vehicle drivers and other road users in West Yorkshire.
-
-Over 90 percent of offences reported by cyclists were for driving without consideration for other road users and without due care and attention.
-This is consistent with the phenomenon of 'close passes' in the cycling safety literature [@aldred2016; @cubbin2024].
-It suggests the OpSnap data holds the potential for further analysis to inform knowledge about the nature of close passes generally.
-For example, it should be possible to identify hotspot locations.
-As with much of the further research outlined here, this the potential to inform preventive policy and practice relating to police interventions, driver behaviour, the design of roads and roadside environments.
-
-Comparison by complainant mode of transport, reported offense type and disposal, subdivided by other characteristics, is needed.
-Determining which type of submission - by which type of road user for example — is more likely to result in a recommendation of court proceedings, may help refine police investigations.
-For example, are submissions by cyclists (or horse riders) more or less likely to result in court proceedings than those by motorists?
-If so, why?
-What is the role of substantive issues, and to what extent are decisions affected by video quality?
-Further work with police investigators should inform best practice in the processing and further investigation of submissions.
-How are disposal options determined?
-Cross-police force comparative analysis may inform national best practice guidelines.
-
-A constraint on the present analysis was the nature of the publicly-available data.
-A pilot project should be undertaken in collaboration with police to establish the potential to further enhance policing and public safety using non-public aspects of the submissions data.
-This would require working partnerships and a secure research platform to meet GDPR requirements and Data Protection Act (2018).
-Such collaborative approaches are increasingly common in health, medicine, and policing research.
-
-Video footage holds significant potential for further analysis, both qualitative and quantitative.
-Police investigations might be improved by research to identify and promote good practice in the assessment of footage, its use in determining disposals, further investigations, and prosecutions.
-What type of footage works best in the courtroom, and how is it best identified and prepared?
-What is the potential for machine learning to identify, clean, and prepare footage of the most serious submissions?
-There are also likely to be lessons that can be learned for how footage is gathered, edited and submitted by road users.
-
-Analysis of video footage should be undertaken to identify risk factors, that is, the types of situations in which crashes, near misses, close passes and other offences occur.
-Such research can inform policy and practice in ways that ameliorate risk.
-Within this area of research, analysis from dash cams is likely to inform different practices than that from cycle-cams, that from pedestrians, horse riders and motorcyclists, and so on.
-
-Some research into the relationship between video footage and supporting written evidence is needed.
-Is written supported evidence always needed?
-Which is deemed most important by police, and which by courts?
-What are the characteristics of strong supporting written evidence, and what are the characteristics of strong video evidence?
-Do both need to be ‘strong’ or can a weakness in one be overcome by particularly strong aspects in the other?
-Research should identify further aspects of good practice for those submitting evidence to police.
-
-The spatial analysis undertaken here was obliged to use data for 'all' video submissions.
-We were unable to cross-tabulate the geographic location of submissions by different road users because, in the publicly-available data, locations are not be matched to individual cases: they were scrambled to ensure anonymity for GDPR purposes.
-An obvious next step, in the context of a secure research platform, would be spatial analysis for different types of road users, for different types of incidents, for incidents resulting in different disposals, and so on.
-Hotspots and spatial clustering are likely to vary by type of road user, type of incident, day or week, time of day, and so on.
-
-Some research into the relationship between video footage and supporting written evidence is needed.
-Is written supported evidence always needed?
-Which is deemed most important by police, and which by courts?
-What are the characteristics of strong supporting written evidence, and what are the characteristics of strong video evidence?
-Do both need to be ‘strong’ or can a weakness in one be overcome by particularly strong aspects in the other?
-Research should identify further aspects of good practice for those submitting evidence to police.
-
-We did not include spatial analysis here, but the data hold that potential.
-The public data ued here only allowed spatial analysis for ‘all’ video submissions.
-That is, we were unable to distinguish the geographic location of submissions by different types of road users because, or for different offence types, because the locations were scrambled for data protection purposes.
-Future research, in the context of a secure data platform, should include spatial analysis for different types of road users, for different types of incidents, for incidents resulting in different disposals.
-Hotspots and spatial clustering will vary by type of road user, type of incident, day or week, time of day, and so on, and parsing the data will produce more informed road safety research.
-
-Accessing OpSnap data in collaboration with police is likely to prove the most fruitful approach for future research.
-Were that impractical, research using existing online videos should be undertaken.
-There are many thousands of such videos on social media in the public realm, including those posted by police forces, and while there are different sampling issues to consider, this sources offers a plausible alternative route to informing road safety.
-
-OpSnap data should be compared to, and integrated with, other road use and road safety datasets.
-Here we offered preliminary comparison to the volume of Stats19 data on road crashes.
-patial analysis integrated with Strava data (road use data by cyclists and pedestrians), Google road use and other data, may facilitate exposure-based measures rather than the counts used here.
-The use of rates will enhance the identification of locations with higher risk other than that due to volume of traffic.
-  Future research should recognise that OpSnapdata holds the potential for use in the evaluation of experimental interventions.
-It may offer the potential for pre-post intervention comparative evaluations using control sites.
-Different road safety interventions imply different resource needs.
-Ideally, evaluations would include cost-benefit analysis of the portfolio of social and economic costs involved, including the cost of death and injury.
-
-# Conclusion
-
-This study offers what we believe is the first study using the open access Operation Snap data.
-It provides proof-of-concept that OpSnap data holds considerable research potential.
-OpSnap data has, we suggest, significant potential to inform policy and practice promoting road safety.
-Drawing on the preliminary analysis of the West Yorkshire data, a  research agenda was outlined.
-Our preliminary list will prove far from exhaustive.
-And while aspects of that agenda may appear ambitious at the time of writing, we suggest that the present study will have achieved one of its aims if this agenda is rapidly superseded.
-
-# Acknowledgements
-
-[To be added after peer review]
-
-<!--
-
-We would like to thank Ian King for making the connections that led to this paper being written.
-Thanks to Roger Beecham who wrote code for visualising some of the tables/figures in the paper, testing and supporting with reproducibility. 
-Thanks to Simon D'vali and colleagues from City of Bradford Metropolitan Council for their support and advice on the project and colleagues from the Safer Cycling Yorkshire group.
-
--->
-
-# References

### 2024-09-16 paper.qmd
Working docx

--- 2024-09-16 paper.qmd
@@
 #| eval: false
-#| label: tbl-locations
-#| tbl-cap: "Most common locations recorded in the OpSnap dataset"
+# #| label: tbl-locations
+# #| tbl-cap: "Most common locations recorded in the OpSnap dataset"
 # d_with_location |>
@@
 #   knitr::kable()
-tbl = d_with_location |>
-  count(location, sort = TRUE) |>
-  mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
-  arrange(desc(n)) |>
-  head(10) |>
-  rename_all(snakecase::to_title_case) |>
-  rename(`Number` = N, `Percent` = `Percent of Records`) |>
-  gt() |>
-  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
-  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
-  tab_style(
-    style = cell_text(size=px(18)),
-    locations = cells_body(columns = Location)
-  ) |>
-  gt_theme_espn()
+# if (!file.exists("tbl-locations.png")) {
+# tbl = d_with_location |>
+#   count(location, sort = TRUE) |>
+#   mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
+#   arrange(desc(n)) |>
+#   head(10) |>
+#   rename_all(snakecase::to_title_case) |>
+#   rename(`Number` = N, `Percent` = `Percent of Records`) |>
+#   gt() |>
+#   gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
+#   cols_width(1 ~ px(400), 2 ~ px(80)) |> 
+#   tab_style(
+#     style = cell_text(size=px(18)),
+#     locations = cells_body(columns = Location)
+#   ) |>
+#   gt_theme_espn()
 # gt::gtsave(tbl, "tbl-locations.png")
+# } else {
 # knitr::include_graphics("tbl-locations.png")
+# }
+# knitr::kable(cars)
 ```
